---
title: "archive plots"
author: "AMP"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Figure 1: Year 1 DOC 
Main Message 1 (Hypothesis rejected): There is a prolonged and sustained increase in dissolved carbon following initial drop in DOC after exposure to salinity.

EDIT to make the fw and c plots as bands. Then show the spatial variability in the sw plot by plotting the individual lysimeters (numbers inside circles for the points?)
straight range stripe for FW and C plots
95 and 5 (25 /75) quartile lines

```{r plot DOC}

figure_legend_plot <- pw_tmp1 %>%
  ggplot(aes(x = collection_datetime, y = doc_mg_l, color = plot)) +
  geom_point()+
  scale_color_manual(values=Anyas_colors) +
  theme(legend.position = "right")

legend <- get_legend(figure_legend_plot)

Figure1_content <- pw_tmp1 %>%
  filter(plot == "Saltwater") %>%
  filter(between(evacuation_date, startstudydate_yr1, endstudydate_yr1)) %>%
    mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = collection_datetime, y = doc_mg_l, color = plot, label= grid)) +
    geom_vline(xintercept = WaterDeliveryStart2022,linetype="dashed",color="grey")+ 
    #geom_vline(xintercept = WaterDeliveryStart1,linetype="dashed",color="grey")+ 
   # geom_vline(xintercept = WaterDeliveryStart2,linetype="dashed",color="grey")+
    annotate("rect", xmin = min(pw_tmp1$collection_datetime), xmax = max(WaterDeliveryStop2), ymin = doc_iqr$q1_doc_mg_l, ymax = doc_iqr$q3_doc_mg_l, fill= c("cyan2", "springgreen2"), alpha = 0.2)+
    geom_point(size = 5, shape = 21, fill = "violetred2", color = "black") +
    geom_text(size = 2, color = "white") +
    ylab("DOC mgC/L")+
    xlab("Date")+
    theme_classic() +
    theme(legend.position="none")

Figure1 <- cowplot::plot_grid(Figure1_content, legend, nrow= 1, rel_widths=c(1, 0.2))

print(Figure1)

plotly::ggplotly(Figure1_content)



cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/Figure1_pwdoc.png",Figure1, dpi=600)

```

```{r}
Figure2a <- pw_doc %>%
  filter(between(evacuation_date, startstudydate_yr1, endstudydate_yr1)) %>%
  mutate(year = lubridate::year(evacuation_date), 
         month = lubridate::month(evacuation_date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(evacuation_date) - 5) %% 12 ) %>%
  group_by(month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year)) %>%
  ungroup() %>%
    ggplot(aes(x = adj_month, y = doc_mg_l, color = plot, shape = year)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l-doc_mg_l_sd, ymax=doc_mg_l+doc_mg_l_sd), size = 1)+
    ylab("DOC mgC/L")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    xlim(0,12)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks()) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "bottom") +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  a) Field Year 1"), aes(x, y, label = label), hjust = -0.1, vjust = 1.1, inherit.aes = FALSE)

print(Figure2a)
```