---
title: "tmp paper graveyard"
author: "AMP"
date: "`r Sys.Date()`"
output: html_document
---


# Graveyard


## CEC from pre event soils

```{r cec}


soils_directory <- "https://docs.google.com/spreadsheets/d/1a-4Nt7jorO301bi3NF9ebGhwNx_hVjPG/edit?gid=456381663#gid=456381663"

directory= file.path(here::here() %>% dirname(), 'TEMPEST-1-Porewater/data_do_not_commit/')

file_path = file.path(directory,"UGA_soils_analysis_2020.xlsx")

googledrive::drive_download(soils_directory, path= file_path, overwrite = TRUE)
```





### Prep data for figure 1: Porewater DOC averages/std by plot and month
Generate for easy comparison between event years and with laboratory data sets.
Can simplify this from previous iterations (see `graveyard_analysis_code.Rmd`) because we just want a monthly timeframe. 
```{r doc avg std by month YEAR 1}

pw_doc_yr1 <- doc_l1_2022 %>%
  mutate(date = as_date(date)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr1, endstudydate)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  # fix the sampling dates in June before the event to the month before
  mutate(adj_month = case_when(adj_month == 1 & date < EventStart2022 ~ adj_month == 0,
                               TRUE ~ adj_month)) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

pw_doc_yr1_2023 <- doc_l1_2023 %>%
  mutate(date = as_date(date)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr1, endstudydate_yr1)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  mutate(adj_month = case_when(adj_month == 0 ~ 12,
                               TRUE ~ adj_month )) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

pw_doc_y1_all <- pw_doc_yr1 %>%
  full_join(pw_doc_yr1_2023) %>%
  mutate(month = case_when(adj_month <= 7 ~ adj_month + 5,
                           adj_month > 7 ~ adj_month - 7))
  

saveRDS(pw_doc_y1_all, "~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP1_Figure1.rds")
```

```{r doc avg std by month YEAR 2}


pw_doc_yr2 <- doc_l1_2023 %>%
  mutate(date = as_date(evacuation_datetime)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr2, endstudydate_yr2)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  # fix the sampling dates in June before the event to the month before
  mutate(adj_month = case_when(adj_month == 1 & date < EventStart2023 ~ adj_month == 0,
                               TRUE ~ adj_month)) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

pw_doc_yr2_2024 <- doc_l1_2024 %>%
  mutate(date = as_date(date)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr2, endstudydate_yr2)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  mutate(adj_month = case_when(adj_month == 0 ~ 12,
                               TRUE ~ adj_month )) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

pw_doc_y2_all <- pw_doc_yr2 %>%
  full_join(pw_doc_yr2_2024) %>%
  mutate(month = case_when(adj_month <= 7 ~ adj_month + 5,
                           adj_month > 7 ~ adj_month - 7))

  

saveRDS(pw_doc_y2_all, "~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP2_Figure1.rds")
```

```{r doc avg std by month YEAR 3}


pw_doc_yr3 <- doc_l1_2024 %>%
  mutate(date = as_date(evacuation_datetime)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr3, endstudydate_yr3)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  # fix the sampling dates in June before the event to the month before
  mutate(adj_month = case_when(adj_month == 1 & date < EventStart2023 ~ adj_month == 0,
                               TRUE ~ adj_month)) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

  
pw_doc_yr3_2025 <- doc_l1_2025 %>%
  mutate(date = as_date(date)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr3, endstudydate_yr3)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  mutate(adj_month = case_when(adj_month == 0 ~ 12,
                               TRUE ~ adj_month )) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

pw_doc_y3_all <- pw_doc_yr3 %>%
  full_join(pw_doc_yr3_2025) %>%
  mutate(month = case_when(adj_month <= 7 ~ adj_month + 5,
                           adj_month > 7 ~ adj_month - 7))

  

saveRDS(pw_doc_y3_all, "~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP3_Figure1.rds")
```

```{r group all doc}

pw_doc_all <- pw_doc_y1_all %>%
  full_join(pw_doc_y2_all) %>%
full_join(pw_doc_y3_all) %>%
  mutate(stamp =lubridate::my(interaction(month, year, sep = " "))) 
  

saveRDS(pw_doc_all, "~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_ALLYEARS_Figure1.rds")

```



```{r save doc precip}
pw_doc_all_wprecip <- pw_doc_all %>%
  left_join(precip, by = c("month", "year"))


write_csv(pw_doc_all_wprecip, "~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_allyrs_withprecip.csv")

```



# Experimental Data Analysis for mechanistic relationships
Workflow 
1) Exploratory Data Analysis: look at DOC vs driver variable relationships in the ColEx data 
2) Bring in a priori hypothesis-based knowledge from table of mechanisms to reduce variable numbers.  https://docs.google.com/spreadsheets/d/1DHOeCKPpyUJbPBAVBy6-vegGymVbECd7rN9bFWm7Y0M/edit?usp=sharing
3) Try Bayesian Networks to discover probabilistic causal structures with bnlearn 
  3a) scale data (try z-scores)
  3b) check multicollinearity with VIFs
  3c) include mechanistic constraints (e.g., Fe/OC influence DOC)
  
### Exploratory data analysis
```{r doc stats}

hist(doc_l1_pw_grid_2022_2026$doc_mg_l)

doc_comparison <- doc_l1_pw_grid_2022_2026 %>%
  filter(plot != "Saltwater") %>%
  mutate(logdoc = log(doc_mg_l),  #DOC is right skewed
  gridID = interaction(plot, grid, sep = "_"),
  month = lubridate::month(evacuation_date),
  year = lubridate::year(evacuation_date),
  dateID = interaction(month, year, sep = "_"))# create unique grid IDs (per plot)

hist(doc_comparison$logdoc)

doc_lme <- lme(
  logdoc ~ plot ,
  random = ~ 1 | gridID, # random effect of grids in the plot
  correlation = corAR1(form = ~ 1 | gridID), # repeated monthly observations are temporally autocorrelated
  data = doc_comparison 
)

summary(doc_lme)

plot(doc_lme)

qqnorm(resid(doc_lme))
qqline(resid(doc_lme))

#one factor model

EMM <- emmeans(doc_lme, ~ plot)

EMM

### pairwise comparisons
pairs(EMM)

```

```{r load ColEx data}
ColEx <- readxl::read_excel("~/GitHub/tempest_ionic_strength/Data/Exp 3 Column Experiment ColEx Data/ColEx_data_allreps_formatted.xlsx") %>% #drop initial for this
  filter(Timepoint != 0)

```

```{r exploratory data analysis}
library(ggcorrplot)

ColEx_corr <- ColEx %>%
  select(-Treatment, -Column, - Timepoint)

cor_matrix <- cor(ColEx_corr, use = "complete.obs")

ggcorrplot(cor_matrix)

```
There are no other variables that stick out than the ones we'd pick a priori that are highly correlated to DOC. 


```{r check lags}
# Check lagged correlations
library(tseries)
ccf(ColEx$pH, ColEx$doc_mgperL, lag.max = 12)
ccf(ColEx$conductivity_uSpercm, ColEx$doc_mgperL, lag.max = 12)

```
 Interpretation:
 - Check lag 0: Strong direct correlation between variable and DOC at the same time point.
 - Lags > 0: Positive correlations here mean variable changes precede DOC changes by the lag amount.
 - Lags < 0: Negative correlations here mean DOC changes precede variable changes by the lag amount.
 - Dashed horizontal lines = confidence bounds, significant correlations occur when values go beyond those lines.

This analysis might be really cool to do with the actual plot data and not the experiment- going to leave out of the model here. 

Narrow down to final variables


```{r check colinearity}
library(car)

# Fit a linear model to check VIF
lm_model <- lm(DOC ~ pH + Fe_OC_ratio + Conductivity, data = scaled_data)
car::vif(lm_model)

# General guideline: VIF > 10 indicates problematic multicollinearity
```


```{r do the Dynamic Bayesian Networks}

library(bnlearn)

# Whitelist specific relationships based on prior knowledge
whitelist <- data.frame(
  from = c("Fe_OC_ratio", "pH"),
  to = c("DOC", "DOC")
)

dag <- hc(data, whitelist = whitelist)
plot(dag)
```


```{r prelim plots}


cdom_indicies_divided_all %>%
    group_by(plot, timedate_bin) %>%
    dplyr::summarise(across(everything(), list(mean = ~mean(., na.rm=TRUE), sd = ~sd(., na.rm=TRUE))), .groups = "keep") %>%
   ungroup() %>%
  ggplot(aes(x=factor(timedate_bin, levels = c("Pre","Mid","Pre Cond Threshold", "Post Cond Threshold")), 
               y=A_T_mean)) +
  geom_col() +
  geom_errorbar(aes(ymin = A_T_mean - A_T_sd, ymax= A_T_mean + A_T_sd), width = 0.2) +
  facet_wrap( ~ plot)

cdom_indicies_divided_all %>%
    group_by(plot, timedate_bin) %>%
    dplyr::summarise(across(everything(), list(mean = ~mean(., na.rm=TRUE), sd = ~sd(., na.rm=TRUE))), .groups = "keep") %>%
   ungroup() %>%
  ggplot(aes(x=factor(timedate_bin, levels = c("Pre","Mid","Pre Cond Threshold", "Post Cond Threshold")), 
               y=SUVA254_mean)) +
  geom_col() +
  geom_errorbar(aes(ymin = SUVA254_mean - SUVA254_sd, ymax= SUVA254_mean + SUVA254_sd), width = 0.2) +
  facet_wrap( ~ plot) +
  xlab("Sampling Timeframe") +
  ylab("SUVA254")+
  theme_classic()

biplot(eemspca)

```

##



#graveyard 

### Supplemental Figure XX: Precip and DOC

```{r precip}
doc_precip <- read_csv("~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_allyrs_withprecip.csv")
precip <- read_csv("~/GitHub/TEMPEST-1-porewater/data/mean_monthly_precip_2006-2026.csv")
```

```{r plots doc rain}
top_y <- 100
bottom_y <- 0

max_precip <- max(doc_precip$mean_monthly_precip_cm, na.rm = TRUE)

scale_factor <- (top_y - bottom_y) / max_precip

precip_axis <- function(x) (top_y - x) / scale_factor

scale_factor

#note this only has precip when we have DOC data...
Figure1_combined <- doc_precip %>%
   ggplot(aes(x = stamp, y = doc_mg_l_mean, color = plot)) +
   geom_rect(aes(xmin = stamp -15,  xmax = stamp + 15, y = top_y - mean_monthly_precip_cm * scale_factor), ymax = top_y, inherit.aes = FALSE, fill = "skyblue", alpha = 0.6) +
   #geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab(" ")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    scale_y_continuous(limits = c(bottom_y, top_y),  name = "DOC (mg/L)", sec.axis = sec_axis(trans = precip_axis , name = "Mean Monthly Precipitation (cm)")
  ) +
    scale_color_manual(values=Anyas_colors) +
   # scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16),  axis.title.x = element_blank() )
          #, axis.title.x = element_blank(), axis.text.x = element_blank() ) 

Figure1_combined 

rain_doc <- doc_precip %>%
   ggplot() +
   geom_point(aes(x = stamp, y = doc_mg_l_mean, color= mean_monthly_precip_cm)) +
   facet_wrap(~plot)+
    theme_classic()

rain_doc

precip_study <- precip %>%
  filter(year > 2021)

   ggplot() +
   geom_col(aes(x = stamp, y =  mean_monthly_precip_cm * 100), data = precip_study) +
   geom_point(aes(x = stamp, y = doc_mg_l_mean, color = plot), data = doc_precip)+
         scale_color_manual(values=Anyas_colors) 



```



## SI 3 Option 2 pH vs DOC

```{r load ph year plot avgs}
ph_2 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_ph_porewater_TMP2.rds") %>%
  mutate(pH_sd = case_when(is.na(pH_sd) ~ 0,
                          TRUE ~ pH_sd))

ph_3 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_ph_porewater_TMP3.rds") %>%
  mutate(pH_sd = case_when(is.na(pH_sd) ~ 0,
                          TRUE ~ pH_sd))

ph_2n3 <- ph_2 %>%
  full_join(ph_3)
```

```{r merging}

ph_doc <- pw_doc_2 %>%
  full_join(pw_doc_3) %>%
  full_join(ph_2n3, by = c("adj_month", "year", "plot")) %>%
  select(-tdn_mg_l_mean, -tdn_mg_l_sd) %>%
  drop_na()
```

```{r plot ph vs doc }

SI3 <- ph_doc %>%
   filter(plot == "Saltwater") %>%
    ggplot(aes(x = pH_mean, y = doc_mg_l_mean)) +
    geom_pointrange(aes(xmin=pH_mean-pH_sd, xmax=pH_mean+pH_sd, ymin= doc_mg_l_mean - doc_mg_l_sd, ymax = doc_mg_l_mean + doc_mg_l_sd), shape=17)+ 
    ylab("porewater DOC")+
    xlab("porewater pH")+
    theme_classic()

SI3

cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/FigureS3.png",SI3, dpi=600)

```



```{r import doc data}

pw_doc_1 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP1_Figure1.rds") %>%
  mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd))

pw_doc_2 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP2_Figure1.rds") %>%
  mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd))

pw_doc_3 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP3_Figure1.rds") %>%
  mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd))

#pw_tmp1 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/doc_porewater_grids_TMP1_forfig1.rds")

#doc_iqr <- readRDS("~/GitHub/TEMPEST-1-porewater/data/doc_porewater_summarystats_forfig1.rds") %>%
#  mutate(collection_datetime = as.POSIXct("2022-05-18 15:59:00", tz= "EST" ),
#         doc_mg_l = NA)

```

```{r plot DOC}

Figure1a <- pw_doc_1 %>%
  group_by(adj_month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year)) %>%
  ungroup() %>%
   ggplot(aes(x = adj_month, y = doc_mg_l_mean, color = plot)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab(" ")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16),  axis.title.x = element_blank() )+
          #, axis.title.x = element_blank(), axis.text.x = element_blank() ) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  a) Field Year 1"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(Figure1a)

Figure1b <- pw_doc_2 %>%
  group_by(adj_month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year)) %>%
  ungroup() %>%
    ggplot(aes(x = adj_month, y = doc_mg_l_mean, color = plot)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab("DOC mgC/L")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1),
         text= element_text(size = 16) ) +
       #  , axis.title.x = element_blank(), axis.text.x = element_blank() ) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  b) Field Year 2"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(Figure1b)

Figure1c <- pw_doc_3 %>%
  group_by(adj_month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year)) %>%
  ungroup() %>%
    ggplot(aes(x = adj_month, y = doc_mg_l_mean, color = plot)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab(" ")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16), 
          axis.title.x = element_blank() ) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  c) Field Year 3"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(Figure1c)

Figure1b_LEGEND <- pw_doc_2 %>%
  group_by(adj_month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year)) %>%
  ungroup() %>%
    ggplot(aes(x = adj_month, y = doc_mg_l_mean, color = plot)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab("DOC mgC/L")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  b) Field Year 2"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

fig1_legend = get_legend(Figure1b_LEGEND)




 Figure1 <- ggarrange(Figure1a, Figure1b, Figure1c, nrow = 3, legend = "bottom", common.legend = TRUE, align = "hv")


 Figure1


 cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/Figure1.png",Figure1, dpi=600,
                   nrow = 3)

all_hands_fig <- ggarrange(Figure1a, Figure1b, Figure1c, nrow = 1, legend = "bottom", common.legend = TRUE, align = "hv")

all_hands_fig

```

```{r graveyard}
Figure1d <- exp_doc %>%
  mutate(Treatment = factor(Treatment, levels = Treatment_order)) %>%
    mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd)) %>%
    ggplot(aes(x = Wash, y = doc_mg_l_mean, color = Treatment)) +
    geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=16, size = 1)+
    ylab("DOC mgC/L")+
    xlab("Experimental Wash")+
   # title("Laboratory Simulation")+
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  d) Laboratory"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(Figure1d)

Figure1d_LEGEND <- exp_doc %>%
  mutate(Treatment = factor(Treatment, levels = Treatment_order)) %>%
    mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd)) %>%
    ggplot(aes(x = Wash, y = doc_mg_l_mean, color = Treatment)) +
    geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=16, size = 1)+
    ylab("DOC mgC/L")+
    xlab("Experimental Wash")+
   # title("Laboratory Simulation")+
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    theme_classic() +
    theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  d) Laboratory"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

fig1c_legend = get_legend(Figure1d_LEGEND)
```


```{r}
SI4b <- exp_nacl_dat %>%
   ggplot(aes(x = Wash, y = Fe_ppm, color = Treatment)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=Fe_ppm-Fe_ppm_stdev, ymax=Fe_ppm+Fe_ppm_stdev), shape=17, size = 1)+
    ylab("Fe mg/L")+
    xlab("Treatment Wash Number")+
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,6)) +
    scale_color_manual(values = PNWColors::pnw_palette("Starfish",6))+
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    geom_text(data = data.frame(x = -Inf, y = Inf, label = "   a) "), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE) +
    theme(
    legend.position = "none",
    legend.title = element_blank(),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_blank(),
    axis.title.x =element_blank(),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_text(size = axis_text_size),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )

print(SI5_a)

SI5_b <- exp_nacl_dat %>%
   ggplot(aes(x = Wash, y = Al_ppm, color = Treatment)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=Al_ppm-Al_ppm_stdev, ymax=Al_ppm+Al_ppm_stdev), shape=17, size = 1)+
    ylab("Al mg/L")+
    xlab("Treatment Wash Number")+
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,6)) +
  scale_color_manual(values = PNWColors::pnw_palette("Starfish",6))+
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
   geom_text(data = data.frame(x = -Inf, y = Inf, label = "   b) "), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE) +
    theme(
    legend.position = "bottom",
    legend.title = element_text(size = axis_title_size),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_text(size = axis_title_size),
    axis.title.x = element_text(size = axis_title_size),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_text(size = axis_text_size),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )

print(SI5_b)

SI5 <- ggarrange(SI5_a, SI5_b, nrow = 2, legend = "bottom", common.legend = TRUE, align = "hv")

SI5

 cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/SI_Figure5.png",SI5, dpi=600,
                  nrow = 2)
```



# Extended Data Figure 4:  a) DOC curves with ASW vs NaCl b) NaCl Fe and Al 

```{r nacl}

exp_nacl_dat <- readxl::read_excel("~/GitHub/tempest_ionic_strength/Data/Exp 2 Ionic Strength NaCl Data/TEMPEST_Ionic_Strength_NaCl_chem_L2.xlsx")  %>%
  mutate(Treatment =  as.factor(Treatment))

nacl_doc <- exp_nacl_dat %>%
  mutate(group = interaction(Exp_Type, Treatment, sep = "_")) %>%
  select(group, Wash, doc_mgperL, doc_mgperL_stdev) %>%
  rename(doc_mg_l_mean= doc_mgperL,
         doc_mg_l_sd = doc_mgperL_stdev)

exp_asw_dat <- read_csv("~/GitHub/tempest_ionic_strength/Data/Exp 1 Ionic Strength ASW Data/TEMPEST_Ionic_Strength_ASW_chem_L2.csv")

asw_doc <- exp_asw_dat %>%
  mutate(group = interaction(Exp_Type, Treatment, sep = "_")) %>%
  select(group, Wash, doc_mg_l_mean, doc_mg_l_sd)

batch_exps <- nacl_doc %>% full_join(asw_doc) %>%
   separate(group, into = c("Experiment", "Level"), sep = "_")

batch_exps_change <- batch_exps %>%
  group_by(Experiment, Level) %>%
  mutate(
    doc_change = doc_mg_l_mean -
      doc_mg_l_mean[Wash == 1]
  ) %>%
  ungroup() %>%
    mutate(
    Level = factor(
      Level,
      levels = c("0", "0.1", "1", "5", "25", "100")
    )
  )
```
```{r batch plots}
SI4a <- batch_exps_change %>%
  ggplot(aes(
    x = Wash,
    y = doc_change,
    color = Level,
    linetype = Experiment
  )) +
  scale_color_manual(values = PNWColors::pnw_palette("Starfish",6))+
  geom_line(linewidth = 0.7) +
  geom_point(size = 2)+
  geom_hline(yintercept = 0, color = "grey60", linewidth = 0.5)+
#  geom_pointrange(
 #   aes(
 #     ymin = doc_mg_l_mean - doc_mg_l_sd,
# #     ymax = doc_mg_l_mean + doc_mg_l_sd
 #   ),
#    size = 0.5
#  ) +
#  scale_shape_manual(values = c(
#    "NaCl" = 16,   # circle
#    "ASW"  = 15    # square
#  )) +
  geom_text(
    data = data.frame(x = -Inf, y = Inf, label = "  a) "),
    aes(x, y, label = label),
    hjust = -0.1,
    vjust = 2,
    inherit.aes = FALSE
  )+
labs(
  y = expression(Delta*"DOC relative to treatment mgC/L"),
  x = "Treatement Wash Number",
  color = "Treatment Type (% relative \n to full strength ASW)"
)+
  scale_x_continuous(
    breaks = pretty_breaks(n = 6),
    limits = c(1, 6)
  ) +
  theme_classic() +
    theme(
    legend.position = "bottom",
    legend.title = element_text(size = legend_text_size),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_text(size = axis_title_size),
    axis.title.x = element_text(size = axis_title_size),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_text(size = axis_text_size),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )

print(SI4a)

```

For example, b) in the batch experiments with NaCl as the only introduced ions, higher ionic strength results in delayed release of greater amounts of Al and Fe from the soils. 

Fe, Al
```{r icp nacl}
metals_long <- exp_nacl_dat %>%
  rename(
    Al_mean = Al_ppm,
    Al_sd   = Al_ppm_stdev,
    Fe_mean = Fe_ppm,
    Fe_sd   = Fe_ppm_stdev
  ) %>%
  pivot_longer(
    cols = c(Al_mean, Al_sd, Fe_mean, Fe_sd),
    names_to = c("Metal", ".value"),
    names_pattern = "(Al|Fe)_(mean|sd)"
  ) %>%
  mutate(
    Treatment = factor(
      Treatment,
      levels = c("0", "0.1", "1", "5", "25", "100")
    )
  )

metal_com <- metals_long %>%
  group_by(Metal, Treatment) %>%
  summarise(
    center_of_mass = sum(Wash * mean, na.rm = TRUE) /
                     sum(mean, na.rm = TRUE),
    .groups = "drop"
  )

```

```{r summary plot metals nacl}
SI4b <- ggplot(
  metal_com,
  aes(
    x = Treatment,
    y = center_of_mass,
    color = Metal,
    shape = Metal,
    group = Metal
  )
) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 3) +
  scale_shape_manual(values = c(16, 15)) +
  scale_y_continuous(
    limits = c(1, 6),
    breaks = 1:6
  ) +
  labs(
    x = "NaCl Treatment \n (% relative to full strength ASW)",
    y = "Mean timing of Fe or Al \n release (wash number)",
    color = NULL,
    shape = NULL
  ) +  geom_text(
    data = data.frame(x = -Inf, y = Inf, label = "  b) "),
    aes(x, y, label = label),
    hjust = -0.1,
    vjust = 2,
    inherit.aes = FALSE
  )+
scale_color_manual(
  values = c(
    "Al" = "#4F6D8A",
    "Fe" = "#8C4A2F"
  )
)+
   theme_classic() +
    theme(
    legend.position = "bottom",
    legend.title = element_text(size = legend_text_size),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_text(size = axis_title_size),
    axis.title.x = element_text(size = axis_title_size),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_text(size = axis_text_size),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )

SI4b

```

```{r si4 combined}

SI4_aligned_panels <- cowplot::align_plots(
  SI4a,
  SI4b,
  align = "v",
  axis = "lr"
)

SI_4 <- cowplot::plot_grid(
  SI4_aligned_panels[[1]],
  SI4_aligned_panels[[2]],
  ncol = 1,
  rel_heights = c(1, 1)
)

SI_4 

cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/SI_Figure4.png", SI_4 , base_width = 7,
  base_height = 7, dpi=600)

```
