---
title: "Porewater DOC Manuscript Analysis"
author: "AMP"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(lubridate)
library(tidyverse)
library(car)
library(emmeans)
library(lme4)
library(lmerTest)  
library(nlme)
library(ggplot2)    # For diagnostic plots
library(broom)      # For tidy model outputs


#double check your wd
getwd()

#clean out your environment 
rm(list = ls())
```

### Set the study dates

```{r event and study dates}
endstudydate = lubridate::as_date("2026-05-31")
startstudydate = lubridate::as_date("2022-05-01")

plotting_end = as.POSIXct("2025-05-31 11:59:59", tz = "EST")
plotting_start = as.POSIXct("2022-05-01 00:00:00", tz = "EST")

endstudydate_yr1 = lubridate::as_date("2023-05-31")
startstudydate_yr1 = lubridate::as_date("2022-05-01")

endstudydate_yr2 = lubridate::as_date("2024-05-31")
startstudydate_yr2 = lubridate::as_date("2023-05-01") 

endstudydate_yr3 = lubridate::as_date("2025-05-31")
startstudydate_yr3 = lubridate::as_date("2024-05-01") 

year1_start = lubridate::as_date("2022-01-01")
year1_stop = lubridate::as_date("2022-12-31")

year2_start = lubridate::as_date("2023-01-01")
year2_stop = lubridate::as_date("2023-12-31")

year3_start = lubridate::as_date("2024-01-01")
year3_stop = lubridate::as_date("2024-12-31")


## TMP 1 
EventStart2022 = lubridate::as_date("2022-06-22")

WaterDeliveryStart2022 = as.POSIXct("2022-06-22 05:30:00", tz = "EST")
WaterDeliveryStop2022 = as.POSIXct("2022-06-22 14:30:00", tz = "EST")

flood_event_datetime_1 <- c(WaterDeliveryStart2022, WaterDeliveryStop2022)

## TMP 2
EventStart2023 = lubridate::as_date("2023-06-06")

WaterDeliveryStart1 = as.POSIXct("2023-06-06 05:30:00", tz = "EST")
WaterDeliveryStop1 = as.POSIXct("2023-06-06 14:30:00", tz = "EST")

WaterDeliveryStart2 = as.POSIXct("2023-06-07 05:30:00", tz = "EST")
WaterDeliveryStop2 = as.POSIXct("2023-06-07 14:30:00", tz = "EST")

flood_event_datetime_2 <- c(WaterDeliveryStart1, WaterDeliveryStop2)

## TMP 3
EventStart2024 = lubridate::as_date("2024-06-11")

WaterDeliveryStart1_2024 = as.POSIXct("2024-06-11 06:10:00", tz = "EST")
WaterDeliveryStart2_2024 = as.POSIXct("2024-06-12 06:08:00", tz = "EST")
WaterDeliveryStart3_2024 = as.POSIXct("2024-06-13 06:21:00", tz = "EST")

WaterDeliveryStop3_2024 = as.POSIXct("2024-06-13 16:29:00", tz = "EST")

flood_event_datetime_3 <- c(WaterDeliveryStart1_2024, WaterDeliveryStop3_2024)

plot_order <- c('Control', 'Freshwater','Saltwater')
time_order <- c('Pre', 'Mid','Post')
```

### Read in DOC data

1) Import DOC grid data for porewater from 2022, 2023 and 2024 events. Then, bind them together and take the average and standard deviation of the values for each event. Do this by the evacuation date, as this will capture multiple tries within the same dateframe. For this particular analysis, it's ok to take the average of the date time of sampling, because we're just going to be using this data to plot through time and do some simple summary statistics.  

```{r read in doc}

#Porewater DOC data, from L1 google drive: 

doc_l1_2022 <- read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_GRIDSONLY_NPOCTDN_2022_L1.csv")
  
doc_l1_2023 <- read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_GRIDSONLY_NPOCTDN_2023_L1.csv")

doc_l1_2024 <- read_csv("~/GitHub/TEMPEST_Porewater/processing_scripts/DOC/TEMPORARY/01_TMP_PW_NPOCTDN_2024_L1_updated.csv")
# C_C6_DOC_20241003 still looks sus

doc_l1_2025 <- read_csv("~/GitHub/TEMPEST_Porewater/processing_scripts/DOC/TEMPORARY/01_TMP_PW_NPOCTDN_2025_L1_updated.csv")

doc_l1_2026 <- read_csv("~/GitHub/TEMPEST_Porewater/processing_scripts/DOC/tmp_doc_2026/COMPASS_TEMPEST_DOC_2026.csv") %>%
  mutate(plot = case_when(plot == "FW" ~ "Freshwater",
                          plot == "C" ~ "Control",
                          plot == "SW" ~ "Saltwater",
                          TRUE ~ plot),
            collection_date = lubridate::as_date(as.character(Collection_Date_YYYYMMDD), format = "%Y%m%d"),
           Collection_End_Time_24hrs = str_replace(Collection_End_Time_24hrs, '\\d+', function(m) str_pad(m, 4, pad = '0', side = ("left"))),
         Collection_End_Time_24hrs = strptime(Collection_End_Time_24hrs, format ="%H%M"),
         Collection_End_Time_24hrs = strftime(Collection_End_Time_24hrs, "%H:%M:%S")) %>%
  mutate(collection_datetime = as.POSIXct(paste(collection_date, Collection_End_Time_24hrs), format = "%Y-%m-%d %H:%M:%S"),
        evacuation_date =  lubridate::as_date(as.character(Evacuation_date_YYYMMDD), format = "%Y%m%d")) %>%
  rename(doc_mg_l = npoc_mgL)


```
```{r combine years}
doc_l1_pw_grids <- doc_l1_2022 %>% ## why are there duplicates inserted here? We will never know, something with the 2022 data....
   mutate(troubleshooting = duplicated(doc_l1_2022)) %>%
   filter(troubleshooting == FALSE) %>%
  full_join(doc_l1_2023) %>%
  full_join(doc_l1_2024) %>%
  full_join(doc_l1_2025) %>%
  mutate(evacuation_date = as_date(evacuation_datetime)) %>%
  select(plot, grid, evacuation_date, collection_datetime, doc_mg_l) %>%
  drop_na() 

nrow(doc_l1_pw_grids)

nrow(unique(doc_l1_pw_grids))

doc_l1_pw_grid_2022_2026 <- doc_l1_pw_grids %>%
full_join(doc_l1_2026) %>%
select(plot, grid, evacuation_date, collection_datetime, doc_mg_l) %>%
  mutate( collection_date = lubridate::as_date(collection_datetime),
    evacuation_month = lubridate::floor_date(evacuation_date, "month"),
    collection_month = lubridate::floor_date(collection_date, "month"),
    evacuation_month_year = format(evacuation_month, "%b %Y"),
    collection_month_year = format(collection_month, "%b %Y"),
    days_from_month_end = lubridate::days_in_month(evacuation_date) - lubridate::day(evacuation_date),
    evac_is_month_end = days_from_month_end <= 2, #if within a few days of the end of the month, we want this as the prior month
    Group = case_when(lubridate::month(collection_date) == 6 ~ collection_date,  # June experimental treatment: keep exact collection date
      evac_is_month_end & evacuation_month != collection_month ~ collection_month, # Edge case: evacuation started on the last day of prior month, so assign the interval to the collection month
      TRUE ~ evacuation_month),    # Default: assign integrated sample to evacuation month
    sampling_interval = paste(evacuation_date, collection_date, sep = "_to_"),
    crosses_month = evacuation_month != collection_month,
    Group = case_when(Group == "2022-12-01" ~ lubridate::as_date("2022-11-01"),
                      Group == "2025-02-01" ~ lubridate::as_date("2025-01-01"),
                           TRUE ~ Group)) %>%
  select(-evacuation_month, -collection_month) %>%
  group_by(Group) %>%
  mutate(nominal_date = mean(evacuation_date) + (mean(collection_date) - mean(evacuation_date))/2) %>%
  #Fix some edge cases in 2023 sampling (straddling months)
  #Fix edge cases during the events:
  mutate( nominal_date = case_when(Group == "2023-02-01" ~ Group,
                                  Group == "2023-10-01" ~ Group,
                                  Group == "2023-06-06" ~ Group,
                                  Group == "2024-06-11" ~ Group,
                                  TRUE ~ nominal_date)) %>% 
  ungroup()

nrow(doc_l1_pw_grid_2022_2026)

nrow(unique(doc_l1_pw_grid_2022_2026))

saveRDS(doc_l1_pw_grid_2022_2026, "~/GitHub/TEMPEST-1-porewater/data/doc_porewatergrids_2022-2026.rds")

doc_pw_studydates <-  doc_l1_pw_grid_2022_2026 %>%
  filter(between(collection_date, startstudydate, endstudydate))
  

saveRDS(doc_pw_studydates, "~/GitHub/TEMPEST-1-porewater/data/doc_porewatergrids_STUDYPERIOD_may2022-may2025.rds")

```
```{r means DOC}

 doc_means_pw <- doc_pw_studydates %>% 
  group_by(nominal_date, plot) %>%
  dplyr::select(plot, doc_mg_l, nominal_date) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

  #select(plot, grid, Group, doc_mg_l, sampling_interval, nominal_date) %>%
  #group_by(plot, Group) %>%
  #dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
  # ungroup() %>%

saveRDS(doc_means_pw, "~/GitHub/TEMPEST-1-porewater/data/DOC_Figure1_FINAL.rds")


 doc_means_pw_all <- doc_l1_pw_grid_2022_2026 %>% 
  group_by(nominal_date, plot) %>%
  dplyr::select(plot, doc_mg_l, nominal_date) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

saveRDS(doc_means_pw_all, "~/GitHub/TEMPEST-1-porewater/data/DOC_FigureS8.rds")


```


# Main Message 1: One salinity exposure permanently changes the soil DOC component of the carbon cycle. 

### Stats for DOC paper

"DOC concentrations did not differ significantly between the freshwater and control plots (GLS: t = 1.90, df = 5.13, p = 0.114; see extended data)."

```{r doc glm means}

pw_doc_cf <-  doc_means_pw  %>%
  mutate(logdoc = log(doc_mg_l_mean)) %>%
  group_by(plot) %>%
  arrange(nominal_date, .by_group = TRUE) %>%
  mutate(index = row_number()) %>%
  ungroup() %>%
  filter(plot != "Saltwater") 

doc_gls <- nlme::gls(
  logdoc ~ plot,
  correlation = corAR1(form = ~ index | plot),
    weights = varIdent(form = ~1 | plot),
  data = pw_doc_cf,
  method = "REML"
)

summary(doc_gls)

plot(doc_gls)

qqnorm(resid(doc_gls))
qqline(resid(doc_gls))

#one factor model

EMM <- emmeans(doc_gls, ~ plot)

EMM

### pairwise comparisons
pairs(EMM)


```
### calculate effect sizes for figure 1 

DOC effect-size workflow:
1. Split data frames; Drop DOC outliers in the control plot. 
2. Collapse repeated samples to Group/plot/grid means.
3. Get datasets ready (unique columns for control and treatment plots). Use same-Group Control when available. If no same-Group Control exists, gapfill from previous mean value and global standard deviation for the plot. if no standard deviation for the treatment plots, use the plot average across all time points. Drop timepoints when data only in control and not in the treatment plots. 
4. Bootstrap spatial variability and write a plotting-ready CSV.

```{r split data frame}

fw_plot <- doc_pw_studydates %>%
  select(plot, grid, collection_datetime, nominal_date, doc_mg_l) %>%
  filter(plot == "Freshwater") %>%
  group_by(nominal_date, plot) %>%
  dplyr::summarise(
    n_timepoints_collapsed = n_distinct(collection_datetime), # for tracking only; note that these count each lysimeter separately in future years when we got better about timepoints 
    treatment_doc_mg_l_mean = mean(doc_mg_l, na.rm = TRUE),
    treatment_doc_mg_l_sd = sd(doc_mg_l, na.rm = TRUE),
    treatment_n_grids = n(),
    treatment_doc_mg_l_se = treatment_doc_mg_l_sd / sqrt(treatment_n_grids),
    .groups = "drop"
  )
  

sw_plot <- doc_pw_studydates %>%
  select(plot, grid, collection_datetime, nominal_date, doc_mg_l) %>%
  filter(plot == "Saltwater") %>%
  group_by(nominal_date, plot) %>%
  dplyr::summarise(
    n_timepoints_collapsed = n_distinct(collection_datetime), # for tracking only; note that these count each lysimeter separately in future years when we got better about timepoints 
    treatment_doc_mg_l_mean = mean(doc_mg_l, na.rm = TRUE),
    treatment_doc_mg_l_sd = sd(doc_mg_l, na.rm = TRUE),
    treatment_n_grids = n(),
    treatment_doc_mg_l_se = treatment_doc_mg_l_sd / sqrt(treatment_n_grids),
    .groups = "drop"
  )
  

c_plot_for_outliers <- doc_pw_studydates %>%
  select(plot, grid, collection_datetime, nominal_date, doc_mg_l) %>%
  filter(plot == "Control") %>%
  mutate(source_row = row_number())

control_median_doc <- median(c_plot_for_outliers$doc_mg_l, na.rm = TRUE)
control_mad_doc <- mad(c_plot_for_outliers$doc_mg_l, constant = 1, na.rm = TRUE)

control_removed_outliers <- c_plot_for_outliers %>%
  mutate(
    control_modified_z = if (control_mad_doc > 0) {
      0.6745 * (doc_mg_l - control_median_doc) / control_mad_doc
    } else {
      NA_real_
    },
    control_mad_outlier = abs(control_modified_z) > 3.5
  ) %>%
  filter(control_mad_outlier)

c_plot <- c_plot_for_outliers %>%
  anti_join(
    control_removed_outliers %>% select(source_row),
    by = "source_row"
  ) %>%
   group_by(nominal_date, plot) %>%
  dplyr::summarise(
    n_timepoints_collapsed = n_distinct(collection_datetime), # for tracking only; note that these count each lysimeter separately in future years when we got better about timepoints 
    control_doc_mg_l_mean = mean(doc_mg_l, na.rm = TRUE),
    control_doc_mg_l_sd = sd(doc_mg_l, na.rm = TRUE),
    control_n_grids = n(),
    control_doc_mg_l_se = control_doc_mg_l_sd / sqrt(control_n_grids),
    .groups = "drop"
  )


fw_c_foreffect <- fw_plot %>%
  full_join(c_plot, by = c("nominal_date")) %>%
  filter(plot.x == "Freshwater") %>% # filter out when only have data for treatment plot 
  #gapfilling when missing control data points; using "typical monthly values" for both mean and std: 
  mutate(month = month(nominal_date)) %>%
  group_by(month) %>%
    mutate(
     plot.y = case_when(is.na(plot.y) ~ "Gapfilled",
                            TRUE ~ plot.y),
  control_n_grids = if_else(is.na(control_n_grids),
  as.integer(round(median(control_n_grids, na.rm = TRUE))),
  control_n_grids),
  treatment_n_grids = if_else(is.na(treatment_n_grids),
  as.integer(round(median(treatment_n_grids, na.rm = TRUE))),
  treatment_n_grids),
    control_doc_mg_l_mean = if_else(is.na(control_doc_mg_l_mean),
    mean(control_doc_mg_l_mean, na.rm = TRUE),
    control_doc_mg_l_mean),
  control_doc_mg_l_sd = if_else(is.na(control_doc_mg_l_sd),
    mean(control_doc_mg_l_sd, na.rm = TRUE),
    control_doc_mg_l_sd),
  treatment_doc_mg_l_sd = if_else(is.na(treatment_doc_mg_l_sd),
    mean(treatment_doc_mg_l_sd, na.rm = TRUE),
    treatment_doc_mg_l_sd),
  control_doc_mg_l_se = control_doc_mg_l_sd / sqrt(control_n_grids),
  treatment_doc_mg_l_se = treatment_doc_mg_l_sd / sqrt(treatment_n_grids),
    treatment_doc_mg_l_se = if_else(
      is.na(treatment_doc_mg_l_se),
      median(treatment_doc_mg_l_se, na.rm = TRUE),
      treatment_doc_mg_l_se)) %>%
  ungroup()


sw_c_foreffect <- sw_plot %>%
  full_join(c_plot, by = c("nominal_date")) %>%
  filter(plot.x == "Saltwater") %>% # filter out when only have data for treatment plot 
  #gapfilling when missing control data points; using "typical monthly values" for both mean and std: 
  mutate(month = month(nominal_date)) %>%
  group_by(month) %>%
   mutate(
     plot.y = case_when(is.na(plot.y) ~ "Gapfilled",
                            TRUE ~ plot.y),
  control_n_grids = if_else(is.na(control_n_grids),
  as.integer(round(median(control_n_grids, na.rm = TRUE))),
  control_n_grids),
  treatment_n_grids = if_else(is.na(treatment_n_grids),
  as.integer(round(median(treatment_n_grids, na.rm = TRUE))),
  treatment_n_grids),
    control_doc_mg_l_mean = if_else(is.na(control_doc_mg_l_mean),
    mean(control_doc_mg_l_mean, na.rm = TRUE),
    control_doc_mg_l_mean),
  control_doc_mg_l_sd = if_else(is.na(control_doc_mg_l_sd),
    mean(control_doc_mg_l_sd, na.rm = TRUE),
    control_doc_mg_l_sd),
  treatment_doc_mg_l_sd = if_else(is.na(treatment_doc_mg_l_sd),
    mean(treatment_doc_mg_l_sd, na.rm = TRUE),
    treatment_doc_mg_l_sd),
  control_doc_mg_l_se = control_doc_mg_l_sd / sqrt(control_n_grids),
  treatment_doc_mg_l_se = treatment_doc_mg_l_sd / sqrt(treatment_n_grids),
    treatment_doc_mg_l_se = if_else(
      is.na(treatment_doc_mg_l_se),
      median(treatment_doc_mg_l_se, na.rm = TRUE),
      treatment_doc_mg_l_se)) %>%
  ungroup()

 
```

```{r effect size}

set.seed(20260603)

n_boot <- 10000

fw_effect <- fw_c_foreffect %>%
  mutate(
    effect_size = (treatment_doc_mg_l_mean - control_doc_mg_l_mean) /
      control_doc_mg_l_mean
  )

sw_effect <- sw_c_foreffect %>%
  mutate(
    effect_size = (treatment_doc_mg_l_mean - control_doc_mg_l_mean) /
      control_doc_mg_l_mean
  )

## bootstrapping 

save_bootstrap_replicates <- FALSE
estimation_method <- "parametric_normal_plot_mean_uncertainty"

#inputs
effect_size_bootstrap_input <- bind_rows(
  fw_effect %>% mutate(treatment_plot = "Freshwater"),
  sw_effect %>% mutate(treatment_plot = "Saltwater")
) %>%
  mutate(
    bootstrap_row_id = row_number(),
    nominal_date = as.Date(nominal_date),
    control_reference = case_when(
      as.character(plot.y) == "Control" ~ "same_nominal_date_control",
      as.character(plot.y) == "Gapfilled" ~ "gapfilled_monthly_control",
      is.na(plot.y) ~ "missing_control",
      TRUE ~ as.character(plot.y)
    ),
    estimation_method = estimation_method,
    bootstrap_status = case_when(
      !is.finite(control_doc_mg_l_mean) | control_doc_mg_l_mean <= 0 ~
        "skipped_bad_control_mean",
      !is.finite(treatment_doc_mg_l_mean) ~ "skipped_bad_treatment_mean",
      !is.finite(control_doc_mg_l_se) | control_doc_mg_l_se < 0 ~
        "skipped_missing_control_se",
      !is.finite(treatment_doc_mg_l_se) | treatment_doc_mg_l_se < 0 ~
        "skipped_missing_treatment_se",
      TRUE ~ "calculated"
    )
  )

calculated_bootstrap_input <- effect_size_bootstrap_input %>%
  filter(bootstrap_status == "calculated")

bootstrap_replicates <- map_dfr(seq_len(nrow(calculated_bootstrap_input)), function(i) {
  this_row <- calculated_bootstrap_input[i, ]

  control_mean_boot <- rnorm(
    n_boot,
    mean = this_row$control_doc_mg_l_mean,
    sd = this_row$control_doc_mg_l_se
  )

  treatment_mean_boot <- rnorm(
    n_boot,
    mean = this_row$treatment_doc_mg_l_mean,
    sd = this_row$treatment_doc_mg_l_se
  )

  tibble(
    bootstrap_row_id = this_row$bootstrap_row_id,
    nominal_date = this_row$nominal_date,
    treatment_plot = this_row$treatment_plot,
    control_reference = this_row$control_reference,
    bootstrap_iteration = seq_len(n_boot),
    control_doc_mg_l_mean_boot = control_mean_boot,
    treatment_doc_mg_l_mean_boot = treatment_mean_boot,
    effect_size_boot = if_else(
      control_doc_mg_l_mean_boot > 0,
      (treatment_doc_mg_l_mean_boot - control_doc_mg_l_mean_boot) /
        control_doc_mg_l_mean_boot,
      NA_real_
    )
  )
})

bootstrap_summary <- bootstrap_replicates %>%
  filter(is.finite(effect_size_boot)) %>%
  group_by(bootstrap_row_id) %>%
  summarise(
    bootstrap_iterations = n(),
    effect_size_boot_mean = mean(effect_size_boot),
    effect_size_boot_sd = sd(effect_size_boot),
    effect_size_boot_median = median(effect_size_boot),
    effect_size_ci_lower = quantile(effect_size_boot, 0.025),
    effect_size_ci_upper = quantile(effect_size_boot, 0.975),
    effect_size_percent_boot_mean = 100 * effect_size_boot_mean,
    effect_size_percent_ci_lower = 100 * effect_size_ci_lower,
    effect_size_percent_ci_upper = 100 * effect_size_ci_upper,
    prob_effect_size_gt_0 = mean(effect_size_boot > 0),
    prob_effect_size_lt_0 = mean(effect_size_boot < 0),
    .groups = "drop"
  )

```

```{r bootstrap final data frame}

effect_size_bootstrap_summary <- effect_size_bootstrap_input %>%
  left_join(bootstrap_summary, by = "bootstrap_row_id") %>%
  arrange(nominal_date, factor(treatment_plot, levels = c("Freshwater", "Saltwater"))) %>%
  filter(bootstrap_status == "calculated") %>%
  select(
    nominal_date,
    treatment_plot,
    control_reference,
    effect_size_boot_mean,
    effect_size_ci_lower,
    effect_size_ci_upper,
    effect_size_percent_boot_mean,
    effect_size_percent_ci_lower,
    effect_size_percent_ci_upper,
    estimation_method)

saveRDS(effect_size_bootstrap_summary, "~/GitHub/TEMPEST-1-porewater/data/DOC_Figure1_effectsizes.rds")

```




# Main Message 2: The possible mechanisms at play here are all changing in concert, which clouds predictability across systems and scales.


### ANALYSIS: When does saltwater plot DOC go out of "normal" variability?  

"Therefore, we suggest that 235 +/- 29 µS/cm may serve as a broad conductivity threshold between depressed versus elevated soil porewater DOC concentrations (see methods for calculation). "

```{r doc change analysis}

#1. calculate control plot baseline 

baseline <- doc_l1_pw_grids %>%
  filter(plot == "Control") %>%
  group_by(evacuation_date) %>%
  summarize(
    baseline_mean = mean(doc_mg_l, na.rm = TRUE),
    baseline_sd = sd(doc_mg_l, na.rm = TRUE) # Include standard deviation for later analysis
  ) 

max_sd <- max(baseline$baseline_sd, na.rm= TRUE)
                                   

# 2. Combine baseline data with saltwater data for comparison.
comparison <- doc_l1_pw_grids  %>%
  filter(plot == "Saltwater") %>%
  group_by(evacuation_date) %>%
  summarize(
    doc_median = median(doc_mg_l, na.rm = TRUE),
    doc_sd = sd(doc_mg_l, na.rm = TRUE) 
  ) %>%
  ungroup()%>%
  left_join(baseline, by = "evacuation_date") %>%
  mutate(baseline_mean = case_when(is.na(baseline_mean) ~ mean(baseline_mean, na.rm =TRUE),
                                   TRUE ~ baseline_mean),
         baseline_sd = case_when(is.na(baseline_sd) ~ max_sd,
                                 TRUE ~ baseline_sd)) %>%
  mutate(
    # Calculate the deviation of saltwater analyte concentration from baseline.
    deviation = doc_median - baseline_mean,
    # Flag dates where the saltwater concentration is significantly higher than baseline
    flag_high_change = deviation > (3 * baseline_sd) # Adjust multiplier based on your needs
  )

# Inspect the comparison dataset
print(comparison)

# 3. Extract the dates where there is a "significant change" in the saltwater plot.
significant_dates <- comparison %>%
  filter(flag_high_change) %>%
  select(evacuation_date, doc_median , deviation) %>%
  mutate(year = lubridate::year(evacuation_date)) %>%
  ungroup()

print(significant_dates)

# 4. Get the "first" date post event each year. 

tmp1_first_dates <- significant_dates %>% 
  filter(year == "2022") %>%
  filter(evacuation_date == min(evacuation_date))

tmp2_first_dates <- significant_dates %>% 
  filter(year == "2023") %>%
  filter(evacuation_date > WaterDeliveryStop2) %>%
  filter(evacuation_date == min(evacuation_date))

tmp3_first_dates <- significant_dates %>% 
  filter(year > 2023) %>%
  filter(evacuation_date > WaterDeliveryStart3_2024) %>%
  filter(evacuation_date == min(evacuation_date))

tmpdates_doc.1 <- tmp1_first_dates %>%
  full_join(tmp2_first_dates ) %>%
  full_join(tmp3_first_dates)


collection_dates <- doc_l1_pw_grids %>%
  filter(plot == "Saltwater") %>%
  filter(evacuation_date %in% tmpdates_doc.1$evacuation_date) %>%
  select(evacuation_date, collection_datetime) %>%
  group_by(evacuation_date) %>%
  summarise(collection_datetime = min(collection_datetime)) %>%
  ungroup()

tmpdates_doc <-  tmpdates_doc.1 %>%
  full_join(collection_dates, by= "evacuation_date") %>%
  mutate(date_type = c("Year1_date", "Year2_date", "Year3_date"))

tmpdates_doc 

 saveRDS(tmpdates_doc, "~/GitHub/TEMPEST-1-porewater/data/dates_DOC_higher_than_variance.rds")
```

```{r time elapsed}

yr1_elapsed_time <- tmpdates_doc %>% 
  mutate(event_stop_date = c(as_date(WaterDeliveryStop2022), as_date(WaterDeliveryStop2), as_date(WaterDeliveryStop3_2024))) %>%
  mutate(elapsed_time_days = evacuation_date- event_stop_date)

yr1_elapsed_time
```







These "significant dates" fall in the 
1st month post event 2022 (i.e. July)
2nd month post event 2023 (i.e. August)
8th month post event 2024 (i.e. January)

### Dissolved organic carbon has a strong and delayed response to saltwater flooding
"The first novel exposure of this temperate forest to saltwater from the nearby estuary across the three years caused lagged and markedly higher DOC concentrations in the saltwater plot (43.1 +/- 20.7 mg C/L) compared to the freshwater (11.5 +/- 5.9 mg C/L) and the control plots (17.9 +/- 8.2 mg C/L), which were relatively stable through time. DOC concentrations in the saltwater plot increased by an average of 140% in comparison to the control plot. "

```{r doc summaries paper}

fw_doc_mean_all <- doc_l1_pw_grids %>%
  filter(plot == "Freshwater") %>%
  summarize(
    fw_mean = mean(doc_mg_l, na.rm = TRUE),
    fw_sd = sd(doc_mg_l, na.rm = TRUE) 
  ) 

fw_doc_mean_all

control_doc_mean_all <- doc_l1_pw_grids %>%
  filter(plot == "Control") %>%
  summarize(
    control_mean = mean(doc_mg_l, na.rm = TRUE),
    control_sd = sd(doc_mg_l, na.rm = TRUE) 
  ) 

control_doc_mean_all

salt_doc_mean_beforeincrease <- doc_l1_pw_grids %>%
   filter(plot == "Saltwater") %>%
  filter(
        (evacuation_date > EventStart2022 & evacuation_date < "2022-07-18") | # year 1
        (evacuation_date > EventStart2023 & evacuation_date < "2023-08-07") |  # year 2 
        (evacuation_date > EventStart2024 & evacuation_date < "2025-01-28") # year 3
        ) %>%
    summarize(
    sw_mean = mean(doc_mg_l, na.rm = TRUE),
    sw_sd = sd(doc_mg_l, na.rm = TRUE) 
  ) 

salt_doc_mean_beforeincrease

salt_doc_mean_afterincrease <- doc_l1_pw_grids %>%
   filter(plot == "Saltwater") %>%
  filter(
        (evacuation_date > "2022-07-18" & evacuation_date < EventStart2023) | # year 1
        (evacuation_date > "2023-08-07" & evacuation_date < EventStart2024 ) |  # year 2 
        (evacuation_date > "2025-01-28" & evacuation_date < endstudydate_yr3) # year 3
        ) %>%
    summarize(
    sw_mean = mean(doc_mg_l, na.rm = TRUE),
    sw_sd = sd(doc_mg_l, na.rm = TRUE) 
  ) 

salt_doc_mean_afterincrease

percent_increase_mean <- ((salt_doc_mean_afterincrease$sw_mean - control_doc_mean_all$control_mean)/ control_doc_mean_all$control_mean) * 100

percent_increase_mean

sd_percent_increase <- (100 / control_doc_mean_all$control_mean) * sqrt(salt_doc_mean_afterincrease$sw_sd^2 + (control_doc_mean_all$control_sd^2 * ((salt_doc_mean_afterincrease$sw_mean - control_doc_mean_all$control_mean) / control_doc_mean_all$control_mean)^2))

sd_percent_increase

```

### Conductivity: What is the conductivity values when DOC is high? 

What is the conductivity when this happens?
- bring in teros conductivity plot averages to ID conductivity values when DOC is high

```{r teros EC}

pw_grids_ec <- read_rds("~/GitHub/TEMPEST-1-porewater/processed data/TMP_PW_EC_15cm_2021-2025_L2.rds")

pw_plot_ec <- pw_grids_ec %>%
  group_by(plot, depth, research_name, datetime_est, yday(datetime_est)) %>%
        summarise(datetime_est = mean(datetime_est),
                  EC_uScm = mean(value, na.rm = TRUE),
                  .groups = "drop")

# pw_plot_sw <- pw_plot_ec %>%
  

# write out for plotting:
saveRDS(pw_plot_ec, "~/GitHub/TEMPEST-1-porewater/data/averaged_by_plot_ec_teros_porewater.rds")

```

Mean plot conductivity values when DOC has reached 3x SD of control plot. 

```{r cond threshold}
dates_4_cond <- tmpdates_doc %>%
  mutate(date = lubridate::date(collection_datetime))

cond_doc_thres <- pw_plot_ec %>%
  filter(plot == "Saltwater") %>%
  mutate(date = lubridate::date(datetime_est)) %>%
  group_by(plot, depth, research_name, date) %>%
  filter(date %in% dates_4_cond$date) %>%
  ungroup() %>%
  summarise(EC_uScm_mean = mean(EC_uScm, na.rm = TRUE),
                  EC_uScm_sd = sd(EC_uScm, na.rm = TRUE),
                  .groups = "drop")

cond_doc_thres

```

### Threshold sensitivity analysis 

```{r, sensitivity}

library(tidyverse)
library(lubridate)

# Sensitivity values to test
sd_multipliers <- seq(1.5, 5, by = 0.5)

# 1. Control baseline by date
baseline <- doc_l1_pw_grids %>%
  filter(plot == "Control") %>%
  group_by(evacuation_date) %>%
  summarize(
    baseline_mean = mean(doc_mg_l, na.rm = TRUE),
    baseline_sd = sd(doc_mg_l, na.rm = TRUE),
    .groups = "drop"
  )

max_sd <- max(baseline$baseline_sd, na.rm = TRUE)

# 2. Saltwater DOC summary by date
salt_doc <- doc_l1_pw_grids %>%
  filter(plot == "Saltwater") %>%
  group_by(evacuation_date) %>%
  summarize(
    doc_median = median(doc_mg_l, na.rm = TRUE),
    doc_mean = mean(doc_mg_l, na.rm = TRUE),
    doc_sd = sd(doc_mg_l, na.rm = TRUE),
    collection_datetime = min(collection_datetime, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(baseline, by = "evacuation_date") %>%
  mutate(
    baseline_mean = if_else(
      is.na(baseline_mean),
      mean(baseline_mean, na.rm = TRUE),
      baseline_mean
    ),
    baseline_sd = if_else(
      is.na(baseline_sd),
      max_sd,
      baseline_sd
    ),
    deviation = doc_median - baseline_mean
  )

# 3. Function to identify first exceedance dates by multiplier
get_first_exceedance <- function(multiplier) {
  
  significant_dates <- salt_doc %>%
    mutate(
      sd_multiplier = multiplier,
      flag_high_change = deviation > multiplier * baseline_sd,
      year = year(evacuation_date)
    ) %>%
    filter(flag_high_change)
  
  tmp1 <- significant_dates %>%
    filter(year == 2022) %>%
    filter(evacuation_date == min(evacuation_date, na.rm = TRUE))
  
  tmp2 <- significant_dates %>%
    filter(year == 2023) %>%
    filter(evacuation_date > WaterDeliveryStop2) %>%
    filter(evacuation_date == min(evacuation_date, na.rm = TRUE))
  
  tmp3 <- significant_dates %>%
    filter(year > 2023) %>%
    filter(evacuation_date > WaterDeliveryStart3_2024) %>%
    filter(evacuation_date == min(evacuation_date, na.rm = TRUE))
  
  bind_rows(tmp1, tmp2, tmp3) %>%
    mutate(
      sd_multiplier = multiplier,
      event_stop_date = case_when(
        year == 2022 ~ as_date(WaterDeliveryStop2022),
        year == 2023 ~ as_date(WaterDeliveryStop2),
        year > 2023 ~ as_date(WaterDeliveryStop3_2024)
      ),
      elapsed_time_days = as.numeric(evacuation_date - event_stop_date)
    )
}

# 4. Run sensitivity analysis
sensitivity_dates <- map_dfr(sd_multipliers, get_first_exceedance)

sensitivity_dates
```

# try this without the outlier, do outlier test on crazy control plot point
```{r, cond sens}
# Prepare daily conductivity data
pw_plot_ec_daily <- pw_plot_ec %>%
  filter(plot == "Saltwater") %>%
  mutate(date = date(datetime_est)) %>%
  group_by(plot, depth, research_name, date) %>%
  summarize(
    EC_uScm_mean_daily = mean(EC_uScm, na.rm = TRUE),
    EC_uScm_sd_daily = sd(EC_uScm, na.rm = TRUE),
    .groups = "drop"
  )

# Join DOC exceedance dates with conductivity
sensitivity_with_ec <- sensitivity_dates %>%
  mutate(date = date(collection_datetime)) %>%
  left_join(
    pw_plot_ec_daily,
    by = "date"
  )

sensitivity_with_ec

sensitivity_summary <- sensitivity_with_ec %>%
  group_by(sd_multiplier) %>%
  summarize(
    n_years_detected = n_distinct(year),
    mean_elapsed_days = mean(elapsed_time_days, na.rm = TRUE),
    sd_elapsed_days = sd(elapsed_time_days, na.rm = TRUE),
    mean_EC_uScm = mean(EC_uScm_mean_daily, na.rm = TRUE),
    sd_EC_uScm = sd(EC_uScm_mean_daily, na.rm = TRUE),
    min_EC_uScm = min(EC_uScm_mean_daily, na.rm = TRUE),
    max_EC_uScm = max(EC_uScm_mean_daily, na.rm = TRUE),
    .groups = "drop"
  )

sensitivity_summary

ggplot(sensitivity_summary, aes(x = sd_multiplier, y = mean_EC_uScm)) +
  geom_point(size = 3) +
  geom_line() +
  geom_errorbar(
    aes(
      ymin = mean_EC_uScm - sd_EC_uScm,
      ymax = mean_EC_uScm + sd_EC_uScm
    ),
    width = 0.1
  ) +
  labs(
    x = "DOC exceedance threshold multiplier",
    y = "Mean conductivity at first DOC exceedance (uS/cm)",
    title = "Sensitivity of conductivity threshold to DOC exceedance criterion"
  ) +
  theme_classic()

ggplot(
  sensitivity_with_ec,
  aes(sd_multiplier, EC_uScm_mean_daily, color = factor(year))
) +
  geom_point(size = 3) +
  geom_line() +
  theme_classic()
```

```{r set dates based on pw_doc_dates}
tmp1_doc_high = as.POSIXct("2022-07-21 15:59:00", tz = "EST")
tmp2_doc_high = as.POSIXct("2023-08-10 15:53:00", tz = "EST")
tmp3_doc_high = as.POSIXct("2025-01-31 20:52:00", tz = "EST")

```

```{r bring in precip}

precip <- read_csv("~/GitHub/TEMPEST-1-porewater/data/daily_precip_2021-2026.csv")

precip_TMP1_sum <- precip %>%
  filter(between(date, EventStart2022, tmp1_doc_high)) %>%
  summarize(precip_cm_sum = sum(precip_cm))

precip_TMP1_sum 

precip_TMP2_sum <- precip %>%
  filter(between(date, EventStart2023, tmp2_doc_high)) %>%
  summarize(precip_cm_sum = sum(precip_cm))

precip_TMP2_sum 

precip_TMP3_sum <- precip %>%
  filter(between(date, EventStart2024, tmp3_doc_high)) %>%
  summarize(precip_cm_sum = sum(precip_cm))

precip_TMP3_sum 

```

# ColEx fluxes 

```{r colex fluxes}

colex_data <- readxl::read_excel("~/GitHub/tempest_ionic_strength/Data/Exp 3 Column Experiment ColEx Data/ColEx_data_allreps_alltreatments_formatted.xlsx")

```

```{r col ex doc flux}

colex_flux <- colex_data %>%
  group_by(Treatment, Column) %>%
  mutate(across(doc_mgperL:S_ppm, ~ .x * (volume_added_mL/1000), .names = "mgtot_{.col}")) %>%
  rename_with(~ str_remove(.x, "_[^_]+$"), starts_with("mgtot_")) %>%
  filter(Timepoint != 1) %>%
  filter(Column != "Treatment Water") %>%
  summarise(across(starts_with("mgtot_"), sum, na.rm = TRUE), .groups = "drop")


colex_flux_tot <- colex_flux%>%
  ungroup() %>%
  group_by(Treatment) %>%
  summarise(across(starts_with("mgtot_"), list(mean = mean, sd = sd), na.rm = TRUE, .names = "{.col}_{.fn}"), .groups = "drop")

```

```{r column soil}

soil_masses <- readxl::read_excel("~/GitHub/TEMPEST-1-porewater/data_do_not_commit/from ed/Column_data_Soil_masses.xlsx",skip=5) %>%
  select(Column, soil_g) %>% drop_na()

colex_soil <- colex_flux %>%
  full_join(soil_masses, by="Column") %>%
  drop_na()

soil_c <- soil_data %>%
  select(plot, C_mean, C_sd, N_mean, N_sd, Na_mean, Na_sd, K_mean, K_sd, Mg_mean, Mg_sd, Ca_mean, Ca_sd, Fe_mean, Fe_sd, Mn_mean, Mn_sd) %>%
  mutate(Treatment = case_when(plot == "Saltwater" ~ "ASW",
                               TRUE ~ plot)) 

#save for BHM

colex_soil_2 <- colex_soil %>%
  select(Treatment, Column, soil_g)

write_csv(colex_soil_2, "~/GitHub/tempest_ionic_strength/BHM/Data/TEMPEST_ColEx_soil_masses.csv")

```
need a monte carlo to propagate uncertainty since we dont have exact soil carbon values for these cores. 
```{r c flux vs c soil}
n_sim <- 10000 #number of simulations 

soils_colex <- colex_soil %>%
  left_join(soil_c, by = "Treatment")

#function to repeat simulation results, drawing on random C value per treatment from a normal distribution of the data.
#Then, calculate C stock - C export 
sim_results <- replicate(n_sim, {
  
  # one C draw per treatment
  C_draws <- soils_colex %>%
    distinct(Treatment, C_mean_mg, C_sd_mg) %>% #C is in % from the UGA values
    mutate(C_draw = rnorm(n(), C_mean_mg, C_sd_mg))
  
  df_sim <- soils_colex %>%
    select(Column, Treatment, soil_g, mgtot_doc) %>%
    left_join(select(C_draws, Treatment, C_draw), by = "Treatment") %>%
    mutate(
      soil_stock_g = soil_g * C_draw, # g soil x g C / 100g soil 
      net_C_g = soil_stock_g - (mgtot_doc/1000)
    )
  
# summarize 

 df_sim %>%
    group_by(Treatment) %>%
    summarise(mean_net_C = mean(net_C), .groups = "drop") %>%
    pull(mean_net_C)

})


```

### Column experiments stats 

"DOC increased markedly in the ASW soil columns after 6 successive washes of low ionic strength water following the initial ASW pulse, with no significant differences in DOC across washes in the control or freshwater column experiments." 

See markdown: soil_column_doc_wash_statistical_test.Rmd (created with Codex)

# Supplemental Materials 

### The changes in concentration observed are coupled with a dramatic shift in the organic matter chemistry (both during and after the event). 

```{r load cdom data}

cdom_yr1 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_cdom_indicies_porewater_TMP1.rds")
cdom_yr2 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_cdom_indicies_porewater_TMP2.rds")
cdom_yr3 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_cdom_indicies_porewater_TMP3.rds")

cdom_allyrs <- readRDS("~/GitHub/TEMPEST-1-porewater/data/cdom_indicies_porewater_2022-2025.rds")
```

```{r load abs data}

abs_allyears <- readRDS("~/GitHub/TEMPEST-1-porewater/data/absorbance_docnormalized_porewater_2022-2025.rds")
```

2022 cdom was collected at the pooled level but DOC was at the grid level. Take the plot level doc averages for the collection dates where this is the case for SUVA.

```{r abs data}

averaged_DOC_2022 <- doc_l1_2022 %>%
  select(plot, grid, collection_datetime, doc_mg_l) %>%
  drop_na() %>%
  mutate(collection_date = lubridate::as_date(collection_datetime)) %>%
  select(-collection_datetime, -grid) %>%
  group_by(collection_date, plot) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup() %>%
  select(-doc_mg_l_sd)

abs_allyears_1 <- abs_allyears %>%
  left_join(averaged_DOC_2022, by=c("collection_date", "plot")) %>%
  mutate(doc_mg_l_filled = coalesce(doc_mg_l, doc_mg_l_mean),
         SUVA254 = case_when(is.na(SUVA254) ~ Abs_254nm / doc_mg_l_filled * 100,
                             TRUE ~ SUVA254),
         SUVA280 = case_when(is.na(SUVA280) ~ Abs_280nm / doc_mg_l_filled * 100,
                             TRUE ~ SUVA280), 
         SUVA350 = case_when(is.na(SUVA350) ~ Abs_350nm / doc_mg_l_filled * 100,
                             TRUE ~ SUVA350),
         SVA412 = case_when(is.na(SVA412) ~ Abs_412nm / doc_mg_l_filled * 100,
                             TRUE ~ SVA412),
         SVA440 = case_when(is.na(SVA440) ~ Abs_440nm / doc_mg_l_filled * 100,
                             TRUE ~ SVA440)) %>%
  select(sample_name, plot, grid, collection_date, S275_295:Sr, SUVA254:SVA440) %>%
  drop_na()
  
saveRDS(abs_allyears_1, "~/GitHub/TEMPEST-1-porewater/data/absorbance_porewater_2022-2025_withpooled.rds")

```


```{r abs eems}

abs_eems_all_years <- cdom_allyrs %>%
  mutate(collection_date = as_date(collection_datetime)) %>%
  select(-S275_295, - S350_400, -Sr) %>%
  left_join(abs_allyears_1, by = c("sample_name", "collection_date", "plot")) %>%
  distinct(sample_name, .keep_all = TRUE)
```

Separate data by system level analysis bins: 
```{r load functions for analysis}

source("~/GitHub/tempest-system-level-analysis/data_package_prep/scripts/tmp_test_functions.R")

#set doc increase dates 

tmpdates_doc1 <- tmpdates_doc %>%
  filter(date_type == "Year1_date") %>%
  select(evacuation_date, collection_datetime) %>%
  rename(datetime = evacuation_date,
         datetime2 = collection_datetime) %>%
  mutate(datetime = as.POSIXct(datetime, tz= "EST"))

tmpdates_doc2 <- tmpdates_doc %>%
  filter(date_type == "Year2_date") %>%
  select(evacuation_date, collection_datetime) %>%
  rename(datetime = evacuation_date,
         datetime2 = collection_datetime) %>%
  mutate(datetime = as.POSIXct(datetime, tz= "EST"))

tmpdates_doc3 <- tmpdates_doc %>%
  filter(date_type == "Year3_date") %>%
  select(evacuation_date, collection_datetime) %>%
  rename(datetime = evacuation_date,
         datetime2 = collection_datetime) %>%
  mutate(datetime = as.POSIXct(datetime, tz= "EST"))

```

NOTE:::: there are some issues with time zones messing up the dates slightly here!!!!
```{r divide data for cdom}

cdom_tmp1 <- abs_eems_all_years  %>%
  rename(datetime = collection_datetime) %>%
  filter(between(datetime, startstudydate_yr1, endstudydate_yr1))

cdom_tmp1_dd <- divide_data(cdom_tmp1, datetime, flood_event_datetime_1, eventB = c(tmpdates_doc1$datetime, tmpdates_doc1$datetime2), eventB_name = "Post") %>%
    mutate(timedate_bin = case_when(timedate_bin == "Btwn Events" ~ "Pre Cond Threshold",
                                  TRUE ~ timedate_bin)) %>%
    mutate(timedate_bin = case_when(timedate_bin == "Post" ~ "Post Cond Threshold",
                                  TRUE ~ timedate_bin))

cdom_tmp2 <- abs_eems_all_years  %>%
  rename(datetime = collection_datetime) %>%
  filter(between(datetime, startstudydate_yr2, endstudydate_yr2))

cdom_tmp2_dd <- divide_data(cdom_tmp2, datetime, flood_event_datetime_2, eventB = c(tmpdates_doc2$datetime, tmpdates_doc2$datetime2), eventB_name = "Post") %>%
     mutate(timedate_bin = case_when(timedate_bin == "Btwn Events" ~ "Pre Cond Threshold",
                                  TRUE ~ timedate_bin)) %>%
  mutate(timedate_bin = case_when(timedate_bin == "Post" ~ "Post Cond Threshold",
                                  TRUE ~ timedate_bin))%>%
  mutate(timedate_bin = case_when(timedate_bin == "Pre" ~ "Post Cond Threshold",
                                  TRUE ~ timedate_bin))

cdom_tmp3 <- abs_eems_all_years %>%
  rename(datetime = collection_datetime) %>%
  filter(between(datetime, startstudydate_yr3, endstudydate_yr3))  

cdom_tmp3_dd <- divide_data(cdom_tmp3, datetime, flood_event_datetime_3, eventB = c(tmpdates_doc3$datetime, tmpdates_doc3$datetime2), eventB_name = "Post") %>%
     mutate(timedate_bin = case_when(timedate_bin == "Btwn Events" ~ "Pre Cond Threshold",
                                  TRUE ~ timedate_bin)) %>%
  mutate(timedate_bin = case_when(timedate_bin == "Post" ~ "Post Cond Threshold",
                                  TRUE ~ timedate_bin))%>%
  mutate(timedate_bin = case_when(timedate_bin == "Pre" ~ "Post Cond Threshold",
                                  TRUE ~ timedate_bin))


cdom_indicies_divided_all <- cdom_tmp1_dd %>%
  full_join(cdom_tmp2_dd) %>%
  full_join(cdom_tmp3_dd) %>%
  drop_na()

colnames(cdom_indicies_divided_all)[match(
  c("FRESH", "A_T", "C_A", "C_M", "C_T", "S275_295", "S350_400", "Sr"),
  colnames(cdom_indicies_divided_all)
)] <- c(
  "β:α",
  "A:T",
  "C:A",
  "C:M",
  "C:T",
  "S275–295",
  "S350–400",
  "SR"
)

saveRDS(cdom_indicies_divided_all, "~/GitHub/TEMPEST-1-porewater/data/cdom_all_forplotting.rds" )

```


```{r try multivariate eems analysis}

eems4multi <- cdom_indicies_divided_all %>%
  select(-sample_name, -grid, -plot, -timedate_bin, -datetime, -collection_date) 

eemspca = prcomp(eems4multi , retx = TRUE, center = TRUE, scale. = TRUE)

eemspca

saveRDS(eemspca, "~/GitHub/TEMPEST-1-porewater/data/cdom_pca.rds")

```

```{r gllm prep}
cdom_stats <- cdom_indicies_divided_all %>%
  mutate(date = as_date(datetime),
         timedate_bin = factor(timedate_bin),
         date = factor(date),
         grid = factor(grid),
         plot = factor(plot),
       spatial_rep = case_when(grid =="POOL" ~ NA,
                               TRUE ~ grid),
        spatial_rep = factor(spatial_rep),
       plot_time = interaction(plot, date, drop = TRUE)) %>%
  filter(
    !is.na(SUVA254),
    !is.na(timedate_bin),
    !is.na(plot),
    !is.na(grid),
    SUVA254 > 0
  )

# Check replication before modeling
table(cdom_stats$plot, cdom_stats$timedate_bin)
table(cdom_stats$plot_time)

```

```{r glmm model}

  # Packages
library(glmmTMB)
library(emmeans)
library(DHARMa)

# Example assumes your data frame is called dat
# Columns:
# y              response variable
# timedate_bin           4-category bin factor
# datetime          sampling timepoint
# grid       repeated sampler ID
# plot           plot ID
# spatial_rep    spatial replicate ID, NA for pooled samples


# GLMM
m1 <- glmmTMB(
  SUVA254 ~ timedate_bin +
    (1 | plot) +
    (1 | grid),
  family = Gamma(link = "log"),
  data = cdom_stats
)

# Model diagnostics
sim <- simulateResiduals(m1)
plot(sim)
testDispersion(sim)
testZeroInflation(sim)

# Test overall bin effect
drop1(m1, test = "Chisq")

# look at responses
emmeans(m1, ~ timedate_bin, type = "response")
pairs(emmeans(m1, ~ timedate_bin), adjust = "tukey", type = "response")

```


loc <- cdom_indicies_divided_all %>% select(plot, timedate_bin)

env1 <- vegan::envfit(nmds, loc, permu = 999)

vegan::ordiplot(nmds, display = 'si', type = 'n')
points(nmds, pch = 19)
vegan::ordiellipse(nmds, groups= cdom_indicies_divided_all$timedate_bin )



### pH: What is the relationship between pH and DOC after each event? 

We can do this analysis a bit differently than conductivity since we should have directly paired datasets.

1. Plot level pH averages and stds
```{r load pH data}

ph_2023 <-  read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_pH_2023_L1.csv") 
ph_2024 <-  read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_pH_2024_L1.csv") 
ph_2025 <-  read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_pH_2025_L1.csv") 

```

```{r ph all years}
ph_l1_pw_grids <- ph_2023 %>%
  full_join(ph_2024) %>%
  full_join(ph_2025) %>%
  dplyr::select(plot, datetime, evacuation_date, pH) %>%
  drop_na() %>%
  mutate(date = lubridate::date(datetime)) %>%
  group_by(plot, evacuation_date)
  
```

Mean plot pH values when DOC has reached 10x SD of control plot. 

```{r ph threshold}
dates_4_ph <- tmpdates_doc %>%
  mutate(date = lubridate::date(collection_datetime))

ph_doc_thres <- ph_l1_pw_grids %>%
  filter(plot == "Saltwater") %>%
  mutate(date = lubridate::date(datetime)) %>%
  group_by(plot, date) %>%
  filter(date %in% dates_4_ph$date) %>%
  ungroup() %>%
  summarise(pH_mean = mean(pH, na.rm = TRUE),
                  pH_sd = sd(pH, na.rm = TRUE),
                  .groups = "drop")

ph_doc_thres

```

2 a. ph time series plots

```{r ph avg std for timeseries plots}
# Some dates not enough water was collected to have multiple samples to get averages and sds for that evacuation period. Most of the time, those can be combined with sampling dates just before or just after as we go out over several multi-day periods. This, however isn't consistent across the FW, EW, or C plots, so it will be a bit tricky to do the gap filling. To make the gap filling more accurate, we'll take the average of those dates, so for example if the dates we're merging is 2022-11-25 and 2022-11-28 we'll have the date be the 26th (but we'll let R do that math).

#First, isolate the trouble dates/times: 

#Control: 
mean_ph_l1_c_to_gapfill <- ph_l1_pw_grids %>%
  filter(plot == "Control") %>%
  filter(evacuation_date %in% c("2023-06-12", "2023-06-14", "2023-07-05", "2023-07-07", "2023-08-04", "2023-08-07","2023-12-05", "2023-12-08")) %>%
  mutate(gapfill_group = case_when( 
    str_detect(evacuation_date,"2023-06-12") | str_detect(evacuation_date,"2023-06-14") ~ "A",
    str_detect(evacuation_date,"2023-07-05") | str_detect(evacuation_date,"2023-07-07") ~ "B",
    str_detect(evacuation_date,"2023-08-04") | str_detect(evacuation_date,"2023-08-07") ~ "C",
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "D",
    TRUE ~ "no")) %>%
   group_by(plot, gapfill_group) %>%
    dplyr::summarise(across(everything(), mean), .groups = "keep") %>%
  mutate(datetime = as.POSIXct(datetime, origin = "1970-01-01"),
         evacuation_date = as_date(evacuation_date)) %>%
  ungroup()%>%
  dplyr::select(-date)

sd_ph_l1_c_to_gapfill <- ph_l1_pw_grids %>%
  select(-datetime) %>%
  filter(plot == "Control") %>%
 filter(evacuation_date %in% c("2023-06-12", "2023-06-14", "2023-07-05", "2023-07-07", "2023-08-04", "2023-08-07","2023-12-05", "2023-12-08")) %>%
  mutate(gapfill_group = case_when( 
    str_detect(evacuation_date,"2023-06-12") | str_detect(evacuation_date,"2023-06-14") ~ "A",
    str_detect(evacuation_date,"2023-07-05") | str_detect(evacuation_date,"2023-07-07") ~ "B",
    str_detect(evacuation_date,"2023-08-04") | str_detect(evacuation_date,"2023-08-07") ~ "C",
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "D",
    TRUE ~ "no")) %>%
   group_by(plot, gapfill_group) %>%
    select(-evacuation_date) %>%
    dplyr::summarise(across(everything(), sd), .groups = "keep") %>%
  rename(pH_sd = pH) %>%
    ungroup() %>%
  dplyr::select(-date)

ph_c_gapfill <- mean_ph_l1_c_to_gapfill %>%
  full_join(sd_ph_l1_c_to_gapfill, by = c("plot", "gapfill_group")) %>%
  select(-gapfill_group)

#Freshwater: 
mean_ph_l1_fw_to_gapfill <- ph_l1_pw_grids %>%
  filter(plot == "Freshwater") %>%
  filter(evacuation_date %in% c("2023-06-12", "2023-06-14", "2023-08-04", "2023-08-07","2023-12-05", "2023-12-08")) %>%
  mutate(gapfill_group = case_when( 
    str_detect(evacuation_date,"2023-06-12") | str_detect(evacuation_date,"2023-06-14") ~ "A",
    str_detect(evacuation_date,"2023-08-04") | str_detect(evacuation_date,"2023-08-07") ~ "B",
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "C",
    TRUE ~ "no")) %>%
   group_by(plot, gapfill_group) %>%
    dplyr::summarise(across(everything(), mean), .groups = "keep") %>%
  mutate(datetime = as.POSIXct(datetime, origin = "1970-01-01"),
         evacuation_date = as_date(evacuation_date)) %>%
  ungroup()%>%
  dplyr::select(-date)

sd_ph_l1_fw_to_gapfill <- ph_l1_pw_grids %>%
  select(-datetime) %>%
  filter(plot == "Freshwater") %>%
  filter(evacuation_date %in% c("2023-06-12", "2023-06-14", "2023-08-04", "2023-08-07","2023-12-05", "2023-12-08")) %>%
  mutate(gapfill_group = case_when( 
    str_detect(evacuation_date,"2023-06-12") | str_detect(evacuation_date,"2023-06-14") ~ "A",
    str_detect(evacuation_date,"2023-08-04") | str_detect(evacuation_date,"2023-08-07") ~ "B",
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "C",
    TRUE ~ "no")) %>%
   group_by(plot, gapfill_group) %>%
    select(-evacuation_date) %>%
    dplyr::summarise(across(everything(), sd), .groups = "keep") %>%
  rename(pH_sd = pH) %>%
    ungroup() %>%
  dplyr::select(-date)

ph_fw_gapfill <- mean_ph_l1_fw_to_gapfill %>%
  full_join(sd_ph_l1_fw_to_gapfill, by = c("plot", "gapfill_group")) %>%
  select(-gapfill_group)

#Seawater: 
mean_ph_l1_sw_to_gapfill <- ph_l1_pw_grids %>%
  filter(plot == "Saltwater") %>%
  filter(evacuation_date %in% c("2023-06-12", "2023-06-14", "2023-07-05", "2023-07-07", "2023-08-04", "2023-08-07","2023-09-29", "2023-10-02", "2023-12-05", "2023-12-08")) %>%
  mutate(gapfill_group = case_when( 
    str_detect(evacuation_date,"2023-06-12") | str_detect(evacuation_date,"2023-06-14") ~ "A",
    str_detect(evacuation_date,"2023-07-05") | str_detect(evacuation_date,"2023-07-07") ~ "B",
    str_detect(evacuation_date,"2023-08-04") | str_detect(evacuation_date,"2023-08-07") ~ "C",
    str_detect(evacuation_date,"2023-09-29") | str_detect(evacuation_date,"2023-10-02") ~ "D",
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "E",
    TRUE ~ "no")) %>%
   group_by(plot, gapfill_group) %>%
    dplyr::summarise(across(everything(), mean), .groups = "keep") %>%
  mutate(datetime = as.POSIXct(datetime, origin = "1970-01-01"),
         evacuation_date = as_date(evacuation_date)) %>%
  ungroup()%>%
  dplyr::select(-date)

sd_ph_l1_sw_to_gapfill <- ph_l1_pw_grids %>%
  select(-datetime) %>%
  filter(plot == "Saltwater") %>%
  filter(evacuation_date %in% c("2023-06-12", "2023-06-14", "2023-07-05", "2023-07-07", "2023-08-04", "2023-08-07","2023-09-29", "2023-10-02", "2023-12-05", "2023-12-08")) %>%
  mutate(gapfill_group = case_when( 
    str_detect(evacuation_date,"2023-06-12") | str_detect(evacuation_date,"2023-06-14") ~ "A",
    str_detect(evacuation_date,"2023-07-05") | str_detect(evacuation_date,"2023-07-07") ~ "B",
    str_detect(evacuation_date,"2023-08-04") | str_detect(evacuation_date,"2023-08-07") ~ "C",
    str_detect(evacuation_date,"2023-09-29") | str_detect(evacuation_date,"2023-10-02") ~ "D",
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "E",
    TRUE ~ "no")) %>%
   group_by(plot, gapfill_group) %>%
    select(-evacuation_date) %>%
    dplyr::summarise(across(everything(), sd), .groups = "keep") %>%
  rename(pH_sd = pH) %>%
    ungroup() %>%
  dplyr::select(-date)

ph_sw_gapfill <- mean_ph_l1_sw_to_gapfill %>%
  full_join(sd_ph_l1_sw_to_gapfill, by = c("plot", "gapfill_group")) %>%
  select(-gapfill_group)

##
#All the rest: 
means_ph_l1_pw_grids <- ph_l1_pw_grids %>%
filter(plot != "Control" | 
    (plot == "Control" & !evacuation_date %in% as.Date(c("2023-06-12", "2023-06-14", "2023-07-05", "2023-07-07", "2023-08-04", "2023-08-07","2023-12-05", "2023-12-08")))) %>%
  filter(plot != "Saltwater" | 
    (plot == "Saltwater" & !evacuation_date %in% as.Date(c("2023-06-12", "2023-06-14", "2023-07-05", "2023-07-07", "2023-08-04", "2023-08-07","2023-09-29", "2023-10-02", "2023-12-05", "2023-12-08")))) %>%
    filter(plot != "Freshwater" | 
    (plot == "Freshwater" & !evacuation_date %in% as.Date(c("2023-06-12", "2023-06-14", "2023-08-04", "2023-08-07","2023-12-05", "2023-12-08")))) %>%
  mutate(evacuation_date = as.numeric(evacuation_date),
         datetime= as.numeric(datetime)) %>%
  group_by(plot, evacuation_date) %>%
  dplyr::summarise(across(everything(), mean), .groups = "keep") %>%
  mutate(datetime = as.POSIXct(datetime, origin = "1970-01-01"),
         evacuation_date = as_date(evacuation_date))


sds_ph_l1_pw_grids <- ph_l1_pw_grids %>%
  select(-datetime) %>%
 filter(plot != "Control" | 
    (plot == "Control" & !evacuation_date %in% as.Date(c("2023-06-12", "2023-06-14", "2023-07-05", "2023-07-07", "2023-08-04", "2023-08-07","2023-12-05", "2023-12-08")))) %>%
  filter(plot != "Saltwater" | 
    (plot == "Saltwater" & !evacuation_date %in% as.Date(c("2023-06-12", "2023-06-14", "2023-07-05", "2023-07-07", "2023-08-04", "2023-08-07","2023-09-29", "2023-10-02", "2023-12-05", "2023-12-08"))))  %>%
    filter(plot != "Freshwater" | 
    (plot == "Freshwater" & !evacuation_date %in% as.Date(c("2023-06-12", "2023-06-14", "2023-08-04", "2023-08-07","2023-12-05", "2023-12-08")))) %>%
  mutate(evacuation_date = as.numeric(evacuation_date)) %>%
  group_by(plot, evacuation_date) %>%
  dplyr::summarise(across(everything(), sd), .groups = "keep") %>%
  mutate(evacuation_date = as_date(evacuation_date)) %>%
  rename(pH_sd = pH) %>%
  dplyr::select(-date)

merging_dfs <- list(means_ph_l1_pw_grids, sds_ph_l1_pw_grids, ph_c_gapfill, ph_sw_gapfill, ph_fw_gapfill)

pw_ph <- purrr::reduce(merging_dfs, full_join) %>%
   dplyr::select(-date)

# write out for plotting:
saveRDS(pw_ph, "~/GitHub/TEMPEST-1-porewater/data/averaged_ph_porewater.rds")

#note this has two dates close in 2025 that we might need to merge together at some point 
```

2 b. Now, get pH at the level as the DOC in terms of per month 

```{r ph avg std by month YEAR 2}


ph_yr2 <-  ph_2023 %>%
  mutate(date = as_date(datetime)) %>%
  select(plot, grid, date, pH) %>%
  drop_na() %>% 
  filter(between(date, startstudydate_yr2, endstudydate_yr2)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  # fix the sampling dates in June before the event to the month before
  mutate(adj_month = case_when(adj_month == 1 & date < EventStart2023 ~ adj_month == 0,
                               TRUE ~ adj_month)) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, pH, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

ph_yr2_2024 <- ph_2024 %>%
  mutate(date = as_date(datetime)) %>%
  select(plot, grid, date, pH) %>%
  drop_na() %>% #
  filter(between(date, startstudydate_yr2, endstudydate_yr2)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  mutate(adj_month = case_when(adj_month == 0 ~ 12,
                               TRUE ~ adj_month )) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, pH, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

ph_y2_all <- ph_yr2 %>%
  full_join(ph_yr2_2024)

  

saveRDS(ph_y2_all, "~/GitHub/TEMPEST-1-porewater/data/averaged_ph_porewater_TMP2.rds")
```

```{r ph avg std by month YEAR 3}


ph_yr3 <- ph_2024 %>%
  mutate(date = as_date(datetime)) %>%
  select(plot, grid, date, pH) %>%
  drop_na() %>% 
  filter(between(date, startstudydate_yr3, endstudydate_yr3)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  # fix the sampling dates in June before the event to the month before
  mutate(adj_month = case_when(adj_month == 1 & date < EventStart2023 ~ adj_month == 0,
                               TRUE ~ adj_month)) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, pH, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

  
ph_yr3_2025 <- ph_2025 %>%
  mutate(date = as_date(datetime)) %>%
  select(plot, grid, date, pH) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr3, endstudydate_yr3)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  mutate(adj_month = case_when(adj_month == 0 ~ 12,
                               TRUE ~ adj_month )) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, pH, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

ph_y3_all <- ph_yr3 %>%
  full_join(ph_yr3_2025)

  

saveRDS(ph_y3_all, "~/GitHub/TEMPEST-1-porewater/data/averaged_ph_porewater_TMP3.rds")
```





# Graveyard


## CEC from pre event soils

```{r cec}


soils_directory <- "https://docs.google.com/spreadsheets/d/1a-4Nt7jorO301bi3NF9ebGhwNx_hVjPG/edit?gid=456381663#gid=456381663"

directory= file.path(here::here() %>% dirname(), 'TEMPEST-1-Porewater/data_do_not_commit/')

file_path = file.path(directory,"UGA_soils_analysis_2020.xlsx")

googledrive::drive_download(soils_directory, path= file_path, overwrite = TRUE)
```

```{r cec 2}

soil_data <- readxl::read_excel("~/GitHub/TEMPEST-1-Porewater/data_do_not_commit//UGA_soils_analysis_2020.xlsx", skip = 8) %>%
  mutate(plot = case_when(stringr::str_detect(Sample, "C:") ~ "Control",
                          stringr::str_detect(Sample, "F:") ~ "Freshwater",
                          stringr::str_detect(Sample, "S:") ~ "Saltwater",
                          TRUE ~ NA),
         depth = case_when(stringr::str_detect(Sample, " 0-5") ~ "0-5",
                           stringr::str_detect(Sample, "5-10") ~ "5-10",
                           stringr::str_detect(Sample, "10-15") ~ "10-15",
                           TRUE ~ "deeper than 15cm"))  %>%
  filter(depth != "deeper than 15cm") %>%
  select(-Lab, - Sample) %>%
    group_by(plot) %>%
      dplyr::summarise(across(everything(), list(mean = ~mean(., na.rm=TRUE), sd = ~sd(., na.rm=TRUE))), .groups = "keep") %>%
   ungroup() %>%
  mutate(depth_mean = "0-15cm") %>%
  select(-depth_sd)

soil_data_bhm <- soil_data %>%
  filter(plot == "Control") %>%
  select(plot, depth_mean, `Base
Saturation_mean`, `Base
Saturation_sd`, CEC_mean, CEC_sd, `pH 2_mean`, `pH 2_sd`, C_mean, C_sd, Na_mean, Na_sd, 
         K_mean, K_sd, Ca_mean, Ca_sd, Mg_mean, Mg_sd, Fe_mean, Fe_sd) %>%
  rename(Base_Saturation_percent = `Base
Saturation_mean`,
         Base_Saturation_percent_stdev = `Base
Saturation_sd`,
         CEC_meqper100g = CEC_mean,
         CEC_meqper100g_stdev = CEC_sd,
         pH = `pH 2_mean`, 
         pH_stdev = `pH 2_sd`,
         SoilC_percent = C_mean, 
         SoilC_percent_stdev = C_sd, 
         Na_ppm = Na_mean, 
         Na_ppm_stdev = Na_sd, 
         K_ppm = K_mean, 
         K_ppm_stdev = K_sd, 
         Ca_ppm = Ca_mean, 
         Ca_ppm_stdev = Ca_sd, 
         Mg_ppm = Mg_mean, 
         Mg_ppm_stdev = Mg_sd, 
         Fe_ppm = Fe_mean, 
         Fe_ppm_stdev = Fe_sd)

write_csv(soil_data_bhm, "~/GitHub/tempest_ionic_strength/BHM/Data/TEMPEST_soildata_2020.csv")

  
```



### Prep data for figure 1: Porewater DOC averages/std by plot and month
Generate for easy comparison between event years and with laboratory data sets.
Can simplify this from previous iterations (see `graveyard_analysis_code.Rmd`) because we just want a monthly timeframe. 
```{r doc avg std by month YEAR 1}

pw_doc_yr1 <- doc_l1_2022 %>%
  mutate(date = as_date(date)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr1, endstudydate)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  # fix the sampling dates in June before the event to the month before
  mutate(adj_month = case_when(adj_month == 1 & date < EventStart2022 ~ adj_month == 0,
                               TRUE ~ adj_month)) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

pw_doc_yr1_2023 <- doc_l1_2023 %>%
  mutate(date = as_date(date)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr1, endstudydate_yr1)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  mutate(adj_month = case_when(adj_month == 0 ~ 12,
                               TRUE ~ adj_month )) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

pw_doc_y1_all <- pw_doc_yr1 %>%
  full_join(pw_doc_yr1_2023) %>%
  mutate(month = case_when(adj_month <= 7 ~ adj_month + 5,
                           adj_month > 7 ~ adj_month - 7))
  

saveRDS(pw_doc_y1_all, "~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP1_Figure1.rds")
```

```{r doc avg std by month YEAR 2}


pw_doc_yr2 <- doc_l1_2023 %>%
  mutate(date = as_date(evacuation_datetime)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr2, endstudydate_yr2)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  # fix the sampling dates in June before the event to the month before
  mutate(adj_month = case_when(adj_month == 1 & date < EventStart2023 ~ adj_month == 0,
                               TRUE ~ adj_month)) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

pw_doc_yr2_2024 <- doc_l1_2024 %>%
  mutate(date = as_date(date)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr2, endstudydate_yr2)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  mutate(adj_month = case_when(adj_month == 0 ~ 12,
                               TRUE ~ adj_month )) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

pw_doc_y2_all <- pw_doc_yr2 %>%
  full_join(pw_doc_yr2_2024) %>%
  mutate(month = case_when(adj_month <= 7 ~ adj_month + 5,
                           adj_month > 7 ~ adj_month - 7))

  

saveRDS(pw_doc_y2_all, "~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP2_Figure1.rds")
```

```{r doc avg std by month YEAR 3}


pw_doc_yr3 <- doc_l1_2024 %>%
  mutate(date = as_date(evacuation_datetime)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr3, endstudydate_yr3)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  # fix the sampling dates in June before the event to the month before
  mutate(adj_month = case_when(adj_month == 1 & date < EventStart2023 ~ adj_month == 0,
                               TRUE ~ adj_month)) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

  
pw_doc_yr3_2025 <- doc_l1_2025 %>%
  mutate(date = as_date(date)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr3, endstudydate_yr3)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  mutate(adj_month = case_when(adj_month == 0 ~ 12,
                               TRUE ~ adj_month )) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

pw_doc_y3_all <- pw_doc_yr3 %>%
  full_join(pw_doc_yr3_2025) %>%
  mutate(month = case_when(adj_month <= 7 ~ adj_month + 5,
                           adj_month > 7 ~ adj_month - 7))

  

saveRDS(pw_doc_y3_all, "~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP3_Figure1.rds")
```

```{r group all doc}

pw_doc_all <- pw_doc_y1_all %>%
  full_join(pw_doc_y2_all) %>%
full_join(pw_doc_y3_all) %>%
  mutate(stamp =lubridate::my(interaction(month, year, sep = " "))) 
  

saveRDS(pw_doc_all, "~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_ALLYEARS_Figure1.rds")

```



```{r save doc precip}
pw_doc_all_wprecip <- pw_doc_all %>%
  left_join(precip, by = c("month", "year"))


write_csv(pw_doc_all_wprecip, "~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_allyrs_withprecip.csv")

```



# Experimental Data Analysis for mechanistic relationships
Workflow 
1) Exploratory Data Analysis: look at DOC vs driver variable relationships in the ColEx data 
2) Bring in a priori hypothesis-based knowledge from table of mechanisms to reduce variable numbers.  https://docs.google.com/spreadsheets/d/1DHOeCKPpyUJbPBAVBy6-vegGymVbECd7rN9bFWm7Y0M/edit?usp=sharing
3) Try Bayesian Networks to discover probabilistic causal structures with bnlearn 
  3a) scale data (try z-scores)
  3b) check multicollinearity with VIFs
  3c) include mechanistic constraints (e.g., Fe/OC influence DOC)
  
### Exploratory data analysis
```{r doc stats}

hist(doc_l1_pw_grid_2022_2026$doc_mg_l)

doc_comparison <- doc_l1_pw_grid_2022_2026 %>%
  filter(plot != "Saltwater") %>%
  mutate(logdoc = log(doc_mg_l),  #DOC is right skewed
  gridID = interaction(plot, grid, sep = "_"),
  month = lubridate::month(evacuation_date),
  year = lubridate::year(evacuation_date),
  dateID = interaction(month, year, sep = "_"))# create unique grid IDs (per plot)

hist(doc_comparison$logdoc)

doc_lme <- lme(
  logdoc ~ plot ,
  random = ~ 1 | gridID, # random effect of grids in the plot
  correlation = corAR1(form = ~ 1 | gridID), # repeated monthly observations are temporally autocorrelated
  data = doc_comparison 
)

summary(doc_lme)

plot(doc_lme)

qqnorm(resid(doc_lme))
qqline(resid(doc_lme))

#one factor model

EMM <- emmeans(doc_lme, ~ plot)

EMM

### pairwise comparisons
pairs(EMM)

```

```{r load ColEx data}
ColEx <- readxl::read_excel("~/GitHub/tempest_ionic_strength/Data/Exp 3 Column Experiment ColEx Data/ColEx_data_allreps_formatted.xlsx") %>% #drop initial for this
  filter(Timepoint != 0)

```

```{r exploratory data analysis}
library(ggcorrplot)

ColEx_corr <- ColEx %>%
  select(-Treatment, -Column, - Timepoint)

cor_matrix <- cor(ColEx_corr, use = "complete.obs")

ggcorrplot(cor_matrix)

```
There are no other variables that stick out than the ones we'd pick a priori that are highly correlated to DOC. 


```{r check lags}
# Check lagged correlations
library(tseries)
ccf(ColEx$pH, ColEx$doc_mgperL, lag.max = 12)
ccf(ColEx$conductivity_uSpercm, ColEx$doc_mgperL, lag.max = 12)

```
 Interpretation:
 - Check lag 0: Strong direct correlation between variable and DOC at the same time point.
 - Lags > 0: Positive correlations here mean variable changes precede DOC changes by the lag amount.
 - Lags < 0: Negative correlations here mean DOC changes precede variable changes by the lag amount.
 - Dashed horizontal lines = confidence bounds, significant correlations occur when values go beyond those lines.

This analysis might be really cool to do with the actual plot data and not the experiment- going to leave out of the model here. 

Narrow down to final variables


```{r check colinearity}
library(car)

# Fit a linear model to check VIF
lm_model <- lm(DOC ~ pH + Fe_OC_ratio + Conductivity, data = scaled_data)
car::vif(lm_model)

# General guideline: VIF > 10 indicates problematic multicollinearity
```


```{r do the Dynamic Bayesian Networks}

library(bnlearn)

# Whitelist specific relationships based on prior knowledge
whitelist <- data.frame(
  from = c("Fe_OC_ratio", "pH"),
  to = c("DOC", "DOC")
)

dag <- hc(data, whitelist = whitelist)
plot(dag)
```


```{r prelim plots}


cdom_indicies_divided_all %>%
    group_by(plot, timedate_bin) %>%
    dplyr::summarise(across(everything(), list(mean = ~mean(., na.rm=TRUE), sd = ~sd(., na.rm=TRUE))), .groups = "keep") %>%
   ungroup() %>%
  ggplot(aes(x=factor(timedate_bin, levels = c("Pre","Mid","Pre Cond Threshold", "Post Cond Threshold")), 
               y=A_T_mean)) +
  geom_col() +
  geom_errorbar(aes(ymin = A_T_mean - A_T_sd, ymax= A_T_mean + A_T_sd), width = 0.2) +
  facet_wrap( ~ plot)

cdom_indicies_divided_all %>%
    group_by(plot, timedate_bin) %>%
    dplyr::summarise(across(everything(), list(mean = ~mean(., na.rm=TRUE), sd = ~sd(., na.rm=TRUE))), .groups = "keep") %>%
   ungroup() %>%
  ggplot(aes(x=factor(timedate_bin, levels = c("Pre","Mid","Pre Cond Threshold", "Post Cond Threshold")), 
               y=SUVA254_mean)) +
  geom_col() +
  geom_errorbar(aes(ymin = SUVA254_mean - SUVA254_sd, ymax= SUVA254_mean + SUVA254_sd), width = 0.2) +
  facet_wrap( ~ plot) +
  xlab("Sampling Timeframe") +
  ylab("SUVA254")+
  theme_classic()

biplot(eemspca)

```

##



