---
title: "Porewater DOC Manuscript Analysis"
author: "AMP"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(lubridate)
library(tidyverse)
library(car)
library(emmeans)
library(lme4)
library(lmerTest)  
library(ggplot2)    # For diagnostic plots
library(broom)      # For tidy model outputs

#double check your wd
getwd()

#clean out your environment 
rm(list = ls())
```

### Set the study dates

```{r event and study dates}
endstudydate = lubridate::as_date("2024-01-31")
startstudydate = lubridate::as_date("2022-05-01")

endstudydate_yr1 = lubridate::as_date("2023-05-31")
startstudydate_yr1 = lubridate::as_date("2022-05-01")

endstudydate_yr2 = lubridate::as_date("2024-05-31")
startstudydate_yr2 = lubridate::as_date("2023-05-01") 

endstudydate_yr3 = lubridate::as_date("2025-05-31")
startstudydate_yr3 = lubridate::as_date("2024-05-01") 

year1_start = lubridate::as_date("2022-01-01")
year1_stop = lubridate::as_date("2022-12-31")

year2_start = lubridate::as_date("2023-01-01")
year2_stop = lubridate::as_date("2023-12-31")

year3_start = lubridate::as_date("2024-01-01")
year3_stop = lubridate::as_date("2024-12-31")

EventStart2022 = lubridate::as_date("2022-06-22")

WaterDeliveryStart2022 = as.POSIXct("2022-06-22 05:30:00", tz = "EST")
WaterDeliveryStop2022 = as.POSIXct("2022-06-22 14:30:00", tz = "EST")

EventStart2023 = lubridate::as_date("2022-06-06")

WaterDeliveryStart1 = as.POSIXct("2023-06-06 05:30:00", tz = "EST")
WaterDeliveryStop1 = as.POSIXct("2023-06-06 14:30:00", tz = "EST")

WaterDeliveryStart2 = as.POSIXct("2023-06-07 05:30:00", tz = "EST")
WaterDeliveryStop2 = as.POSIXct("2023-06-07 14:30:00", tz = "EST")

plot_order <- c('Control', 'Freshwater','Saltwater')
time_order <- c('Pre', 'Mid','Post')
```

### DOC

1) Import DOC grid data for porewater from 2022 and 2023 events. Then, bind them together and take the average and standard deviation of the values for each event. Do this by the evacuation date, as this will capture multiple tries within the same dateframe. For this particular analysis, it's ok to take the average of the date time of sampling, because we're just going to be using this data to plot through time and do some simple summary statistics.  

```{r read in doc}

#Porewater DOC data, from L1 google drive: 

doc_l1_2022 <- read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_GRIDSONLY_NPOCTDN_2022_L1.csv")
  
doc_l1_2023 <- read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_GRIDSONLY_NPOCTDN_2023_L1.csv")

doc_l1_2024 <- read_csv("~/GitHub/TEMPEST_Porewater/processing_scripts/DOC/TEMPORARY/01_TMP_PW_NPOCTDN_2024_L1.csv")

doc_l1_2025 <- read_csv("~/GitHub/TEMPEST_Porewater/processing_scripts/DOC/TEMPORARY/01_TMP_PW_NPOCTDN_2025_L1.csv")
```
```{r combine years}
doc_l1_pw_grids <- doc_l1_2022 %>%
  full_join(doc_l1_2023) %>%
  full_join(doc_l1_2024) %>%
  full_join(doc_l1_2025) %>%
  mutate(evacuation_date = as_date(evacuation_datetime)) %>%
  select(plot, grid, evacuation_date, collection_datetime, doc_mg_l, tdn_mg_l) %>%
  drop_na() #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
```
Porewater DOC data manipulations for Figure 1 
<!-- include only year 1 data:  -->

<!-- ```{r DOC data crunch} -->
<!-- #1) filter dataframe to YR1 Event data -->

<!-- doc_TMP1 <- doc_l1_pw_grids %>% -->
<!--   filter(between(evacuation_date, startstudydate_yr1, endstudydate_yr1))  -->

<!-- saveRDS(doc_TMP1, "~/GitHub/TEMPEST-1-porewater/data/doc_porewater_grids_TMP1_forfig1.rds") -->

<!-- #2) get min, max, IQR for control and freshwater plots -->

<!-- doc_iqr <- doc_TMP1 %>% -->
<!--   filter(plot != "Saltwater") %>% -->
<!--   group_by(plot) %>% -->
<!--   summarise(min_doc_mg_l = min(doc_mg_l), -->
<!--             max_doc_mg_l = max(doc_mg_l), -->
<!--             iqr_doc_mg_l = IQR(doc_mg_l), -->
<!--             q1_doc_mg_l = quantile(doc_mg_l, 0.25), -->
<!--             q3_doc_mg_l = quantile(doc_mg_l, 0.75)) -->

<!-- saveRDS(doc_iqr, "~/GitHub/TEMPEST-1-porewater/data/doc_porewater_summarystats_forfig1.rds") -->
<!-- ``` -->

# Main Message 1: One salinity exposure permanently changes the soil DOC component of the carbon cycle. 

### Prep data for figure 1: Porewater DOC averages/std by plot and month
Generate for  comparison with laboratory data sets.
Can simplify this from previous iterations because we just want a monthly timeframe. 
```{r doc avg std by month YEAR 1}

pw_doc_yr1 <- doc_l1_2022 %>%
  mutate(date = as_date(date)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr1, endstudydate)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  # fix the sampling dates in June before the event to the month before
  mutate(adj_month = case_when(adj_month == 1 & date < EventStart2022 ~ adj_month == 0,
                               TRUE ~ adj_month)) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

pw_doc_yr1_2023 <- doc_l1_2023 %>%
  mutate(date = as_date(date)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr1, endstudydate_yr1)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  mutate(adj_month = case_when(adj_month == 0 ~ 12,
                               TRUE ~ adj_month )) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

pw_doc_y1_all <- pw_doc_yr1 %>%
  full_join(pw_doc_yr1_2023)
  

saveRDS(pw_doc_y1_all, "~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP1_Figure1.rds")
```

```{r doc avg std by month YEAR 2}


pw_doc_yr2 <- doc_l1_2023 %>%
  mutate(date = as_date(evacuation_datetime)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr2, endstudydate_yr2)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  # fix the sampling dates in June before the event to the month before
  mutate(adj_month = case_when(adj_month == 1 & date < EventStart2023 ~ adj_month == 0,
                               TRUE ~ adj_month)) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

pw_doc_yr2_2024 <- doc_l1_2024 %>%
  mutate(date = as_date(date)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr2, endstudydate_yr2)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  mutate(adj_month = case_when(adj_month == 0 ~ 12,
                               TRUE ~ adj_month )) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

pw_doc_y2_all <- pw_doc_yr2 %>%
  full_join(pw_doc_yr2_2024)

  

saveRDS(pw_doc_y2_all, "~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP2_Figure1.rds")
```

```{r doc avg std by month YEAR 3}


pw_doc_yr3 <- doc_l1_2024 %>%
  mutate(date = as_date(evacuation_datetime)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr3, endstudydate_yr3)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  # fix the sampling dates in June before the event to the month before
  mutate(adj_month = case_when(adj_month == 1 & date < EventStart2023 ~ adj_month == 0,
                               TRUE ~ adj_month)) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

  
pw_doc_yr3_2025 <- doc_l1_2025 %>%
  mutate(date = as_date(date)) %>%
  select(plot, grid, date, doc_mg_l, tdn_mg_l) %>%
  drop_na() %>% #dont need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
  filter(between(date, startstudydate_yr3, endstudydate_yr3)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  mutate(adj_month = case_when(adj_month == 0 ~ 12,
                               TRUE ~ adj_month )) %>%
  group_by(adj_month, year, plot) %>%
  dplyr::select(plot, doc_mg_l, tdn_mg_l, adj_month, year) %>%
  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup()

pw_doc_y3_all <- pw_doc_yr3 %>%
  full_join(pw_doc_yr3_2025)

  

saveRDS(pw_doc_y3_all, "~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP3_Figure1.rds")
```

# Main Message 2: The possible mechanisms at play here are all changing in concert, which clouds predictability across systems and scales.

### precip

```{r precip}

rain_month_yr1 <- read_csv("~/GitHub/TEMPEST-1-porewater/processed data/TMP_rain_dailyavg_2022-2023.csv") %>%
 filter(between(date, startstudydate_yr1, endstudydate_yr1))  %>%
mutate(month = lubridate::month(date),
         year = lubridate::year(date),
        # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  group_by(adj_month, year) %>%
  summarise(mean_monthly_rain = mean(daily_total_rain_cm),
            sd_monthly_rain = sd(daily_total_rain_cm))


rain_month_yr2 <- read_csv("~/GitHub/TEMPEST-1-porewater/processed data/TMP_rain_dailyavg_2022-2023.csv") %>%
 filter(between(date, startstudydate_yr2, endstudydate_yr2))  %>%
mutate(month = lubridate::month(date),
         year = lubridate::year(date),
        # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  group_by(adj_month, year) %>%
  summarise(mean_monthly_rain = mean(daily_total_rain_cm),
            sd_monthly_rain = sd(daily_total_rain_cm))

```



#### Relationship between conductivity and DOC

```{r cond vs doc load data}
cond_doc <-  read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_fluxes/Lysimeter_EC_DOC_forcorrelations_SALTWATER.csv") %>%
  dplyr::filter(collection_datetime > WaterDeliveryStart2022) %>%
  ungroup() %>%
  filter(plot == "Saltwater") %>%
  group_by(collection_date) %>%
  dplyr::select(collection_date, mean_EC_uScm_15cm, max_EC_uScm_15cm, doc_mg_l) %>%
 #  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup() %>%
  mutate(date = as_date(collection_date)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  mutate(adj_month = case_when(adj_month == 0 ~ 12,
                               adj_month == 1 & collection_date == "2023-06-03" ~ 12,
                               TRUE ~ adj_month)) 
 # dplyr::filter(adj_month != 1)

saveRDS(cond_doc, "~/GitHub/TEMPEST-1-porewater/data/DOC_cond_data_for_relationships_SALT.rds")
```

do this for the control plot, for comparison

```{r cond vs doc load data Control}
cond_doc_C <-  read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_fluxes/Lysimeter_EC_DOC_forcorrelations_CONTROL.csv") %>%
  dplyr::filter(collection_datetime > WaterDeliveryStart2022) %>%
  ungroup() %>%
  filter(plot == "Control") %>%
  group_by(collection_date) %>%
  dplyr::select(collection_date, mean_EC_uScm_15cm, max_EC_uScm_15cm, doc_mg_l) %>%
 #  dplyr::summarise(across(everything(), list(mean = ~mean(.), sd = ~sd(.))), .groups = "keep") %>%
   ungroup() %>%
  mutate(date = as_date(collection_date)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  mutate(adj_month = case_when(adj_month == 0 ~ 12,
                               adj_month == 1 & collection_date == "2023-06-03" ~ 12,
                               TRUE ~ adj_month)) 
 # dplyr::filter(adj_month != 1)

saveRDS(cond_doc_C, "~/GitHub/TEMPEST-1-porewater/data/DOC_cond_data_for_relationships_CONTROLPLOT.rds")
```

```{r make line for fig 5}

start_a <- max(cond_doc$doc_mg_l) 
start_b <- -0.001

# Fit the model using nls
fit_nls <- nls(doc_mg_l ~ a * exp(b * mean_EC_uScm_15cm), data = cond_doc, start = list(a = start_a, b = start_b))

# Summary of the model
summary(fit_nls)
# Extract fitted values
coefficients <- coef(fit_nls)
a <- coefficients["a"]
b <- coefficients["b"]

# Create a prediction data frame
new_data <- data.frame(mean_EC_uScm_15cm = seq(min(cond_doc$mean_EC_uScm_15cm), max(cond_doc$mean_EC_uScm_15cm), length.out = 100))
new_data$pred <- predict(fit_nls, newdata = new_data)

# Compute standard error of the fit
se_fit <- sqrt(sum(residuals(fit_nls)^2) / df.residual(fit_nls))
new_data$lower <- new_data$pred - 1.96 * se_fit
new_data$upper <- new_data$pred + 1.96 * se_fit

```
# multivariate prep

### Plot level pH averages and stds
```{r load pH data}

ph <-  read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_pH_2023_L1.csv") 
ph_2024 <-  read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_pH_2024_L1.csv") 

```

```{r ph all years}
ph_l1_pw_grids <- ph %>%
  full_join(ph_2024) %>%
  dplyr::select(plot, datetime, evacuation_date, pH) %>%
  drop_na() %>%
  mutate(date = lubridate::date(datetime)) %>%
  group_by(plot, evacuation_date)
  
```

```{r ph avg std}
# Some dates not enough water was collected to have multiple samples to get averages and sds for that evacuation period. Most of the time, those can be combined with sampling dates just before or just after as we go out over several multi-day periods. This, however isn't consistent across the FW, EW, or C plots, so it will be a bit tricky to do the gap filling. To make the gap filling more accurate, we'll take the average of those dates, so for example if the dates we're merging is 2022-11-25 and 2022-11-28 we'll have the date be the 26th (but we'll let R do that math).

#First, isolate the trouble dates/times: 

#Control: 
mean_ph_l1_c_to_gapfill <- ph_l1_pw_grids %>%
  filter(plot == "Control") %>%
  filter(evacuation_date %in% c("2023-06-12", "2023-06-14", "2023-07-05", "2023-07-07", "2023-08-04", "2023-08-07","2023-12-05", "2023-12-08")) %>%
  mutate(gapfill_group = case_when( 
    str_detect(evacuation_date,"2023-06-12") | str_detect(evacuation_date,"2023-06-14") ~ "A",
    str_detect(evacuation_date,"2023-07-05") | str_detect(evacuation_date,"2023-07-07") ~ "B",
    str_detect(evacuation_date,"2023-08-04") | str_detect(evacuation_date,"2023-08-07") ~ "C",
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "D",
    TRUE ~ "no")) %>%
   group_by(plot, gapfill_group) %>%
    dplyr::summarise(across(everything(), mean), .groups = "keep") %>%
  mutate(datetime = as.POSIXct(datetime, origin = "1970-01-01"),
         evacuation_date = as_date(evacuation_date)) %>%
  ungroup()%>%
  dplyr::select(-date)

sd_ph_l1_c_to_gapfill <- ph_l1_pw_grids %>%
  select(-datetime) %>%
  filter(plot == "Control") %>%
 filter(evacuation_date %in% c("2023-06-12", "2023-06-14", "2023-07-05", "2023-07-07", "2023-08-04", "2023-08-07","2023-12-05", "2023-12-08")) %>%
  mutate(gapfill_group = case_when( 
    str_detect(evacuation_date,"2023-06-12") | str_detect(evacuation_date,"2023-06-14") ~ "A",
    str_detect(evacuation_date,"2023-07-05") | str_detect(evacuation_date,"2023-07-07") ~ "B",
    str_detect(evacuation_date,"2023-08-04") | str_detect(evacuation_date,"2023-08-07") ~ "C",
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "D",
    TRUE ~ "no")) %>%
   group_by(plot, gapfill_group) %>%
    select(-evacuation_date) %>%
    dplyr::summarise(across(everything(), sd), .groups = "keep") %>%
  rename(pH_sd = pH) %>%
    ungroup() %>%
  dplyr::select(-date)

ph_c_gapfill <- mean_ph_l1_c_to_gapfill %>%
  full_join(sd_ph_l1_c_to_gapfill, by = c("plot", "gapfill_group")) %>%
  select(-gapfill_group)

#Freshwater: 
mean_ph_l1_fw_to_gapfill <- ph_l1_pw_grids %>%
  filter(plot == "Freshwater") %>%
  filter(evacuation_date %in% c("2023-06-12", "2023-06-14", "2023-08-04", "2023-08-07","2023-12-05", "2023-12-08")) %>%
  mutate(gapfill_group = case_when( 
    str_detect(evacuation_date,"2023-06-12") | str_detect(evacuation_date,"2023-06-14") ~ "A",
    str_detect(evacuation_date,"2023-08-04") | str_detect(evacuation_date,"2023-08-07") ~ "B",
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "C",
    TRUE ~ "no")) %>%
   group_by(plot, gapfill_group) %>%
    dplyr::summarise(across(everything(), mean), .groups = "keep") %>%
  mutate(datetime = as.POSIXct(datetime, origin = "1970-01-01"),
         evacuation_date = as_date(evacuation_date)) %>%
  ungroup()%>%
  dplyr::select(-date)

sd_ph_l1_fw_to_gapfill <- ph_l1_pw_grids %>%
  select(-datetime) %>%
  filter(plot == "Freshwater") %>%
  filter(evacuation_date %in% c("2023-06-12", "2023-06-14", "2023-08-04", "2023-08-07","2023-12-05", "2023-12-08")) %>%
  mutate(gapfill_group = case_when( 
    str_detect(evacuation_date,"2023-06-12") | str_detect(evacuation_date,"2023-06-14") ~ "A",
    str_detect(evacuation_date,"2023-08-04") | str_detect(evacuation_date,"2023-08-07") ~ "B",
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "C",
    TRUE ~ "no")) %>%
   group_by(plot, gapfill_group) %>%
    select(-evacuation_date) %>%
    dplyr::summarise(across(everything(), sd), .groups = "keep") %>%
  rename(pH_sd = pH) %>%
    ungroup() %>%
  dplyr::select(-date)

ph_fw_gapfill <- mean_ph_l1_fw_to_gapfill %>%
  full_join(sd_ph_l1_fw_to_gapfill, by = c("plot", "gapfill_group")) %>%
  select(-gapfill_group)

#Seawater: 
mean_ph_l1_sw_to_gapfill <- ph_l1_pw_grids %>%
  filter(plot == "Saltwater") %>%
  filter(evacuation_date %in% c("2023-06-12", "2023-06-14", "2023-07-05", "2023-07-07", "2023-08-04", "2023-08-07","2023-09-29", "2023-10-02", "2023-12-05", "2023-12-08")) %>%
  mutate(gapfill_group = case_when( 
    str_detect(evacuation_date,"2023-06-12") | str_detect(evacuation_date,"2023-06-14") ~ "A",
    str_detect(evacuation_date,"2023-07-05") | str_detect(evacuation_date,"2023-07-07") ~ "B",
    str_detect(evacuation_date,"2023-08-04") | str_detect(evacuation_date,"2023-08-07") ~ "C",
    str_detect(evacuation_date,"2023-09-29") | str_detect(evacuation_date,"2023-10-02") ~ "D",
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "E",
    TRUE ~ "no")) %>%
   group_by(plot, gapfill_group) %>%
    dplyr::summarise(across(everything(), mean), .groups = "keep") %>%
  mutate(datetime = as.POSIXct(datetime, origin = "1970-01-01"),
         evacuation_date = as_date(evacuation_date)) %>%
  ungroup()%>%
  dplyr::select(-date)

sd_ph_l1_sw_to_gapfill <- ph_l1_pw_grids %>%
  select(-datetime) %>%
  filter(plot == "Saltwater") %>%
  filter(evacuation_date %in% c("2023-06-12", "2023-06-14", "2023-07-05", "2023-07-07", "2023-08-04", "2023-08-07","2023-09-29", "2023-10-02", "2023-12-05", "2023-12-08")) %>%
  mutate(gapfill_group = case_when( 
    str_detect(evacuation_date,"2023-06-12") | str_detect(evacuation_date,"2023-06-14") ~ "A",
    str_detect(evacuation_date,"2023-07-05") | str_detect(evacuation_date,"2023-07-07") ~ "B",
    str_detect(evacuation_date,"2023-08-04") | str_detect(evacuation_date,"2023-08-07") ~ "C",
    str_detect(evacuation_date,"2023-09-29") | str_detect(evacuation_date,"2023-10-02") ~ "D",
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "E",
    TRUE ~ "no")) %>%
   group_by(plot, gapfill_group) %>%
    select(-evacuation_date) %>%
    dplyr::summarise(across(everything(), sd), .groups = "keep") %>%
  rename(pH_sd = pH) %>%
    ungroup() %>%
  dplyr::select(-date)

ph_sw_gapfill <- mean_ph_l1_sw_to_gapfill %>%
  full_join(sd_ph_l1_sw_to_gapfill, by = c("plot", "gapfill_group")) %>%
  select(-gapfill_group)

##
#All the rest: 
means_ph_l1_pw_grids <- ph_l1_pw_grids %>%
filter(plot != "Control" | 
    (plot == "Control" & !evacuation_date %in% as.Date(c("2023-06-12", "2023-06-14", "2023-07-05", "2023-07-07", "2023-08-04", "2023-08-07","2023-12-05", "2023-12-08")))) %>%
  filter(plot != "Saltwater" | 
    (plot == "Saltwater" & !evacuation_date %in% as.Date(c("2023-06-12", "2023-06-14", "2023-07-05", "2023-07-07", "2023-08-04", "2023-08-07","2023-09-29", "2023-10-02", "2023-12-05", "2023-12-08")))) %>%
    filter(plot != "Freshwater" | 
    (plot == "Freshwater" & !evacuation_date %in% as.Date(c("2023-06-12", "2023-06-14", "2023-08-04", "2023-08-07","2023-12-05", "2023-12-08")))) %>%
  mutate(evacuation_date = as.numeric(evacuation_date),
         datetime= as.numeric(datetime)) %>%
  group_by(plot, evacuation_date) %>%
  dplyr::summarise(across(everything(), mean), .groups = "keep") %>%
  mutate(datetime = as.POSIXct(datetime, origin = "1970-01-01"),
         evacuation_date = as_date(evacuation_date))


sds_ph_l1_pw_grids <- ph_l1_pw_grids %>%
  select(-datetime) %>%
 filter(plot != "Control" | 
    (plot == "Control" & !evacuation_date %in% as.Date(c("2023-06-12", "2023-06-14", "2023-07-05", "2023-07-07", "2023-08-04", "2023-08-07","2023-12-05", "2023-12-08")))) %>%
  filter(plot != "Saltwater" | 
    (plot == "Saltwater" & !evacuation_date %in% as.Date(c("2023-06-12", "2023-06-14", "2023-07-05", "2023-07-07", "2023-08-04", "2023-08-07","2023-09-29", "2023-10-02", "2023-12-05", "2023-12-08"))))  %>%
    filter(plot != "Freshwater" | 
    (plot == "Freshwater" & !evacuation_date %in% as.Date(c("2023-06-12", "2023-06-14", "2023-08-04", "2023-08-07","2023-12-05", "2023-12-08")))) %>%
  mutate(evacuation_date = as.numeric(evacuation_date)) %>%
  group_by(plot, evacuation_date) %>%
  dplyr::summarise(across(everything(), sd), .groups = "keep") %>%
  mutate(evacuation_date = as_date(evacuation_date)) %>%
  rename(pH_sd = pH) %>%
  dplyr::select(-date)

merging_dfs <- list(means_ph_l1_pw_grids, sds_ph_l1_pw_grids, ph_c_gapfill, ph_sw_gapfill, ph_fw_gapfill)

pw_ph <- purrr::reduce(merging_dfs, full_join) %>%
   dplyr::select(-date)

# write out for plotting:
saveRDS(pw_ph, "~/GitHub/TEMPEST-1-porewater/data/averaged_ph_porewater.rds")
```


# teros conductivity plot averages

```{r teros EC}

pw_grids_ec <- read_rds("~/GitHub/TEMPEST-1-porewater/processed data/TMP_PW_EC_15cm_2022-2023_L2.rds")

pw_plot_ec <- pw_grids_ec %>%
  group_by(plot, depth, research_name, datetime_est, yday(datetime_est)) %>%
        summarise(datetime_est = mean(datetime_est),
                  EC_uScm = mean(EC_uScm, na.rm = TRUE),
                  .groups = "drop")

# write out for plotting:
saveRDS(pw_plot_ec, "~/GitHub/TEMPEST-1-porewater/data/averaged_by_plot_ec_teros_porewater.rds")

```




# Misc graveyard analyses

# Plot level DOC averages and stds
```{r doc avg stds} 
#can't get both mean and sd to work across everything for some reason, so calculate separate then recombine.  
# Some dates not enough water was collected to have multiple samples to get averages and sds for that evacuation period. Most of the time, those can be combined with sampling dates just before or just after as we go out over several multi-day periods. This, however isn't consistent across the FW, EW, or C plots, so it will be a bit tricky to do the gap filling. To make the gap filling more accurate, we'll take the average of those dates, so for example if the dates we're merging is 2022-11-25 and 2022-11-28 we'll have the date be the 26th (but we'll let R do that math).

#First, isolate the trouble dates/times:  

#Control: 
mean_doc_l1_c_to_gapfill <- doc_l1_pw_grids %>% 
select(-grid) %>% 
  filter(plot == "Control") %>% 
  filter(evacuation_date %in% c("2022-09-09", "2022-09-12", "2022-11-28", "2022-11-25", "2023-01-30", "2023-02-02", "2023-07-05", "2023-07-07", "2023-12-05", "2023-12-08")) %>% 
  mutate(gapfill_group = case_when(  
    str_detect(evacuation_date,"2022-09-09") | str_detect(evacuation_date,"2022-09-12") ~ "A", 
    str_detect(evacuation_date,"2022-11-28") | str_detect(evacuation_date,"2022-11-25") ~ "B", 
    str_detect(evacuation_date,"2023-01-30") | str_detect(evacuation_date,"2023-02-02") ~ "C", 
    str_detect(evacuation_date,"2023-07-05") | str_detect(evacuation_date,"2023-07-07") ~ "D", 
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "E", 
    TRUE ~ "no")) %>% 
   group_by(plot, gapfill_group) %>% 
    dplyr::summarise(across(everything(), mean), .groups = "keep") %>% 
  mutate(collection_datetime = as.POSIXct(collection_datetime, origin = "1970-01-01"), 
         evacuation_date = as_date(evacuation_date)) %>% 
  ungroup() 
 
sd_doc_l1_c_to_gapfill <- doc_l1_pw_grids %>% 
  select(-collection_datetime, -grid) %>% 
  filter(plot == "Control") %>% 
  filter(evacuation_date %in% c("2022-09-09", "2022-09-12", "2022-11-28", "2022-11-25", "2023-01-30", "2023-02-02", "2023-07-05", "2023-07-07", "2023-12-05", "2023-12-08")) %>% 
  mutate(gapfill_group = case_when(  
    str_detect(evacuation_date,"2022-09-09") | str_detect(evacuation_date,"2022-09-12") ~ "A", 
    str_detect(evacuation_date,"2022-11-28") | str_detect(evacuation_date,"2022-11-25") ~ "B", 
    str_detect(evacuation_date,"2023-01-30") | str_detect(evacuation_date,"2023-02-02") ~ "C", 
    str_detect(evacuation_date,"2023-07-05") | str_detect(evacuation_date,"2023-07-07") ~ "D", 
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "E", 
    TRUE ~ "no")) %>% 
   group_by(plot, gapfill_group) %>% 
    select(-evacuation_date) %>% 
    dplyr::summarise(across(everything(), sd), .groups = "keep") %>% 
  rename(doc_mg_l_sd = doc_mg_l, 
         tdn_mg_l_sd = tdn_mg_l) %>% 
    ungroup()  
 
c_gapfill <- mean_doc_l1_c_to_gapfill %>% 
  full_join(sd_doc_l1_c_to_gapfill, by = c("plot", "gapfill_group")) %>% 
  select(-gapfill_group) 
 
#Estuarine Water 
mean_doc_l1_sw_to_gapfill <- doc_l1_pw_grids %>% 
  select(-grid) %>% 
  filter(plot == "Saltwater") %>% 
  filter(evacuation_date %in% c("2022-07-18", "2022-07-14", "2022-08-08", "2022-08-05","2022-09-09", "2022-09-12","2022-11-28", "2022-11-25", "2023-12-05", "2023-12-08", "2023-01-30", "2023-02-02")) %>% 
  mutate(gapfill_group = case_when(  
    str_detect(evacuation_date,"2022-07-18") | str_detect(evacuation_date,"2022-07-14") ~ "A", 
    str_detect(evacuation_date,"2022-08-08") | str_detect(evacuation_date,"2022-08-05") ~ "B", 
    str_detect(evacuation_date,"2022-09-09") | str_detect(evacuation_date,"2022-09-12") ~ "C", 
    str_detect(evacuation_date,"2022-11-25") | str_detect(evacuation_date,"2022-11-28") ~ "D", 
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "E", 
    str_detect(evacuation_date,"2023-01-30") | str_detect(evacuation_date,"2023-02-02") ~ "F", 
    TRUE ~ "no")) %>% 
   group_by(plot, gapfill_group) %>% 
    dplyr::summarise(across(everything(), mean), .groups = "keep") %>% 
    ungroup()  
 
sd_doc_l1_sw_to_gapfill <- doc_l1_pw_grids %>% 
  select(-collection_datetime, -grid) %>% 
  filter(plot == "Saltwater") %>% 
   filter(evacuation_date %in% c("2022-07-18", "2022-07-14", "2022-08-08", "2022-08-05","2022-09-09", "2022-09-12","2022-11-28", "2022-11-25", "2023-12-05", "2023-12-08", "2023-01-30", "2023-02-02")) %>% 
  mutate(gapfill_group = case_when(  
    str_detect(evacuation_date,"2022-07-18") | str_detect(evacuation_date,"2022-07-14") ~ "A", 
    str_detect(evacuation_date,"2022-08-08") | str_detect(evacuation_date,"2022-08-05") ~ "B", 
    str_detect(evacuation_date,"2022-09-09") | str_detect(evacuation_date,"2022-09-12") ~ "C", 
    str_detect(evacuation_date,"2022-11-25") | str_detect(evacuation_date,"2022-11-28") ~ "D", 
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "E", 
    str_detect(evacuation_date,"2023-01-30") | str_detect(evacuation_date,"2023-02-02") ~ "F", 
    TRUE ~ "no")) %>% 
   group_by(plot, gapfill_group) %>% 
    select(-evacuation_date) %>% 
    dplyr::summarise(across(everything(), sd), .groups = "keep") %>% 
  rename(doc_mg_l_sd = doc_mg_l, 
         tdn_mg_l_sd = tdn_mg_l) %>% 
    ungroup() 
 
sw_gapfill <- mean_doc_l1_sw_to_gapfill %>% 
  full_join(sd_doc_l1_sw_to_gapfill, by = c("plot", "gapfill_group")) %>% 
  select(-gapfill_group) 
 
#Freshwater:  
mean_doc_l1_fw_to_gapfill <- doc_l1_pw_grids %>% 
    select(-grid) %>% 
  filter(plot == "Freshwater") %>% 
  filter(evacuation_date %in% c("2022-09-09", "2022-09-12","2022-11-28", "2022-11-25", "2023-01-30", "2023-02-02", "2023-07-05", "2023-07-07", "2023-09-29", "2023-10-02","2023-12-05", "2023-12-08")) %>% 
  mutate(gapfill_group = case_when(  
    str_detect(evacuation_date,"2022-09-09") | str_detect(evacuation_date,"2022-09-12") ~ "A", 
    str_detect(evacuation_date,"2022-11-28") | str_detect(evacuation_date,"2022-11-25") ~ "B", 
    str_detect(evacuation_date,"2023-01-30") | str_detect(evacuation_date,"2023-02-02") ~ "C", 
    str_detect(evacuation_date,"2023-07-05") | str_detect(evacuation_date,"2023-07-07") ~ "D", 
    str_detect(evacuation_date,"2023-09-29") | str_detect(evacuation_date,"2023-10-02") ~ "E", 
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "F", 
    TRUE ~ "no")) %>% 
   group_by(plot, gapfill_group) %>% 
    dplyr::summarise(across(everything(), mean), .groups = "keep") %>% 
    ungroup() 
 
sd_doc_l1_fw_to_gapfill <- doc_l1_pw_grids %>% 
  select(-collection_datetime, -grid) %>% 
  filter(plot == "Freshwater") %>% 
  filter(evacuation_date %in% c("2022-09-09", "2022-09-12","2022-11-28", "2022-11-25", "2023-01-30", "2023-02-02", "2023-07-05", "2023-07-07", "2023-09-29", "2023-10-02","2023-12-05", "2023-12-08")) %>% 
  mutate(gapfill_group = case_when(  
    str_detect(evacuation_date,"2022-09-09") | str_detect(evacuation_date,"2022-09-12") ~ "A", 
    str_detect(evacuation_date,"2022-11-28") | str_detect(evacuation_date,"2022-11-25") ~ "B", 
    str_detect(evacuation_date,"2023-01-30") | str_detect(evacuation_date,"2023-02-02") ~ "C", 
    str_detect(evacuation_date,"2023-07-05") | str_detect(evacuation_date,"2023-07-07") ~ "D", 
    str_detect(evacuation_date,"2023-09-29") | str_detect(evacuation_date,"2023-10-02") ~ "E", 
    str_detect(evacuation_date,"2023-12-05") | str_detect(evacuation_date,"2023-12-08") ~ "F", 
    TRUE ~ "no")) %>% 
  group_by(plot, gapfill_group) %>% 
  dplyr::select(-evacuation_date) %>% 
  dplyr::summarise(across(everything(), sd), .groups = "keep") %>%
  rename(doc_mg_l_sd = doc_mg_l, 
         tdn_mg_l_sd = tdn_mg_l) %>% 
    ungroup()  
 
fw_gapfill <- mean_doc_l1_fw_to_gapfill %>% 
  full_join(sd_doc_l1_fw_to_gapfill, by = c("plot", "gapfill_group")) %>% 
  select(-gapfill_group) 
 
 
#All the rest:  
means_doc_l1_pw_grids <- doc_l1_pw_grids %>% 
 select(-grid) %>% 
filter(plot != "Control" |  
 (plot == "Control" & !evacuation_date %in% as.Date(c("2022-09-09", "2022-09-12", "2022-11-28", "2022-11-25", "2023-01-30", "2023-02-02", "2023-07-05", "2023-07-07", "2023-12-05", "2023-12-08")))) %>% 
 filter(plot != "Saltwater" | 
(plot == "Saltwater" & !evacuation_date %in% as.Date(c("2022-07-18", "2022-07-14", "2022-08-08", "2022-08-05","2022-09-09", "2022-09-12","2022-11-28", "2022-11-25", "2023-12-05", "2023-12-08", "2023-01-30", "2023-02-02")))) %>% 
filter(plot != "Freshwater" | 
  (plot == "Freshwater" & !evacuation_date %in% as.Date(c("2022-09-09", "2022-09-12","2022-11-28", "2022-11-25", "2023-01-30", "2023-02-02", "2023-07-05", "2023-07-07", "2023-09-29", "2023-10-02","2023-12-05", "2023-12-08")))) %>% 
  mutate(evacuation_date = as.numeric(evacuation_date), 
     collection_datetime= as.numeric(collection_datetime)) %>% 
group_by(plot, evacuation_date) %>% 
dplyr::summarise(across(everything(), mean), .groups = "keep") %>%
mutate(collection_datetime = as.POSIXct(collection_datetime, origin = "1970-01-01"), 
 evacuation_date = as_date(evacuation_date))

sds_doc_l1_pw_grids <- doc_l1_pw_grids %>% 
  select(-collection_datetime, -grid) %>% 
  filter(plot != "Control" |  
    (plot == "Control" & !evacuation_date %in% as.Date(c("2022-09-09", "2022-09-12", "2022-11-28", "2022-11-25", "2023-01-30", "2023-02-02", "2023-07-05", "2023-07-07", "2023-12-05", "2023-12-08")))) %>% 
  filter(plot != "Saltwater" | 
   (plot == "Saltwater" & !evacuation_date %in% as.Date(c("2022-07-18", "2022-07-14", "2022-08-08", "2022-08-05","2022-09-09", "2022-09-12","2022-11-28", "2022-11-25", "2023-12-05", "2023-12-08", "2023-01-30", "2023-02-02"))))  %>% 
  filter(plot != "Freshwater" |  
    (plot == "Freshwater" & !evacuation_date %in% as.Date(c("2022-09-09", "2022-09-12","2022-11-28", "2022-11-25", "2023-01-30", "2023-02-02", "2023-07-05", "2023-07-07", "2023-09-29", "2023-10-02","2023-12-05", "2023-12-08")))) %>% 
  mutate(evacuation_date = as.numeric(evacuation_date)) %>% 
  group_by(plot, evacuation_date) %>% 
  dplyr::summarise(across(everything(), sd), .groups = "keep") %>% 
  mutate(evacuation_date = as_date(evacuation_date)) %>% 
  rename(doc_mg_l_sd = doc_mg_l, 
         tdn_mg_l_sd = tdn_mg_l) 


merging_dfs <- list(means_doc_l1_pw_grids, sds_doc_l1_pw_grids, c_gapfill, sw_gapfill, fw_gapfill)

pw_doc <- purrr::reduce(merging_dfs, full_join)
  
#write out for plotting: 
saveRDS(pw_doc, "~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_plotlevel.rds")
```
  
### run stats for DOC: 

First source the functions from TSLA:

```{r tsla functions}

# set event window to use for all data frame function calls
flood_event_date <- c(as_date("2022-06-22"))

# set event window to use for all data frame function calls
flood_event_datetime <- c(as_datetime("2022-06-22 5:30:00", tz = "EST"), as_datetime("2022-06-22 14:30:00", tz = "EST"))

EVENT_START <- as_datetime("2022-06-22 05:30:00", tz = "EST")
EVENT_STOP <- as_datetime("2022-06-22 14:30:00", tz = "EST")

STUDY_START <- as_datetime("2022-05-01 00:00:00", tz = "EST")

STUDY_END <- as_datetime("2023-05-31 00:00:00", tz = "EST")

source("~/GitHub/tempest-system-level-analysis/scripts/tmp_test_functions.R")

# check them out: 
calc_dist_metrics

```

```{r baci functions}
source("~/GitHub/TEMPEST-1-porewater/scripts/analysis_scripts/BACI_stats_function_fromDivideData.R")

```

```{r doc baci stats}
doc_baci_sw_2022 <- doc_l1_pw_grids %>%
    filter(collection_datetime < endstudydate & collection_datetime > startstudydate) %>%
    divide_data(collection_datetime,flood_event_datetime)
#  baci_stats(var = doc_mg_l, plot = "Saltwater",  p_value_threshold = 0.05)

```


[insert some version of BACI stats here]

## OM Chemistry

### CDOM 
```{r read in cdom}

#Porewater CDOM data, from L1 google drive: 

cdom_l1_2022 <- read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_GRIDS_POOL_Indicies_CDOM_2022_L1.csv")
  
cdom_l1_2023 <- read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_GRIDS_POOL_Indicies_CDOM_2023_L1.csv")
```
```{r merging cdom}
cdom_l1_pw_pool <- cdom_l1_2022 %>%
  full_join(cdom_l1_2023) %>%
  dplyr::filter(grid == "POOL") %>%
  select(plot, grid, evacuation_date, collection_datetime, S275_295, S350_400, Sr:M_RSU) 
```
```{r bring in pooled DOC data for merging}
doc_l1_pooled_2022 <- read_csv("~/GitHub/TEMPEST_Porewater/processing_scripts/DOC/TEMPORARY/TMP_PW_POOLED_NPOCTDN_2022_L1.csv")

doc_l1_pooled_2023 <- read_csv("~/GitHub/TEMPEST_Porewater/processing_scripts/DOC/TEMPORARY/TMP_PW_POOLED_NPOCTDN_2023_L1.csv")

doc_l1_pooled_all <- doc_l1_pooled_2022  %>%
  full_join(doc_l1_pooled_2023)

# merge cdom data with the pooled DOC data to calculate SUVA
cdom_wdoc_pooled <- cdom_l1_pw_pool %>%
  mutate(date = as_date(collection_datetime)) %>%
  left_join(doc_l1_pooled_all, by = c("plot", "grid", "date", "evacuation_date")) %>%
  mutate(suva_254 = Abs_254nm / doc_mg_l *100 ) %>%
  #remove instances where pairs are unmatched returning NAs
  drop_na()

cdom_nodoc <- cdom_l1_pw_pool %>%
  mutate(date = as_date(collection_datetime)) %>%
  anti_join(doc_l1_pooled_all, by = c("plot", "grid", "date"))

# datetimes for the rest are off by a day or so from the averaged DOC data; I am not really sure why this is. going to match them manually since only 17 samples we are dealing with- that can be consolidated into 11 unique instances by evacuation date groups: 
cdom_nodoc_1 <- cdom_nodoc %>%
    select(-grid) %>%
    mutate(gapfill_group = case_when( 
    str_detect(evacuation_date,"2022-06-13") | str_detect(evacuation_date,"2022-06-15") ~ "A",
    str_detect(evacuation_date,"2022-07-14") | str_detect(evacuation_date,"2022-07-18") ~ "B",
    str_detect(evacuation_date,"2022-09-09") ~ "C",
    str_detect(evacuation_date,"2022-06-18") | str_detect(evacuation_date,"2022-06-22") ~ "D",
    str_detect(evacuation_date,"2022-07-18") ~ "E",
    TRUE ~ "no")) %>%
   group_by(plot, gapfill_group) %>%
   dplyr::select(-evacuation_date, -collection_datetime, -date) %>% # these get averaged if not. Will adopt the dates in the porewater DOC average sheet.  
   dplyr::summarise(across(everything(), mean), .groups = "keep")

#for these samples, use the plot averages at similar dates to calculate SUVA:
pw_doc_1 <- pw_doc %>%
  mutate(gapfill_group = case_when( 
    str_detect(evacuation_date,"2022-06-09") ~ "A", #these all have collection dates of 6-14 in the avg doc sheet. 
    str_detect(evacuation_date,"2022-07-14") ~ "B", #these all have collection dates of 7-18 in the avg doc sheet. 
    str_detect(evacuation_date,"2022-09-10") | str_detect(evacuation_date,"2022-09-11") ~ "C", #these are closest to the 9/9 date in the cdom sheet. 
    str_detect(evacuation_date,"2022-06-18") | str_detect(evacuation_date,"2022-06-22") ~ "D",
    str_detect(evacuation_date,"2022-07-18") ~ "E",
    TRUE ~ "no")) %>%
   dplyr::filter(gapfill_group != "no") %>%
   group_by(plot, gapfill_group) %>% #ok if dates/times get averaged here. 
   dplyr::summarise(across(everything(), mean), .groups = "keep") %>%
   dplyr::select(plot, gapfill_group, evacuation_date, collection_datetime, doc_mg_l)

cdom_plot_avgsuva <- cdom_nodoc_1 %>%
  left_join(pw_doc_1, by= c("plot",  "gapfill_group")) %>%
  mutate(suva_254 = Abs_254nm / doc_mg_l *100 )

cdom_all <- cdom_wdoc_pooled %>%
  full_join(cdom_plot_avgsuva) %>%
  dplyr::select(plot, evacuation_date, collection_datetime, doc_mg_l, suva_254, S275_295:M_RSU) %>%
  mutate(date = as_date(collection_datetime))


#write out for plotting:

saveRDS(cdom_all, "~/GitHub/TEMPEST-1-porewater/data/averaged_CDOM_POOL_porewater_figures.rds")
```

```{r cdom tsla metrics}
# tsla metrics for YR 1 data: 
cdom_long <- cdom_all %>%
    filter(between(date, startstudydate_yr1, endstudydate_yr1)) %>%
  dplyr::select(-doc_mg_l) %>%
  pivot_longer(cols = -c(plot, evacuation_date, collection_datetime, date),
               names_to = "metric",
               values_to = "values")

#assign timepoints and write out for various plotting efforts
timepoints_cdom <- divide_data(cdom_long, date, flood_event_date)   


sum_cdom <- timepoints_cdom %>%
  group_by(plot, metric, timedate_bin) %>%
  summarize(value_mean = mean(values), value_sd = sd(values), count = n()) 

# write out for plotting:
saveRDS(sum_cdom, "~/GitHub/TEMPEST-1-porewater/data/avg_metrics_premidpost_cdom_porewater.rds")

```

dates_cdom_avg <- cdom_all%>%
  dplyr::select(-evacuation_date, -collection_datetime) %>%
  group_by(plot, date, timedate_bin) %>%
  filter(between(date, startstudydate_yr1, endstudydate)) %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(date) - 5) %% 12 ) %>%
  # fix the sampling dates in June before the event to the month before
  mutate(adj_month = case_when(adj_month == 1 & date < EventStart2022 ~ adj_month == 0,
                               TRUE ~ adj_month)) %>%
  group_by(adj_month, year, plot) %>%
  
    summarise( across( 
      .cols = everything(),
      .fns = list(mean = ~ mean(.), sd = ~ sd(.)),
      .names = "{.col}_{.fn}"),
    .groups = "keep")

# write out for plotting:
saveRDS(dates_cdom_avg, "~/GitHub/TEMPEST-1-porewater/data/avg_relabund_dates_cdom_porewater.rds")


#### FTICRMS 
```{r load fticrms}
# load data
pw_ft <- read_rds("~/GitHub/TEMPEST-1-porewater/processed data/TMP_PW_POOL_FTICRMS_May2022-May2023_L1.rds") %>%
  mutate(datetime_est = force_tz(collection_datetime, tzone = "EST")) %>% 
   group_by(Event, plot, grid, date, sample_name, Class_detailed) %>%
    dplyr::summarise(abund = sum(presence))  %>%
    ungroup() %>% 
    # create a new column for total counts per core assignment
    # and then calculate relative abundance  
    group_by(Event, plot, grid, date, sample_name) %>% 
    dplyr::mutate(total = sum(abund),
                  relabund  = round((abund/total)*100,2)) %>%
  ungroup() %>%
  dplyr::select(date, plot, Class_detailed, relabund) %>% 
  group_by(date, plot, Class_detailed) %>%
  zscore_standard(vars = "relabund") %>% ungroup()

pw_ft %>% summary()

```

```{r tsla metrics}
#assign timepoints and write out for various plotting efforts
timepoints_fticrms <- divide_data(pw_ft, date, flood_event_date)   

dates_fticrms_avg <- timepoints_fticrms %>%
  group_by(plot, Class_detailed, date, timedate_bin) %>%
  summarize(relabund_mean = mean(relabund), relabund_sd = sd(relabund)) %>%
  mutate(plot = case_when(plot == "FW" ~ "Freshwater",
                          plot == "SW" ~ "Saltwater",
                          plot == "C" ~ "Control",
                          TRUE ~ plot))

# write out for plotting:
saveRDS(dates_fticrms_avg, "~/GitHub/TEMPEST-1-porewater/data/avg_relabund_dates_fticrms_porewater.rds")


sum_fticrms <- timepoints_fticrms%>%
  group_by(plot, Class_detailed, timedate_bin) %>%
  summarize(relabund_mean = mean(relabund), relabund_sd = sd(relabund)) %>%
  mutate(plot = case_when(plot == "FW" ~ "Freshwater",
                          plot == "SW" ~ "Saltwater",
                          plot == "C" ~ "Control",
                          TRUE ~ plot))

# write out for plotting:
saveRDS(sum_fticrms, "~/GitHub/TEMPEST-1-porewater/data/avg_relativeabundance_premidpost_fticrms_porewater.rds")
```

# OM chemistry stats

Stats for aromatic content in FTICRMS signal:
Trying a simple repeated measures anova first: 
```{r anova aromatic fticrms}

dates_fticrms_avg_stats <- timepoints_fticrms %>%
  mutate(plot = case_when(plot == "FW" ~ "Freshwater",
                          plot == "SW" ~ "Saltwater",
                          plot == "C" ~ "Control",
                          TRUE ~ plot)) %>%
  mutate(plot = factor(plot, levels = plot_order),
         timedate_bin = factor(timedate_bin, levels = time_order)) %>%
filter(Class_detailed == "highly aromatic")

lmm_res <- lmer(relabund ~ plot + (1|timedate_bin), data = dates_fticrms_avg_stats )
summary(lmm_res)

#baci_stats(dates_fticrms_avg_stats, var = relabund, plot = "Saltwater",  p_value_threshold = 0.05)
  
```

