---
title: "Porewater DOC Manuscript Analysis"
author: "AMP"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(ggpubr)
library(viridis)
library(scales)
library(tidyverse)
library(ggnewscale)

#clean out your environment 
rm(list = ls())


#double check your wd
getwd()


```
### Set the study dates

```{r event and study dates}
endstudydate = lubridate::as_date("2025-05-31")
startstudydate = lubridate::as_date("2022-05-01")

x_limits <- as.Date(c(startstudydate, endstudydate))

plotting_end = as.POSIXct("2025-05-31 11:59:59", tz = "EST")
plotting_start = as.POSIXct("2022-05-01 00:00:00", tz = "EST")

endstudydate_yr1 = lubridate::as_date("2023-05-31")
startstudydate_yr1 = lubridate::as_date("2022-05-01")

endplotdate_yr1 = lubridate::as_date("2023-04-30")

year1_start = lubridate::as_date("2022-01-01")
year1_stop = lubridate::as_date("2022-12-31")

year2_start = lubridate::as_date("2023-01-01")
year2_stop = lubridate::as_date("2023-12-31")

year3_start = lubridate::as_date("2024-01-01")
year3_stop = lubridate::as_date("2024-12-31")

WaterDeliveryStart2022 = as.POSIXct("2022-06-22 05:30:00", tz = "EST")
WaterDeliveryStop2022 = as.POSIXct("2022-06-22 14:30:00", tz = "EST")

WaterDeliveryStart1 = as.POSIXct("2023-06-06 05:30:00", tz = "EST")
WaterDeliveryStop1 = as.POSIXct("2023-06-06 14:30:00", tz = "EST")

WaterDeliveryStart2 = as.POSIXct("2023-06-07 05:30:00", tz = "EST")
WaterDeliveryStop2 = as.POSIXct("2023-06-07 14:30:00", tz = "EST")

WaterDeliveryStart1_2024 = as.POSIXct("2024-06-11 06:10:00", tz = "EST")
WaterDeliveryStart2_2024 = as.POSIXct("2024-06-12 06:08:00", tz = "EST")
WaterDeliveryStart3_2024 = as.POSIXct("2024-06-13 06:21:00", tz = "EST")

flood_event_datetime <- c(as_date("2022-06-22"))
```

```{r theme set up}

Anyas_theme = 
        theme_set(theme_bw() +
        theme(panel.background = element_blank(),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(face = "bold", size = 14),
        legend.position = "bottom",
        legend.title = element_blank(),
        strip.text = element_text(size=16)))

plot_order <- c('Control', 'Freshwater','Saltwater')
time_order <- c('Pre', 'Mid','Post')

Anyas_colors = c("springgreen2", "cyan2", "violetred2")

Anyas_colors_figure1 =  c( "cyan2","violetred2")

Treatment_order <- c('Control', 'Artificial Freshwater','Artificial Saltwater')

base_text_size <- 11
axis_text_size <- 10
axis_title_size <- 11
legend_text_size <- 10
panel_label_size <- 11

event_dates <- sort(as.Date(c(
  WaterDeliveryStart2022,
  WaterDeliveryStart1,
  WaterDeliveryStart1_2024
)))

figure_x_limits <- c(
  as.Date(format(startstudydate, "%Y-%m-01")),
  as.Date(endstudydate)
)

figure_x_breaks <- seq(
  from = figure_x_limits[1],
  to = figure_x_limits[2],
  by = "1 month"
)

figure_x_labels <- function(x) {
  month_num <- format(x, "%m")
  month_label <- format(x, "%b")

  ifelse(
    month_num %in% c("01", "06"),
    paste0(month_label, "\n", format(x, "%Y")),
    ""
  )
}

plot_shapes <- c(
  "Freshwater" = 16,
  "Saltwater" = 17
)

event_labels <- data.frame(
  x = event_dates,
  label = c("T1", "T2", "T3")
)


```

# Figure 1: 

data: 
```{r load doc data}

doc_all <- read_rds("~/GitHub/TEMPEST-1-porewater/data/DOC_Figure1_FINAL.rds")

doc_sw <- doc_all %>%
  filter(plot == "Saltwater") %>%
   mutate(nominal_datetime = as.POSIXct(nominal_date)) %>%
  filter(between(nominal_date, startstudydate, endstudydate))

doc_effect_size <- read_rds("~/GitHub/TEMPEST-1-porewater/data/DOC_Figure1_effectsizes.rds") %>%
   mutate(nominal_datetime = as.POSIXct(nominal_date))

```

```{r bring in SPC}

ec_to_doc_scale <- 100 / 1735.222 # doc max range vs ec max
#ec_scaled_offset <- doc_range[1] - ec_range[1] * ec_to_doc_scale

pw_ec_plotting <- read_rds("~/GitHub/TEMPEST-1-porewater/data/averaged_by_plot_ec_teros_porewater.rds") %>%
  mutate(EC_scaled_to_doc_axis = EC_uScm * ec_to_doc_scale) %>% #scale EC to plot 
  filter(plot =="Saltwater") %>%
  filter(between(datetime_est, startstudydate, endstudydate))

 pw_ec_all <- read_rds("~/GitHub/TEMPEST-1-porewater/processed data/TMP_PW_EC_15cm_2021-2025_L2.rds")

pw_doc_dates <- read_rds("~/GitHub/TEMPEST-1-porewater/data/dates_DOC_higher_than_variance.rds")

pw_doc_dates
```

```{r set dates based on pw_doc_dates}
tmp1_doc_high = as.POSIXct("2022-07-21 15:59:00", tz = "EST")
tmp2_doc_high = as.POSIXct("2023-08-10 15:53:00", tz = "EST")
tmp3_doc_high = as.POSIXct("2025-01-31 20:52:00", tz = "EST")

```

## a) DOC effect sizes 
Supporting Main Message: This phenomenon can be replicated in the laboratory; AND we can repeat it in the field the second year “reset” (SI Figure X)

```{r fig 1a plot}

Figure1a <- doc_effect_size %>%
  filter(between(nominal_date, startstudydate, endstudydate)) %>%
  select(
    treatment_plot,
    nominal_date,
    effect_size_boot_mean,
    effect_size_ci_lower,
    effect_size_ci_upper,
    effect_size_percent_boot_mean,
    effect_size_percent_ci_lower,
    effect_size_percent_ci_upper
  ) %>%
  mutate(
    nominal_date = as.Date(nominal_date),
    plot = factor(treatment_plot, levels = plot_order),
    event_period = cut(
      nominal_date,
      breaks = c(as.Date("1900-01-01"), event_dates, as.Date("2100-01-01")),
      labels = c(
        "before_2022",
        "after_2022_before_2023",
        "after_2023_before_2024",
        "after_2024"
      ),
      right = FALSE),
  line_group = interaction(plot, event_period, drop = TRUE)
  ) %>%
  ggplot(aes(
    x = nominal_date,
    y = effect_size_boot_mean,
    color = plot,
    fill = plot,
    shape = plot
  )) +
  geom_vline(xintercept = WaterDeliveryStart2022, linetype = "dashed", color = "black") +
  geom_vline(xintercept = WaterDeliveryStart1, linetype = "dashed", color = "black") +
  geom_vline(xintercept = WaterDeliveryStart1_2024, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 0, linetype = "solid", color = "darkgrey") +
  geom_ribbon(
    aes(
      ymin = effect_size_ci_lower,
      ymax = effect_size_ci_upper,
      group = line_group
    ),
    alpha = 0.18,
    color = NA
  ) +
  geom_line(
    aes(group = line_group),
    linewidth = 0.6,
    alpha = 0.30
  ) +
  geom_point(
    size = 1.8
  ) +
  labs(
    x = NULL,
    y = "DOC effect size \n (treatment-control) / control", #  \n (% difference from Control Plot)
    color = NULL,
    fill = NULL,
    shape = NULL
  ) +
  scale_color_manual(
  values = Anyas_colors_figure1,
  breaks = names(plot_shapes)
) +
scale_fill_manual(
  values = Anyas_colors_figure1,
  guide = "none"
) +
scale_shape_manual(
  values = plot_shapes,
  breaks = names(plot_shapes)
) +
guides(
  color = guide_legend(nrow = 1),
  shape = guide_legend(nrow = 1),
  fill = "none"
) +
scale_x_date(
  limits = figure_x_limits,
  breaks = figure_x_breaks,
  labels = figure_x_labels,
  expand = expansion(mult = c(0, 0))
) +
  theme_classic() +
  geom_text(
    data = data.frame(x = -Inf, y = Inf, label = "  a) "),
    aes(x, y, label = label),
    hjust = -0.1,
    vjust = 2,
    inherit.aes = FALSE
  )+
  geom_label(
  data = event_labels,
  aes(x = event_dates, y = Inf, label = c("T1","T2","T3")),
  inherit.aes = FALSE,
  vjust = 3,
  label.size = 0.25,
  label.padding = unit(0.12, "lines"),
  fill = "white",
  color = "black",
  fontface = "bold",
  size = 4
) +
    theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_text(size = axis_title_size),
    axis.title.x = element_blank(),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_blank(),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )
    
    
print(Figure1a)

```

## b) Cond and DOC in SW plot only
```{r fig 1b plot}

library(ggnewscale)

ec_axis_col <- "#1F4E79"

doc_event_dates <- sort(as.Date(c(
  tmp1_doc_high,
  tmp2_doc_high,
  tmp3_doc_high
)))

doc_sw_plot <- doc_sw %>%
  mutate(
    nominal_date = as.Date(nominal_date),
    nominal_datetime = as.POSIXct(nominal_date),
    high_doc_window = case_when(
      nominal_date >= as.Date("2022-07-16") &
        nominal_date < as.Date(WaterDeliveryStart1) ~ "After conductivity threshold",

      nominal_date >= as.Date("2023-08-07") &
        nominal_date < as.Date(WaterDeliveryStart1_2024) ~ "After conductivity threshold",

      nominal_date >= as.Date("2025-01-30")  ~ "After conductivity threshold",

      TRUE ~ "Before conductivity threshold"
    ))

fig1b <-
  ggplot() +
  geom_point(
    data = pw_ec_plotting,
    aes(x = datetime_est, y = EC_scaled_to_doc_axis, color = plot),
    alpha = 0.45,
    size = 1
  ) +
  scale_color_manual(values = c("Saltwater" = ec_axis_col)) +

  geom_text(
    data = data.frame(x = -Inf, y = Inf, label = "  b) "),
    aes(x, y, label = label),
    hjust = -0.1,
    vjust = 2,
    inherit.aes = FALSE
  ) +
  ggnewscale::new_scale_color() +

  geom_vline(xintercept = WaterDeliveryStart2022, linetype = "dashed", color = "black") +
  geom_vline(xintercept = WaterDeliveryStart1, linetype = "dashed", color = "black") +
  geom_vline(xintercept = WaterDeliveryStart1_2024, linetype = "dashed", color = "black") +
 # geom_vline(xintercept = tmp1_doc_high, linetype = "dashed", color = "black") +
 # geom_vline(xintercept = tmp2_doc_high, linetype = "dashed", color = "black") +
 # geom_vline(xintercept = tmp3_doc_high, linetype = "dashed", color = "black") +

geom_pointrange(
  data = doc_sw_plot,
  aes(
    x = nominal_datetime,
    y = doc_mg_l_mean,
    ymin = doc_mg_l_mean - doc_mg_l_sd,
    ymax = doc_mg_l_mean + doc_mg_l_sd,
    fill = high_doc_window
  ),
  shape = 24,
  size = 1,
    stroke = 0.7,
  linewidth = 0.6,
  linetype = "solid"
) +
scale_fill_manual(
  values = c(
    "Before conductivity threshold" = "white",
    "After conductivity threshold" = "#B6854D"
  )
) +
  xlab("Date") +
scale_x_date(
  limits = figure_x_limits,
  breaks = figure_x_breaks,
  labels = figure_x_labels,
  expand = expansion(mult = c(0, 0))
)+
  scale_y_continuous(
    name = "Saltwater plot DOC at 15 cm (mg/L)",
    sec.axis = sec_axis(
      ~ (. - 1) / ec_to_doc_scale,
      name = "Saltwater plot mean soil \n conductivity at 15 cm (µS/cm)"
    )
  ) +
  theme_classic() +
theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_text(size = axis_title_size),
    axis.title.x = element_blank(),
     axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(2, 5.5, 5.5, 5.5),
    axis.title.y.right = element_text(color = ec_axis_col),
axis.text.y.right = element_text(color = ec_axis_col),
axis.ticks.y.right = element_line(color = ec_axis_col),
axis.line.y.right = element_line(color = ec_axis_col)
  )


fig1b


```

## Combined Figure 1:    
```{r figure 1}
panel_a <- Figure1a +
  theme(legend.position = "none")

panel_b <- fig1b +
  theme(legend.position = "none")

plotting = c("Before conductivity threshold", "After conductivity threshold", "Saltwater")


legend_a_plot <- tibble(
  x = c(0.42, 0.58),
  label = c("Freshwater plot", "Saltwater plot"),
  plot = factor(c("Freshwater", "Saltwater"), levels = plot_order)
) %>%
  ggplot(aes(x = x, y = 1, color = plot, shape = plot)) +
  geom_segment(aes(x = x - 0.035, xend = x + 0.02, yend = 1), linewidth = 0.7, alpha = 0.45) +
  geom_point(size = 3) +
  geom_text(aes(x = x + 0.045, label = label), color = "black", hjust = 0, size = legend_text_size / ggplot2::.pt) +
  scale_color_manual(values = Anyas_colors_figure1, breaks = names(plot_shapes))+
  coord_cartesian(xlim = c(0.28, 0.78), ylim = c(0.8, 1.2), clip = "off") +
  theme_void() +
  theme(legend.position = "none")


legend_b_plot <- tibble(
  plot = factor(plotting, levels = plotting),
  x = c(0.16, 0.46, 0.76),
  label = c(
    "DOC before conductivity threshold",
    "DOC after conductivity threshold",
    "Soil conductivity"
  ),
  color_value = c(
    "black",
    "#B6854D",
    "#1F4E79"
  ),
  point_shape = c(24, 17, 16),
  point_size = c(4.0, 4.0, 1.8)
) %>%
  ggplot(aes(x = x, y = 1)) +
  geom_point(aes(color = color_value, shape = factor(point_shape), size = point_size), show.legend = FALSE) +
  geom_text(aes(x = x + 0.035, label = label), color = "black", hjust = 0, size = legend_text_size / ggplot2::.pt) +
  scale_color_identity() +
  scale_shape_manual(values = c("24" = 24, "16" = 16, "17" = 17)) +
  scale_size_identity() +
  coord_cartesian(xlim = c(0.1, 1.02), ylim = c(0.8, 1.2), clip = "off") +
  theme_void() +
  theme(legend.position = "none")

aligned_panels <- cowplot::align_plots(
  panel_a,
  panel_b,
  align = "v",
  axis = "lr"
)

Figure1 <- cowplot::plot_grid(
  aligned_panels[[1]],
  legend_a_plot,
  aligned_panels[[2]],
  legend_b_plot,
  ncol = 1,
  rel_heights = c(1, 0.16, 1, 0.20)
)

Figure1

cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/Figure1_updated.png",Figure1, base_width = 8.5,
  base_height = 7, dpi=600)

```

# Extended Data Figures


# Extended Data Figure 1: Precip and SPC all plots

These factors in isolation or combination may explain large interannual variation in the rate that soil conductivity declined (Extended Data Figure 1), and therefore the timing of the abrupt shift in carbon partitioning between solid and dissolved forms.


```{r precip data}

precip_data <- read_csv("~/GitHub/TEMPEST-1-porewater/data/daily_precip_2021-2026.csv")

pw_ec_all <- read_rds("~/GitHub/TEMPEST-1-porewater/data/averaged_by_plot_ec_teros_porewater.rds")

```

```{r SI 1}

spc <- pw_ec_all %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = datetime_est, y = EC_uScm, color = plot)) +
    geom_point()+
  geom_vline(xintercept = WaterDeliveryStart2022, linetype = "dashed", color = "black") +
  geom_vline(xintercept = WaterDeliveryStart1, linetype = "dashed", color = "black") +
  geom_vline(xintercept = WaterDeliveryStart1_2024, linetype = "dashed", color = "black") +
    ylab("mean Porewater \n Conductivity (uS/cm)")+
    xlab("Date")+
    scale_color_manual(values=Anyas_colors) +
  scale_x_date(
  limits = figure_x_limits,
  breaks = figure_x_breaks,
  labels = figure_x_labels,
  expand = expansion(mult = c(0, 0))
)+
  geom_text(
    data = data.frame(x = -Inf, y = Inf, label = "  b) "),
    aes(x, y, label = label),
    hjust = -0.1,
    vjust = 2,
    inherit.aes = FALSE
  )+
    theme_classic() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_text(size = axis_title_size),
    axis.title.x = element_blank(),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_text(size = axis_text_size),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )

spc 

precip <-   precip_data %>%
  ggplot(aes(x = date, y = precip_cm)) +
    geom_col(color = "lightblue")+
   geom_vline(xintercept = WaterDeliveryStart2022, linetype = "dashed", color = "black") +
  geom_vline(xintercept = WaterDeliveryStart1, linetype = "dashed", color = "black") +
  geom_vline(xintercept = WaterDeliveryStart1_2024, linetype = "dashed", color = "black") +
    ylab("Daily Precip (cm)")+
    xlab("Date")+
  scale_x_date(
  limits = figure_x_limits,
  breaks = figure_x_breaks,
  labels = figure_x_labels,
  expand = expansion(mult = c(0, 0))
)+  
  geom_text(
    data = data.frame(x = -Inf, y = Inf, label = "  a) "),
    aes(x, y, label = label),
    hjust = -0.1,
    vjust = 2,
    inherit.aes = FALSE
  )+
  geom_label(
  data = event_labels,
  aes(x = event_dates, y = Inf, label = c("T1","T2","T3")),
  inherit.aes = FALSE,
  vjust = 3,
  label.size = 0.25,
  label.padding = unit(0.12, "lines"),
  fill = "white",
  color = "black",
  fontface = "bold",
  size = 4
) +
    theme_classic() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_text(size = axis_title_size),
    axis.title.x = element_blank(),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_blank(),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )


precip


SI1_aligned_panels <- cowplot::align_plots(
  spc,
  precip,
  align = "v",
  axis = "lr"
)

SI_1 <- cowplot::plot_grid(
  SI1_aligned_panels[[2]],
  SI1_aligned_panels[[1]],
  ncol = 1,
  rel_heights = c(0.5, 1)
)

SI_1 

cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/SI_Figure1.png", SI_1 , base_width = 8.5,
  base_height = 7, dpi=600)

```


# Extended Data Figure 2: Column Experiment DOC

The defining characteristic of the field response — a delayed and persistent DOC increase following salinity exposure — was reproduced in intact soil cores in the absence of living vegetation and ecosystem-scale biological feedbacks. 

```{r load col ex}
# exp_doc <- read_csv("~/GitHub/tempest_ionic_strength/Data/Processed Data/DOC/TEMPEST_ColEx_NPOC_TDN_meansforCDOM.csv") %>%
#   mutate(Wash = case_when(is.na(Wash) ~ 0,
#                           TRUE ~ Wash),
#     Treatment = case_when(Treatment == "AS" ~ "Artificial Saltwater",
#                                Treatment == "FW" ~ "Artificial Freshwater",
#                                Treatment == "Control" ~ "Control",
#                                Treatment == "SWE" ~ "Artificial Saltwater",
#                                Treatment == "CL" ~ "Artificial Freshwater",
#                           TRUE ~ Treatment)) %>%
#   mutate(Treatment = factor(Treatment, levels = Treatment_order))

exp_doc <- read_csv("~/GitHub/tempest_ionic_strength/Data/Exp 3 Column Experiment ColEx Data/TEMPEST_Col_Ex_all_chem_averaged_L2.csv") %>%
  mutate(Wash = case_when(is.na(Timepoint) ~ 0,
                          TRUE ~ Timepoint),
    Treatment = case_when(Treatment == "SW" ~ "Artificial Saltwater",
                               Treatment == "FW" ~ "Artificial Freshwater",
                               Treatment == "Control" ~ "Control",
                               Treatment == "SWE" ~ "Artificial Saltwater",
                               Treatment == "CL" ~ "Artificial Freshwater",
                          TRUE ~ Treatment)) %>%
  mutate(Treatment = factor(Treatment, levels = Treatment_order))
```

```{r col ex}
SI2 <- exp_doc %>%
   ggplot(aes(x = Wash, y = doc_mgperL_avg, color = Treatment)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mgperL_avg-doc_mgperL_sd, ymax=doc_mgperL_avg+doc_mgperL_sd), shape=17, size = 1)+
    ylab("DOC mg/L")+
    xlab("Treatment Wash Number")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(n=6), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 14)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  Column Experiment"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE) +
    theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_text(size = axis_title_size),
    axis.title.x = element_text(size = axis_title_size),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_text(size = axis_text_size),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )

print(SI2)

cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/SI_Figure2.png",SI2, dpi=600)
```
# Extended Data Figure 3: CDOM
Likewise, if DOC increases were related only to increased microbial metabolism, we would expect an increase in non-colored, small molecular weight OM, whereas we observed instead an increase in colored DOC that coincides with the increase in DOC concentrations (Extended Data Figure 3). 

```{r load eems data}

cdom_indicies_divided_all <- readRDS("~/GitHub/TEMPEST-1-porewater/data/cdom_all_forplotting.rds")

eemspca <- readRDS("~/GitHub/TEMPEST-1-porewater/data/cdom_pca.rds")

```

```{r pca}

library(dplyr)
library(ggplot2)
library(ggfortify)
library("wesanderson")

cdom_indicies_divided_all_graphing <- cdom_indicies_divided_all %>%
  mutate(
    color_group = case_when(
      plot == "Saltwater" &
        timedate_bin == "Post Cond Threshold" ~ "Saltwater Post Threshold",
      timedate_bin == "Pre" ~ "Pre Event",
      timedate_bin == "Mid" ~ "Mid Event(s)",
      TRUE ~ "Pre Threshold"
    ),
    color_group = factor(
      color_group,
      levels = c("Pre Event", "Mid Event(s)", "Pre Threshold", "Saltwater Post Threshold")
    )
  )

pal2 <- c(
  "Pre Event"            = "#B8D3E3",  # pale powder blue
  "Mid Event(s)"            = "#7EA9C4",  # muted steel blue
  "Pre Threshold"       = "#3E6F97",  # deep dusty blue
  "Saltwater Post Threshold" = "#D69C4E"      # Darjeeling2 brown
)

SI3 <-
  autoplot(
  eemspca,
  data = cdom_indicies_divided_all_graphing,
  shape = "plot",
  colour = "color_group",
  size = 2,
  loadings = TRUE,
  loadings.colour = "black",
  loadings.label = TRUE,
  loadings.label.size = 2,
    loadings.label.colour = "black",  # text
   loadings.label.repel = TRUE
) +
  scale_shape_manual( values = c("Control" = 0, "Freshwater" = 1, "Saltwater" = 17)  )+
  scale_colour_manual(values = pal2)+
   theme_classic() +
    theme(legend.position = "right", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 14)) +
    theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_text(size = axis_title_size),
    axis.title.x = element_text(size = axis_title_size),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_text(size = axis_text_size),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )

SI3

cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/SI_Figure3.png",SI3, dpi=600)

```


# Extended Data Figure 4:  a) DOC curves with ASW vs NaCl b) NaCl Fe and Al 

```{r nacl}

exp_nacl_dat <- readxl::read_excel("~/GitHub/tempest_ionic_strength/Data/Exp 2 Ionic Strength NaCl Data/TEMPEST_Ionic_Strength_NaCl_chem_L2.xlsx")  %>%
  mutate(Treatment =  as.factor(Treatment))

nacl_doc <- exp_nacl_dat %>%
  mutate(group = interaction(Exp_Type, Treatment, sep = "_")) %>%
  select(group, Wash, doc_mgperL, doc_mgperL_stdev) %>%
  rename(doc_mg_l_mean= doc_mgperL,
         doc_mg_l_sd = doc_mgperL_stdev)

exp_asw_dat <- read_csv("~/GitHub/tempest_ionic_strength/Data/Exp 1 Ionic Strength ASW Data/TEMPEST_Ionic_Strength_ASW_chem_L2.csv")

asw_doc <- exp_asw_dat %>%
  mutate(group = interaction(Exp_Type, Treatment, sep = "_")) %>%
  select(group, Wash, doc_mg_l_mean, doc_mg_l_sd)

batch_exps <- nacl_doc %>% full_join(asw_doc) %>%
   separate(group, into = c("Experiment", "Level"), sep = "_")

batch_exps_change <- batch_exps %>%
  group_by(Experiment, Level) %>%
  mutate(
    doc_change = doc_mg_l_mean -
      doc_mg_l_mean[Wash == 1]
  ) %>%
  ungroup() %>%
    mutate(
    Level = factor(
      Level,
      levels = c("0", "0.1", "1", "5", "25", "100")
    )
  )
```
```{r batch plots}
SI4a <- batch_exps_change %>%
  ggplot(aes(
    x = Wash,
    y = doc_change,
    color = Level,
    linetype = Experiment
  )) +
  scale_color_manual(values = PNWColors::pnw_palette("Starfish",6))+
  geom_line(linewidth = 0.7) +
  geom_point(size = 2)+
  geom_hline(yintercept = 0, color = "grey60", linewidth = 0.5)+
#  geom_pointrange(
 #   aes(
 #     ymin = doc_mg_l_mean - doc_mg_l_sd,
# #     ymax = doc_mg_l_mean + doc_mg_l_sd
 #   ),
#    size = 0.5
#  ) +
#  scale_shape_manual(values = c(
#    "NaCl" = 16,   # circle
#    "ASW"  = 15    # square
#  )) +
  geom_text(
    data = data.frame(x = -Inf, y = Inf, label = "  a) "),
    aes(x, y, label = label),
    hjust = -0.1,
    vjust = 2,
    inherit.aes = FALSE
  )+
labs(
  y = expression(Delta*"DOC relative to treatment mgC/L"),
  x = "Treatement Wash Number",
  color = "Treatment Type (% relative \n to full strength ASW)"
)+
  scale_x_continuous(
    breaks = pretty_breaks(n = 6),
    limits = c(1, 6)
  ) +
  theme_classic() +
    theme(
    legend.position = "bottom",
    legend.title = element_text(size = legend_text_size),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_text(size = axis_title_size),
    axis.title.x = element_text(size = axis_title_size),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_text(size = axis_text_size),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )

print(SI4a)

```

For example, b) in the batch experiments with NaCl as the only introduced ions, higher ionic strength results in delayed release of greater amounts of Al and Fe from the soils. 

Fe, Al
```{r icp nacl}
metals_long <- exp_nacl_dat %>%
  rename(
    Al_mean = Al_ppm,
    Al_sd   = Al_ppm_stdev,
    Fe_mean = Fe_ppm,
    Fe_sd   = Fe_ppm_stdev
  ) %>%
  pivot_longer(
    cols = c(Al_mean, Al_sd, Fe_mean, Fe_sd),
    names_to = c("Metal", ".value"),
    names_pattern = "(Al|Fe)_(mean|sd)"
  ) %>%
  mutate(
    Treatment = factor(
      Treatment,
      levels = c("0", "0.1", "1", "5", "25", "100")
    )
  )

metal_com <- metals_long %>%
  group_by(Metal, Treatment) %>%
  summarise(
    center_of_mass = sum(Wash * mean, na.rm = TRUE) /
                     sum(mean, na.rm = TRUE),
    .groups = "drop"
  )

```

```{r summary plot metals nacl}
SI4b <- ggplot(
  metal_com,
  aes(
    x = Treatment,
    y = center_of_mass,
    color = Metal,
    shape = Metal,
    group = Metal
  )
) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 3) +
  scale_shape_manual(values = c(16, 15)) +
  scale_y_continuous(
    limits = c(1, 6),
    breaks = 1:6
  ) +
  labs(
    x = "NaCl Treatment \n (% relative to full strength ASW)",
    y = "Mean timing of Fe or Al \n release (wash number)",
    color = NULL,
    shape = NULL
  ) +  geom_text(
    data = data.frame(x = -Inf, y = Inf, label = "  b) "),
    aes(x, y, label = label),
    hjust = -0.1,
    vjust = 2,
    inherit.aes = FALSE
  )+
scale_color_manual(
  values = c(
    "Al" = "#4F6D8A",
    "Fe" = "#8C4A2F"
  )
)+
   theme_classic() +
    theme(
    legend.position = "bottom",
    legend.title = element_text(size = legend_text_size),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_text(size = axis_title_size),
    axis.title.x = element_text(size = axis_title_size),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_text(size = axis_text_size),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )

SI4b

```

```{r si4 combined}

SI4_aligned_panels <- cowplot::align_plots(
  SI4a,
  SI4b,
  align = "v",
  axis = "lr"
)

SI_4 <- cowplot::plot_grid(
  SI4_aligned_panels[[1]],
  SI4_aligned_panels[[2]],
  ncol = 1,
  rel_heights = c(1, 1)
)

SI_4 

cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/SI_Figure4.png", SI_4 , base_width = 7,
  base_height = 7, dpi=600)

```

# Extended Data Figure 5: a) pH, b) Fe, Al, or Ca ColEx c) Fe, Al, or Ca TMP1


```{r load pH data}

ph <-  read_rds("~/GitHub/TEMPEST-1-porewater/data/averaged_ph_porewater.rds") 

```

### 5a
```{r pH timeseries}

startstudydate2 = lubridate::as_date("2023-05-01")

figure_y_limits <- c(as.Date(format(startstudydate2, "%Y-%m-01")), as.Date(endstudydate))

SI_5a <- ph %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = datetime, y = pH, color = plot)) +
 geom_vline(xintercept = WaterDeliveryStart2022, linetype = "dashed", color = "black") +
  geom_vline(xintercept = WaterDeliveryStart1, linetype = "dashed", color = "black") +
  geom_vline(xintercept = WaterDeliveryStart1_2024, linetype = "dashed", color = "black") +
   geom_pointrange(aes(ymin=pH-pH_sd, ymax=pH+pH_sd), shape=17, size=1)+
    scale_x_date(
  limits = figure_y_limits ,
  breaks = figure_x_breaks,
  labels = figure_x_labels,
  expand = expansion(mult = c(0, 0))
)+
  geom_label(
  data = event_labels,
  aes(x = event_dates, y = Inf, label = c("T1","T2","T3")),
  inherit.aes = FALSE,
  vjust = 2,
  label.size = 0.25,
  label.padding = unit(0.12, "lines"),
  fill = "white",
  color = "black",
  fontface = "bold",
  size = 4
)+
    ylab("porewater pH")+
    xlab("Date")+
    ylim(0,10)+
    scale_color_manual(values=Anyas_colors) +
    theme_classic() +
    theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_text(size = axis_title_size),
    axis.title.x = element_text(size = axis_title_size),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_text(size = axis_text_size),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )

SI_5a


```

# Supplemental Figure 4: ICP data from column experiment 

```{r icp column}

exp_col_dat <- read_csv("~/GitHub/tempest_ionic_strength/Data/Exp 3 Column Experiment ColEx Data/TEMPEST_Col_Ex_all_chem_averaged_L2.csv") %>%
  mutate(Wash = case_when(is.na(Timepoint) ~ 0,
                          TRUE ~ Timepoint),
    Treatment = case_when(Treatment == "SW" ~ "Artificial Saltwater",
                               Treatment == "FW" ~ "Artificial Freshwater",
                               Treatment == "Control" ~ "Control",
                               Treatment == "SWE" ~ "Artificial Saltwater",
                               Treatment == "CL" ~ "Artificial Freshwater",
                          TRUE ~ Treatment)) %>%
  mutate(Treatment = factor(Treatment, levels = Treatment_order))


```
Fe, Al, Ca
```{r icp col ex}
SI4_a <- exp_col_dat %>%
  filter(Treatment == "Artificial Saltwater")  %>%
   ggplot(aes(x = Wash, y = Fe_ppm_avg), color = "#C38D18") +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=Fe_ppm_avg-Fe_ppm_sd, ymax=Fe_ppm_avg+Fe_ppm_sd), shape=17, size = 1, color = "#C38D18")+
    ylab("Fe mg/L")+
    xlab(" ")+
   # title("Field Manipulation") +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "   a) "), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE) +
    theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_text(size = axis_title_size),
    axis.title.x = element_text(size = axis_title_size),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_text(size = axis_text_size),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )

print(SI4_a)

SI4_b <- exp_col_dat %>%
  filter(Treatment == "Artificial Saltwater")  %>%
   ggplot(aes(x = Wash, y = Al_ppm_avg), color = "#DCDCDA") +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=Al_ppm_avg-Al_ppm_sd, ymax=Al_ppm_avg+Al_ppm_sd), shape=17, size = 1, color = "#DCDCDA")+
    ylab("Al mg/L")+
    xlab(" ")+
   # title("Field Manipulation") +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "   b) "), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)+
    theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_text(size = axis_title_size),
    axis.title.x = element_text(size = axis_title_size),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_text(size = axis_text_size),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )

print(SI4_b)

SI4_c <- exp_col_dat %>%
  filter(Treatment == "Artificial Saltwater")  %>%
   ggplot(aes(x = Wash, y = Ca_ppm_avg), color = "#009DBD") +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=Ca_ppm_avg-Ca_ppm_sd, ymax=Ca_ppm_avg+Ca_ppm_sd), shape=17, size = 1, color = "#009DBD")+
    ylab("Ca mg/L")+
    xlab("Treatment Wash Number")+
   # title("Field Manipulation") +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "   c)"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)+
    theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_text(size = axis_title_size),
    axis.title.x = element_text(size = axis_title_size),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_text(size = axis_text_size),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )

print(SI4_c)

SI4 <- ggarrange(SI4_a, SI4_b, SI4_c, nrow = 3, legend = "bottom", common.legend = TRUE, align = "hv")

SI4

 cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/SI_Figure4.png",SI4, dpi=600,
                  nrow = 3)
```




# Supplemental Figure 6: ICP data from YR 1 through time 


```{r load ICP data}

icp <-  read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_POOL_ICP_2022_L1.csv") 

```

```{r icp timeseries}
ca <- icp %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = date, y = Ca_mgL_Porewater_average, color = plot)) +
    geom_point(shape=17, size=4)+
    geom_vline(xintercept = as.Date(WaterDeliveryStart2022),linetype="dashed",color="blue")+ 
    geom_vline(xintercept = as.Date("2022-07-21 15:59:00"),linetype="solid",color="grey")+
    ylab("Porewater Ca (mg/L)")+
    xlab("")+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16))

mg <- icp %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = date, y = Mg_mgL_Porewater_average, color = plot)) +
    geom_point(shape=17, size=4)+
    geom_vline(xintercept = as.Date(WaterDeliveryStart2022),linetype="dashed",color="grey")+ 
    ylab("Porewater Mg2+ mg/L")+
    xlab("Date")+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()

na <- icp %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = date, y = Na_mgL_Porewater_average, color = plot)) +
    geom_point(shape=17, size=4)+
    geom_vline(xintercept = as.Date(WaterDeliveryStart2022),linetype="dashed",color="grey")+ 
    ylab("Porewater Na+ mg/L")+
    xlab("Date")+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()

fe <-icp %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = date, y = Fe_mgL_Porewater_average, color = plot)) +
    geom_point(shape=17, size=4)+
    geom_vline(xintercept = as.Date(WaterDeliveryStart2022),linetype="dashed",color="blue")+ 
    geom_vline(xintercept = as.Date("2022-07-21 15:59:00"),linetype="solid",color="grey")+ 
    ylab("Porewater Fe (mg/L)")+
    xlab("")+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))  +
  theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16))

al <- icp %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = date, y = Al_mgL_Porewater_average, color = plot)) +
    geom_point(shape=17, size=4)+
    geom_vline(xintercept = as.Date(WaterDeliveryStart2022),linetype="dashed",color="blue")+ 
    geom_vline(xintercept = as.Date("2022-07-21 15:59:00"),linetype="solid",color="grey")+ 
    ylab("Porewater Al (mg/L)")+
    xlab("Date")+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))  +
  theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16))

SI6 <- ggarrange(fe, al, ca, nrow = 1, legend = "bottom", common.legend = TRUE, align = "hv")

SI6

 cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/FigureS6.png",SI6, dpi=600,
                  nrow = 1)

```

# Extended Data Figure 6: Jim's model output 

See: 

# Extended Data Figure 7: Jim made this
Incoming salts interact with soil exchange sites and mineralogy, which results in different recovery trajectories as the salinity of treatment waters increases

See: 


# Extended Data Figure 8: DOC after TMP 3

```{r load doc data through May 2026}

doc_after_t3 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/DOC_FigureS8.rds") %>%
  filter(nominal_date > WaterDeliveryStart1_2024 )
```

```{r plot fig S8}

breaks_si8 <- seq(
  from = min(doc_after_t3$nominal_date),
  to = max(doc_after_t3$nominal_date),
  by = "1 month"
)

FigS8 <- doc_after_t3 %>%
  ggplot(aes(
    x = nominal_date,
    y = doc_mg_l_mean,
    color = plot,
    shape = plot
  )) +
  geom_vline(xintercept = WaterDeliveryStart1_2024, linetype = "dashed", color = "black") +
geom_pointrange(
  aes(
    x = nominal_date,
    y = doc_mg_l_mean,
    ymin = doc_mg_l_mean - doc_mg_l_sd,
    ymax = doc_mg_l_mean + doc_mg_l_sd,
    color = plot
  ),
  size = 1,
    stroke = 0.7,
  linewidth = 0.6,
  linetype = "solid"
)  +
  geom_point(
    size = 1.8
  ) +
  labs(
    x = "Date",
    y = "DOC (mgC/L)" ,
    color = NULL,
    fill = NULL,
    shape = NULL
  ) +
  scale_color_manual(
  values = Anyas_colors
) +
    geom_label(
  data = event_labels,
  aes(x = event_dates[3], y = Inf, label = c("T3")),
  inherit.aes = FALSE,
  vjust = 2,
  label.size = 0.25,
  label.padding = unit(0.12, "lines"),
  fill = "white",
  color = "black",
  fontface = "bold",
  size = 4
)+
scale_shape_manual(
  values = c(15,20,17)
) +
scale_x_date(
  breaks = breaks_si8,
  labels = ifelse(seq_along(breaks_si8) %% 3 == 1,
                  format(breaks_si8, "%b\n%Y"),
                  "")
) +
  theme_classic() +
    theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_text(size = axis_title_size),
    axis.title.x = element_text(size = axis_title_size),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_text(size = axis_text_size),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length.x.bottom = unit(0.2, "cm"),
    axis.minor.ticks.x.bottom  = element_line(
    linewidth = 0.5,
    color = "black"
  ),
  axis.minor.ticks.length.x.bottom = unit(0.1, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )
    

FigS8 

cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/SI_Figure8.png",FigS8, dpi=600)
```





#graveyard 

### Supplemental Figure XX: Precip and DOC

```{r precip}
doc_precip <- read_csv("~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_allyrs_withprecip.csv")
precip <- read_csv("~/GitHub/TEMPEST-1-porewater/data/mean_monthly_precip_2006-2026.csv")
```

```{r plots doc rain}
top_y <- 100
bottom_y <- 0

max_precip <- max(doc_precip$mean_monthly_precip_cm, na.rm = TRUE)

scale_factor <- (top_y - bottom_y) / max_precip

precip_axis <- function(x) (top_y - x) / scale_factor

scale_factor

#note this only has precip when we have DOC data...
Figure1_combined <- doc_precip %>%
   ggplot(aes(x = stamp, y = doc_mg_l_mean, color = plot)) +
   geom_rect(aes(xmin = stamp -15,  xmax = stamp + 15, y = top_y - mean_monthly_precip_cm * scale_factor), ymax = top_y, inherit.aes = FALSE, fill = "skyblue", alpha = 0.6) +
   #geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab(" ")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    scale_y_continuous(limits = c(bottom_y, top_y),  name = "DOC (mg/L)", sec.axis = sec_axis(trans = precip_axis , name = "Mean Monthly Precipitation (cm)")
  ) +
    scale_color_manual(values=Anyas_colors) +
   # scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16),  axis.title.x = element_blank() )
          #, axis.title.x = element_blank(), axis.text.x = element_blank() ) 

Figure1_combined 

rain_doc <- doc_precip %>%
   ggplot() +
   geom_point(aes(x = stamp, y = doc_mg_l_mean, color= mean_monthly_precip_cm)) +
   facet_wrap(~plot)+
    theme_classic()

rain_doc

precip_study <- precip %>%
  filter(year > 2021)

   ggplot() +
   geom_col(aes(x = stamp, y =  mean_monthly_precip_cm * 100), data = precip_study) +
   geom_point(aes(x = stamp, y = doc_mg_l_mean, color = plot), data = doc_precip)+
         scale_color_manual(values=Anyas_colors) 



```



## SI 3 Option 2 pH vs DOC

```{r load ph year plot avgs}
ph_2 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_ph_porewater_TMP2.rds") %>%
  mutate(pH_sd = case_when(is.na(pH_sd) ~ 0,
                          TRUE ~ pH_sd))

ph_3 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_ph_porewater_TMP3.rds") %>%
  mutate(pH_sd = case_when(is.na(pH_sd) ~ 0,
                          TRUE ~ pH_sd))

ph_2n3 <- ph_2 %>%
  full_join(ph_3)
```

```{r merging}

ph_doc <- pw_doc_2 %>%
  full_join(pw_doc_3) %>%
  full_join(ph_2n3, by = c("adj_month", "year", "plot")) %>%
  select(-tdn_mg_l_mean, -tdn_mg_l_sd) %>%
  drop_na()
```

```{r plot ph vs doc }

SI3 <- ph_doc %>%
   filter(plot == "Saltwater") %>%
    ggplot(aes(x = pH_mean, y = doc_mg_l_mean)) +
    geom_pointrange(aes(xmin=pH_mean-pH_sd, xmax=pH_mean+pH_sd, ymin= doc_mg_l_mean - doc_mg_l_sd, ymax = doc_mg_l_mean + doc_mg_l_sd), shape=17)+ 
    ylab("porewater DOC")+
    xlab("porewater pH")+
    theme_classic()

SI3

cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/FigureS3.png",SI3, dpi=600)

```



```{r import doc data}

pw_doc_1 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP1_Figure1.rds") %>%
  mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd))

pw_doc_2 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP2_Figure1.rds") %>%
  mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd))

pw_doc_3 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP3_Figure1.rds") %>%
  mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd))

#pw_tmp1 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/doc_porewater_grids_TMP1_forfig1.rds")

#doc_iqr <- readRDS("~/GitHub/TEMPEST-1-porewater/data/doc_porewater_summarystats_forfig1.rds") %>%
#  mutate(collection_datetime = as.POSIXct("2022-05-18 15:59:00", tz= "EST" ),
#         doc_mg_l = NA)

```

```{r plot DOC}

Figure1a <- pw_doc_1 %>%
  group_by(adj_month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year)) %>%
  ungroup() %>%
   ggplot(aes(x = adj_month, y = doc_mg_l_mean, color = plot)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab(" ")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16),  axis.title.x = element_blank() )+
          #, axis.title.x = element_blank(), axis.text.x = element_blank() ) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  a) Field Year 1"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(Figure1a)

Figure1b <- pw_doc_2 %>%
  group_by(adj_month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year)) %>%
  ungroup() %>%
    ggplot(aes(x = adj_month, y = doc_mg_l_mean, color = plot)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab("DOC mgC/L")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1),
         text= element_text(size = 16) ) +
       #  , axis.title.x = element_blank(), axis.text.x = element_blank() ) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  b) Field Year 2"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(Figure1b)

Figure1c <- pw_doc_3 %>%
  group_by(adj_month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year)) %>%
  ungroup() %>%
    ggplot(aes(x = adj_month, y = doc_mg_l_mean, color = plot)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab(" ")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16), 
          axis.title.x = element_blank() ) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  c) Field Year 3"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(Figure1c)

Figure1b_LEGEND <- pw_doc_2 %>%
  group_by(adj_month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year)) %>%
  ungroup() %>%
    ggplot(aes(x = adj_month, y = doc_mg_l_mean, color = plot)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab("DOC mgC/L")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  b) Field Year 2"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

fig1_legend = get_legend(Figure1b_LEGEND)




 Figure1 <- ggarrange(Figure1a, Figure1b, Figure1c, nrow = 3, legend = "bottom", common.legend = TRUE, align = "hv")


 Figure1


 cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/Figure1.png",Figure1, dpi=600,
                   nrow = 3)

all_hands_fig <- ggarrange(Figure1a, Figure1b, Figure1c, nrow = 1, legend = "bottom", common.legend = TRUE, align = "hv")

all_hands_fig

```

```{r graveyard}
Figure1d <- exp_doc %>%
  mutate(Treatment = factor(Treatment, levels = Treatment_order)) %>%
    mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd)) %>%
    ggplot(aes(x = Wash, y = doc_mg_l_mean, color = Treatment)) +
    geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=16, size = 1)+
    ylab("DOC mgC/L")+
    xlab("Experimental Wash")+
   # title("Laboratory Simulation")+
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  d) Laboratory"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(Figure1d)

Figure1d_LEGEND <- exp_doc %>%
  mutate(Treatment = factor(Treatment, levels = Treatment_order)) %>%
    mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd)) %>%
    ggplot(aes(x = Wash, y = doc_mg_l_mean, color = Treatment)) +
    geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=16, size = 1)+
    ylab("DOC mgC/L")+
    xlab("Experimental Wash")+
   # title("Laboratory Simulation")+
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    theme_classic() +
    theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  d) Laboratory"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

fig1c_legend = get_legend(Figure1d_LEGEND)
```



SI4b <- exp_nacl_dat %>%
   ggplot(aes(x = Wash, y = Fe_ppm, color = Treatment)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=Fe_ppm-Fe_ppm_stdev, ymax=Fe_ppm+Fe_ppm_stdev), shape=17, size = 1)+
    ylab("Fe mg/L")+
    xlab("Treatment Wash Number")+
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,6)) +
    scale_color_manual(values = PNWColors::pnw_palette("Starfish",6))+
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    geom_text(data = data.frame(x = -Inf, y = Inf, label = "   a) "), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE) +
    theme(
    legend.position = "none",
    legend.title = element_blank(),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_blank(),
    axis.title.x =element_blank(),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_text(size = axis_text_size),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )

print(SI5_a)

SI5_b <- exp_nacl_dat %>%
   ggplot(aes(x = Wash, y = Al_ppm, color = Treatment)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=Al_ppm-Al_ppm_stdev, ymax=Al_ppm+Al_ppm_stdev), shape=17, size = 1)+
    ylab("Al mg/L")+
    xlab("Treatment Wash Number")+
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,6)) +
  scale_color_manual(values = PNWColors::pnw_palette("Starfish",6))+
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
   geom_text(data = data.frame(x = -Inf, y = Inf, label = "   b) "), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE) +
    theme(
    legend.position = "bottom",
    legend.title = element_text(size = axis_title_size),
    legend.text = element_text(size = legend_text_size),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    text = element_text(size = base_text_size),
    axis.text = element_text(size = axis_text_size),
    axis.title = element_text(size = axis_title_size),
    axis.title.x = element_text(size = axis_title_size),
    axis.title.y.right = element_text(color = "transparent", size = axis_title_size),
    axis.text.x = element_text(size = axis_text_size),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.ticks.length = unit(0.2, "cm"),
    plot.margin = margin(5.5, 5.5, 2, 5.5)
  )

print(SI5_b)

SI5 <- ggarrange(SI5_a, SI5_b, nrow = 2, legend = "bottom", common.legend = TRUE, align = "hv")

SI5

 cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/SI_Figure5.png",SI5, dpi=600,
                  nrow = 2)

