---
title: "Porewater DOC Manuscript Analysis"
author: "AMP"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(ggpubr)
library(viridis)
library(scales)
library(tidyverse)

#clean out your environment 
rm(list = ls())

Anyas_theme = 
        theme_set(theme_bw() +
        theme(panel.background = element_blank(),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(face = "bold", size = 14),
        legend.position = "bottom",
        legend.title = element_blank(),
        strip.text = element_text(size=16)))

plot_order <- c('Control', 'Freshwater','Saltwater')
time_order <- c('Pre', 'Mid','Post')

Anyas_colors = c("springgreen2", "cyan2", "violetred2")

Anyas_colors_figure1 =  c("violetred2", "cyan2","springgreen2")

Treatment_order <- c('Control', 'Artificial Freshwater','Artificial Saltwater')

#double check your wd
getwd()


```

### Set the study dates

```{r event and study dates}
endstudydate = lubridate::as_date("2024-01-31")
startstudydate = lubridate::as_date("2022-05-01")

endstudydate_yr1 = lubridate::as_date("2023-05-31")
startstudydate_yr1 = lubridate::as_date("2022-05-01")

endplotdate_yr1 = lubridate::as_date("2023-04-30")

year1_start = lubridate::as_date("2022-01-01")
year1_stop = lubridate::as_date("2022-12-31")

year2_start = lubridate::as_date("2023-01-01")
year2_stop = lubridate::as_date("2023-12-31")

year3_start = lubridate::as_date("2024-01-01")
year3_stop = lubridate::as_date("2024-12-31")

WaterDeliveryStart2022 = as.POSIXct("2022-06-22 05:30:00", tz = "EST")
WaterDeliveryStop2022 = as.POSIXct("2022-06-22 14:30:00", tz = "EST")

WaterDeliveryStart1 = as.POSIXct("2023-06-06 05:30:00", tz = "EST")
WaterDeliveryStop1 = as.POSIXct("2023-06-06 14:30:00", tz = "EST")

WaterDeliveryStart2 = as.POSIXct("2023-06-07 05:30:00", tz = "EST")
WaterDeliveryStop2 = as.POSIXct("2023-06-07 14:30:00", tz = "EST")

WaterDeliveryStart1_2024 = as.POSIXct("2024-06-11 06:10:00", tz = "EST")
WaterDeliveryStart2_2024 = as.POSIXct("2024-06-12 06:08:00", tz = "EST")
WaterDeliveryStart3_2024 = as.POSIXct("2024-06-13 06:21:00", tz = "EST")

flood_event_datetime <- c(as_date("2022-06-22"))
```

```{r import doc data}

pw_doc_1 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP1_Figure1.rds") %>%
  mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd))

pw_doc_2 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP2_Figure1.rds") %>%
  mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd))

pw_doc_3 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP3_Figure1.rds") %>%
  mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd))

#pw_tmp1 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/doc_porewater_grids_TMP1_forfig1.rds")

#doc_iqr <- readRDS("~/GitHub/TEMPEST-1-porewater/data/doc_porewater_summarystats_forfig1.rds") %>%
#  mutate(collection_datetime = as.POSIXct("2022-05-18 15:59:00", tz= "EST" ),
#         doc_mg_l = NA)

exp_doc <- read_csv("~/GitHub/tempest_ionic_strength/Data/Processed Data/DOC/TEMPEST_ColEx_NPOC_TDN_meansforCDOM.csv") %>%
  mutate(Wash = case_when(is.na(Wash) ~ 0,
                          TRUE ~ Wash),
    Treatment = case_when(Treatment == "AS" ~ "Artificial Saltwater",
                               Treatment == "FW" ~ "Artificial Freshwater",
                               Treatment == "Control" ~ "Control",
                               Treatment == "SWE" ~ "Artificial Saltwater",
                               Treatment == "CL" ~ "Artificial Freshwater",
                          TRUE ~ Treatment))

```

# Figure 1: Year 1, 2, 3 DOC Field
Supporting Main Message: This phenomenon can be replicated in the laboratory; AND we can repeat it in the field the second year “reset” 

```{r plot DOC}

Figure1a <- pw_doc_1 %>%
  group_by(adj_month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year)) %>%
  ungroup() %>%
   ggplot(aes(x = adj_month, y = doc_mg_l_mean, color = plot)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab(" ")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16), axis.title.x = element_blank(), axis.text.x = element_blank() ) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  a) Field Year 1"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(Figure1a)

Figure1b <- pw_doc_2 %>%
  group_by(adj_month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year)) %>%
  ungroup() %>%
    ggplot(aes(x = adj_month, y = doc_mg_l_mean, color = plot)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab("DOC mgC/L")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1),
         text= element_text(size = 16), axis.title.x = element_blank(), axis.text.x = element_blank() ) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  b) Field Year 2"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(Figure1b)

Figure1c <- pw_doc_3 %>%
  group_by(adj_month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year)) %>%
  ungroup() %>%
    ggplot(aes(x = adj_month, y = doc_mg_l_mean, color = plot)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab(" ")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  c) Field Year 3"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(Figure1c)

Figure1b_LEGEND <- pw_doc_2 %>%
  group_by(adj_month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year)) %>%
  ungroup() %>%
    ggplot(aes(x = adj_month, y = doc_mg_l_mean, color = plot)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab("DOC mgC/L")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  b) Field Year 2"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

fig1_legend = get_legend(Figure1b_LEGEND)




Figure1 <- ggarrange(Figure1a, Figure1b, Figure1c, nrow = 3, legend = "bottom", common.legend = TRUE, align = "hv")


Figure1


cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/Figure1_pw_exp_doc.png",Figure1, dpi=600,
                   nrow = 3)


```

#Figure 4 OM chemistry shifts
The changes in concentration observed are coupled with a dramatic shift in the organic matter chemistry (both during and after the event). 
```{r load om chem data}

fticrms_relabund_avg <- read_rds("~/GitHub/TEMPEST-1-porewater/data/avg_relativeabundance_premidpost_fticrms_porewater.rds")

fticrms_relabund_dates <-read_rds("~/GitHub/TEMPEST-1-porewater/data/avg_relabund_dates_fticrms_porewater.rds")

```

```{r load cdom data}

cdom_metrics_avg <- read_rds("~/GitHub/TEMPEST-1-porewater/data/avg_metrics_premidpost_cdom_porewater.rds") %>%
  mutate(value_sd = case_when(is.na(value_sd) ~ 0,
                              TRUE ~ value_sd))

```

## Figure 4 A

```{r om chemistry plot}
colorblind_friendly_custom <- c("#66C2A5", "#FC8D62", "#FFD92F", "#E78AC3", "#A6D854", "#8DA0CB")

figure3a <- fticrms_relabund_avg %>%
  mutate(timedate_bin = factor(timedate_bin, levels = c("Pre", "Mid", "Post"))) %>%
  arrange(timedate_bin) %>%
  group_by(plot, timedate_bin) %>%
  ggplot() +
  geom_bar(aes(x = timedate_bin, y = relabund_mean, fill = Class_detailed), stat= "identity") +
    labs(x = "Event Timepoint", y = "Relative Abundance (%)", fill = "Assigned FTICR-MS groupings", title = "a)")+
    scale_fill_manual(values = colorblind_friendly_custom)  + 
  facet_wrap(~factor(plot, level=plot_order), nrow = 1, 
             labeller = label_wrap_gen(10)) +
  theme_classic() +
  theme(legend.position = "bottom",
        text = element_text(size = 14),          # Global text size
        axis.title = element_text(size=16),      # Axis title size
        axis.text = element_text(size=14),       # Axis text size
        legend.title = element_text(size=14),    # Legend title size
        legend.text = element_text(size=12),     # Legend text size
        strip.text = element_text(size=14),      # Facet title size
        plot.title = element_text(size=18))      # Plot title size
  

figure3a <- fticrms_relabund_avg %>%
  mutate(timedate_bin = factor(timedate_bin, levels = c("Pre", "Mid", "Post"))) %>%
  arrange(timedate_bin) %>%
  filter(Class_detailed == "highly aromatic") %>%
  group_by(plot, timedate_bin) %>%
  ggplot() +
  geom_bar(aes(x = timedate_bin, y = relabund_mean), stat= "identity") +
    labs(x = "Event Timepoint", y = "Relative Abundance of aliphatics (%)",  title = "a)")+
    scale_fill_manual(values = colorblind_friendly_custom)  + 
  facet_wrap(~factor(plot, level=plot_order), nrow = 1, 
             labeller = label_wrap_gen(10)) +
  theme_classic() +
  theme(legend.position = "bottom",
        text = element_text(size = 14),          # Global text size
        axis.title = element_text(size=16),      # Axis title size
        axis.text = element_text(size=14),       # Axis text size
        legend.title = element_text(size=14),    # Legend title size
        legend.text = element_text(size=12),     # Legend text size
        strip.text = element_text(size=14),      # Facet title size
        plot.title = element_text(size=18))      # Plot title size

figure3b <- cdom_metrics_avg %>%
  mutate(timedate_bin = factor(timedate_bin, levels = c("Pre", "Mid", "Post"))) %>%
  arrange(timedate_bin) %>%
  group_by(plot, timedate_bin) %>%
  filter(metric == "suva_254") %>%
  ggplot() +
  geom_bar(aes(x = timedate_bin, y = value_mean), stat= "identity") +
  geom_errorbar(aes(x = timedate_bin, ymin = value_mean-value_sd, ymax = value_mean + value_sd))+
    labs(x = "Event Timepoint", y = "SUVA", title = "b)")+
  facet_wrap(~factor(plot, level=plot_order), nrow = 1, 
             labeller = label_wrap_gen(10)) +
  theme_classic() +
  theme(legend.position = "bottom",
        text = element_text(size = 14),          # Global text size
        axis.title = element_text(size=16),      # Axis title size
        axis.text = element_text(size=14),       # Axis text size
        legend.title = element_text(size=14),    # Legend title size
        legend.text = element_text(size=12),     # Legend text size
        strip.text = element_text(size=14),      # Facet title size
        plot.title = element_text(size=18))      # Plot title size

#cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/figure3_pw_FTICRMS.png",figure3, dpi=300, base_asp = 1.5, base_height = 8, base_width = 14)
```

```{r timeline}
##
fticrms_relabund_dates %>%
  filter(Class_detailed == "highly aromatic") %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = date, y = relabund_mean, color = plot)) +
    geom_vline(xintercept = flood_event_datetime,linetype="dashed",color="grey")+ 
   geom_pointrange(aes(ymin=relabund_mean-relabund_sd, ymax=relabund_mean+relabund_sd), shape=17, size=1)+
    ylab("Relative abundance of highly aromatic group (%)")+
    xlab("Date")+
    ylim(0,40)+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()
```

## Figure 4 B 
- Slope Ratio (proxy for HMW/LMW) through year (FW & SW plots, TMP 1 & 2) https://aslopubs.onlinelibrary.wiley.com/doi/abs/10.4319/lo.2008.53.3.0955
- SUVA (using average DOC for the plots?) (FW & SW plots, TMP 1 & 2)

EDITS:spatial variability in conductivity is small, but spatial characteristics of DOM chemistry/character may be important/driving trends; and or spatial distribution of aggregates
- plot year 2 as a function of DOC vs chemistry metrics (SUVA/ SR)
## Figure 5 Mechanisms

```{r load cond doc data}
cond_doc <- read_rds("~/GitHub/TEMPEST-1-porewater/data/DOC_cond_data_for_relationships_SALT.rds")

cond_doc_control <- read_rds("~/GitHub/TEMPEST-1-porewater/data/DOC_cond_data_for_relationships_CONTROLPLOT.rds")

```

```{r make line for fig 5}

start_a <- max(cond_doc$doc_mg_l) 
start_b <- -0.001

# Fit the model using nls
fit_nls <- nls(doc_mg_l ~ a * exp(b * mean_EC_uScm_15cm), data = cond_doc, start = list(a = start_a, b = start_b))

# Summary of the model
summary(fit_nls)
# Extract fitted values
coefficients <- coef(fit_nls)
a <- coefficients["a"]
b <- coefficients["b"]

# Create a prediction data frame
new_data <- data.frame(mean_EC_uScm_15cm = seq(min(cond_doc$mean_EC_uScm_15cm), max(cond_doc$mean_EC_uScm_15cm), length.out = 100))
new_data$pred <- predict(fit_nls, newdata = new_data)

# Compute standard error of the fit
se_fit <- sqrt(sum(residuals(fit_nls)^2) / df.residual(fit_nls))
new_data$lower <- new_data$pred - 1.96 * se_fit
new_data$upper <- new_data$pred + 1.96 * se_fit

```

```{r figure 5}
pal= wesanderson::wes_palette("Zissou1", 21, type = "continuous")

condFig <- cond_doc %>%
  ggplot() +
  geom_point(data= , aes(x = max_EC_uScm_15cm, y = doc_mg_l, color = adj_month), shape = 16, size=5) +
  stat_smooth(aes(x = mean_EC_uScm_15cm, y = doc_mg_l),
    method="nls",
    formula = y ~ a * exp(b * x),
    method.args = list(start = list(a = a, b = b)),    # Starting values for a and b
    se=FALSE) +
  scale_color_gradientn(colors = pal)+
 # geom_ribbon(data = new_data, aes(x= max_EC_uScm_15cm, ymin = lower, ymax = upper), fill = "blue", alpha = 0.2) +
  ylab("DOC (mg/L)") +
  xlab("Mean Soil Conductivity during lysimeter collection") +
  theme_classic()

condFig

plotly::ggplotly(condFig)
```


# Supplemental Figures
1. SPC through year (SW plot, TMP 1 & 2)
2. pH through year (FW and SW plot, TMP 2)
3. DO before/during/after event (FW and SW plot, TMP 1 & 2? Discuss with Peter)

# Supplemental Figure 1: SPC timeseries 
```{r bring in SPC}

pw_ec <- read_rds("~/GitHub/TEMPEST-1-porewater/data/averaged_by_plot_ec_teros_porewater.rds")

pw_ec_all <- read_rds("~/GitHub/TEMPEST-1-porewater/processed data/TMP_PW_EC_15cm_2021-2025_L2.rds")

pw_doc_dates <- read_rds("~/GitHub/TEMPEST-1-porewater/data/dates_DOC_higher_than_variance.rds")

pw_doc_dates
```

```{r set dates based on pw_doc_dates}
tmp1_doc_high = as.POSIXct("2022-07-21 15:59:00", tz = "EST")
tmp2_doc_high = as.POSIXct("2023-08-10 15:53:00", tz = "EST")
tmp3_doc_high = as.POSIXct("2025-01-31 20:52:00", tz = "EST")

```

```{r spc timeseries}

#rectangles from cond_doc_threshold in `1_manuscript_anlaysis_porewaterDOC.Rmd`

pw_ec %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = datetime_est, y = EC_uScm, color = plot)) +
    geom_point()+
    geom_vline(xintercept = tmp1_doc_high,linetype="dashed",color="black")+ 
    geom_vline(xintercept = tmp2_doc_high,linetype="dashed",color="black")+ 
    geom_vline(xintercept = tmp3_doc_high,linetype="dashed",color="black")+ 
     annotate("rect", xmin= min(pw_ec$datetime_est) , xmax = max(pw_ec$datetime_est), ymin = (235.3203 - 28.85163), ymax = (235.3203 + 28.85163), fill= "grey", alpha = 0.2)+
    ylab("mean Porewater Conductivity (uS/cm)")+
    xlab("Date")+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()




```

pw_ec_all %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = datetime_est, y = EC_uScm, color = plot)) +
    geom_point()+
    geom_vline(xintercept = WaterDeliveryStart2022,linetype="dashed",color="grey")+ 
    geom_vline(xintercept = WaterDeliveryStart2,linetype="dashed",color="grey")+ 
    geom_vline(xintercept = WaterDeliveryStart1,linetype="dashed",color="grey")+ 
    ylab("mean Porewater Conductivity (uS/cm)")+
    xlab("Date")+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()


# Supplmental Figure 2: pH timeseries 

```{r load pH data}

ph <-  read_rds("~/GitHub/TEMPEST-1-porewater/data/averaged_ph_porewater.rds") 

```


```{r pH timeseries}
ph %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = datetime, y = pH, color = plot)) +
    geom_vline(xintercept = WaterDeliveryStart1,linetype="dashed",color="grey")+ 
    geom_vline(xintercept = WaterDeliveryStart2,linetype="dashed",color="grey")+
    geom_vline(xintercept = WaterDeliveryStart1_2024,linetype="dashed",color="grey")+
   geom_vline(xintercept = WaterDeliveryStart2_2024,linetype="dashed",color="grey")+
   geom_vline(xintercept = WaterDeliveryStart3_2024,linetype="dashed",color="grey")+
   geom_pointrange(aes(ymin=pH-pH_sd, ymax=pH+pH_sd), shape=17, size=1)+
   annotate("rect", xmin= min(ph$datetime) , xmax = max(ph$datetime), ymin = (6.576429 - 0.4931158), ymax = (6.576429 + 0.4931158), fill= "blue", alpha = 0.2)+
    ylab("porewater pH")+
    xlab("Date")+
    ylim(0,10)+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()
```

# Option 2 pH vs DOC

```{r load ph year plot avgs}
ph_2 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_ph_porewater_TMP2.rds") %>%
  mutate(pH_sd = case_when(is.na(pH_sd) ~ 0,
                          TRUE ~ pH_sd))

ph_3 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_ph_porewater_TMP3.rds") %>%
  mutate(pH_sd = case_when(is.na(pH_sd) ~ 0,
                          TRUE ~ pH_sd))

ph_2n3 <- ph_2 %>%
  full_join(ph_3)
```

```{r merging}

ph_doc <- pw_doc_2 %>%
  full_join(pw_doc_3) %>%
  full_join(ph_2n3, by = c("adj_month", "year", "plot")) %>%
  select(-tdn_mg_l_mean, -tdn_mg_l_sd) %>%
  drop_na()
```

```{r plot ph vs doc }

ph_doc %>%
   filter(plot == "Saltwater") %>%
    ggplot(aes(x = pH_mean, y = doc_mg_l_mean)) +
    geom_pointrange(aes(xmin=pH_mean-pH_sd, xmax=pH_mean+pH_sd, ymin= doc_mg_l_mean - doc_mg_l_sd, ymax = doc_mg_l_mean + doc_mg_l_sd), shape=17)+ 
    ylab("porewater DOC")+
    xlab("porewater pH")+
    theme_classic()
```




```{r load ICP data}

icp <-  read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_POOL_ICP_2022_L1.csv") 

```

```{r icp timeseries}
icp %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = date, y = Ca_mgL_Porewater_average, color = plot)) +
    geom_point(shape=17, size=4)+
    geom_vline(xintercept = as.Date(WaterDeliveryStart2022),linetype="dashed",color="grey")+ 
    ylab("Porewater Ca2+ mg/L")+
    xlab("Date")+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()

icp %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = date, y = Mg_mgL_Porewater_average, color = plot)) +
    geom_point(shape=17, size=4)+
    geom_vline(xintercept = as.Date(WaterDeliveryStart2022),linetype="dashed",color="grey")+ 
    ylab("Porewater Mg2+ mg/L")+
    xlab("Date")+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()
```

```{r graveyard}
Figure1d <- exp_doc %>%
  mutate(Treatment = factor(Treatment, levels = Treatment_order)) %>%
    mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd)) %>%
    ggplot(aes(x = Wash, y = doc_mg_l_mean, color = Treatment)) +
    geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=16, size = 1)+
    ylab("DOC mgC/L")+
    xlab("Experimental Wash")+
   # title("Laboratory Simulation")+
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  d) Laboratory"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(Figure1d)

Figure1d_LEGEND <- exp_doc %>%
  mutate(Treatment = factor(Treatment, levels = Treatment_order)) %>%
    mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd)) %>%
    ggplot(aes(x = Wash, y = doc_mg_l_mean, color = Treatment)) +
    geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=16, size = 1)+
    ylab("DOC mgC/L")+
    xlab("Experimental Wash")+
   # title("Laboratory Simulation")+
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    theme_classic() +
    theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  d) Laboratory"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

fig1c_legend = get_legend(Figure1d_LEGEND)
```


