---
title: "Porewater DOC Manuscript Analysis"
author: "AMP"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(ggpubr)
library(viridis)
library(scales)
Anyas_theme = 
        theme_set(theme_bw() +
        theme(panel.background = element_blank(),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(face = "bold", size = 14),
        legend.position = "bottom",
        legend.title = element_blank(),
        strip.text = element_text(size=16)))

plot_order <- c('Control', 'Freshwater','Estuarine-water')
time_order <- c('Pre', 'Mid','Post')

Anyas_colors = c("springgreen2", "cyan2", "violetred2")

Treatment_order <- c('Control', 'Artificial Freshwater','Artificial Estuarine-water')

#double check your wd
getwd()
```

### Set the study dates

```{r event and study dates}
endstudydate = lubridate::as_date("2024-01-31")
startstudydate = lubridate::as_date("2022-05-01")

endstudydate_yr1 = lubridate::as_date("2023-05-31")
startstudydate_yr1 = lubridate::as_date("2022-05-01")

year1_start = lubridate::as_date("2022-01-01")
year1_stop = lubridate::as_date("2022-12-31")

year2_start = lubridate::as_date("2023-01-01")
year2_stop = lubridate::as_date("2023-12-31")

year3_start = lubridate::as_date("2024-01-01")
year3_stop = lubridate::as_date("2024-12-31")

WaterDeliveryStart2022 = as.POSIXct("2022-06-22 05:30:00", tz = "EST")
WaterDeliveryStop2022 = as.POSIXct("2022-06-22 14:30:00", tz = "EST")

WaterDeliveryStart1 = as.POSIXct("2023-06-06 05:30:00", tz = "EST")
WaterDeliveryStop1 = as.POSIXct("2023-06-06 14:30:00", tz = "EST")

WaterDeliveryStart2 = as.POSIXct("2023-06-07 05:30:00", tz = "EST")
WaterDeliveryStop2 = as.POSIXct("2023-06-07 14:30:00", tz = "EST")
```

```{r import doc data}

pw_doc <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater.rds")

exp_doc <- read_csv("~/GitHub/tempest_ionic_strength/Data/Processed Data/DOC/TEMPEST_ColEx_NPOC_TDN_meansforCDOM.csv") %>%
  mutate(Wash = case_when(is.na(Wash) ~ 0,
                          TRUE ~ Wash),
    Treatment = case_when(Treatment == "AS" ~ "Artificial Estuarine-water",
                               Treatment == "FW" ~ "Artificial Freshwater",
                               Treatment == "DI" ~ "Control",
                               Treatment == "SWE" ~ "Artificial Estuarine-water",
                               Treatment == "CL" ~ "Artificial Freshwater",
                          TRUE ~ Treatment))

```


# Figure 1: Year 1 DOC 
Main Message 1 (Hypothesis rejected): There is a prolonged and sustained increase in dissolved carbon following initial drop in DOC after exposure to salinity.

```{r plot DOC}

Figure1 <- pw_doc %>%
  filter(between(evacuation_date, startstudydate_yr1, endstudydate_yr1)) %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = collection_datetime, y = doc_mg_l, color = plot)) +
    geom_vline(xintercept = WaterDeliveryStart2022,linetype="dashed",color="grey")+ 
    geom_vline(xintercept = WaterDeliveryStart1,linetype="dashed",color="grey")+ 
    geom_vline(xintercept = WaterDeliveryStart2,linetype="dashed",color="grey")+
   geom_pointrange(aes(ymin=doc_mg_l-doc_mg_l_sd, ymax=doc_mg_l+doc_mg_l_sd), shape=17, size=1)+
    ylab("DOC mgC/L")+
    xlab("Date")+
    ylim(0,100)+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()

print(Figure1)

plotly::ggplotly(Figure1)

cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/Figure1_pwdoc.png",Figure1, dpi=300)

```

# Figure 2: Year 1 & 2 DOC 
Supporting Main Message: This phenomenon can be replicated in the laboratory; AND we can repeat it in the field the second year “reset” 

```{r plot DOC}

Figure2a <- pw_doc %>%
  filter(between(evacuation_date, startstudydate_yr1, endstudydate)) %>%
  mutate(year = lubridate::year(evacuation_date), 
         month = lubridate::month(evacuation_date),
           # Adjust months: May = 0, June = 1, ..., April = 11
         adj_month = (month(evacuation_date) - 5) %% 12 ) %>%
  group_by(month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year))%>%
  ungroup() %>%
    ggplot(aes(x = adj_month, y = doc_mg_l, color = plot, shape = year)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l-doc_mg_l_sd, ymax=doc_mg_l+doc_mg_l_sd), size = 1)+
    ylab("DOC mgC/L")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(0,100)+
    xlim(0,12)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks()) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "bottom") +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  a) Field"), aes(x, y, label = label), hjust = -0.1, vjust = 1.1, inherit.aes = FALSE)

print(Figure2a)

Figure2b <- exp_doc %>%
  mutate(Treatment = factor(Treatment, levels = Treatment_order)) %>%
    ggplot(aes(x = Wash, y = doc_mg_l_mean, color = Treatment)) +
    geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab("DOC mgC/L")+
    xlab("Experimental Wash")+
   # title("Laboratory Simulation")+
    ylim(0,100)+
    xlim(0,12)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks()) +
    theme_classic() +
    theme(legend.position = "bottom") +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  b) Laboratory"), aes(x, y, label = label), hjust = -0.1, vjust = 1.1, inherit.aes = FALSE)

print(Figure2b)

Figure2 <- ggarrange(Figure2a, Figure2b)

Figure2

cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/Figure2_pw_exp_doc.png",Figure2, dpi=300, base_asp = 3.1)

```

#Figure 4 OM chemistry shifts
The changes in concentration observed are coupled with a dramatic shift in the organic matter chemistry (both during and after the event). 
```{r load om chem data}

fticrms_relabund_avg <- read_rds("~/GitHub/TEMPEST-1-porewater/data/avg_relativeabundance_premidpost_fticrms_porewater.rds")

fticrms_relabund_dates <-read_rds("~/GitHub/TEMPEST-1-porewater/data/avg_relabund_dates_fticrms_porewater.rds")

```

## Figure 4 A

```{r om chemistry plot}
colorblind_friendly_custom <- c("#66C2A5", "#FC8D62", "#FFD92F", "#E78AC3", "#A6D854", "#8DA0CB")
flood_event_datetime <- c(as_date("2022-06-22"))

figure3 <- fticrms_relabund_avg %>%
  mutate(timedate_bin = factor(timedate_bin, levels = c("Pre", "Mid", "Post"))) %>%
  arrange(timedate_bin) %>%
  group_by(plot, timedate_bin) %>%
  ggplot() +
  geom_bar(aes(x = timedate_bin, y = relabund_mean, fill = Class_detailed), stat= "identity") +
    labs(x = "Event Timepoint", y = "Relative Abundance (%)", fill = "Assigned FTICR-MS groupings", title = "a)")+
    scale_fill_manual(values = colorblind_friendly_custom)  + 
  facet_wrap(~factor(plot, level=plot_order), nrow = 1, 
             labeller = label_wrap_gen(10)) +
  theme_classic() +
  theme(legend.position = "bottom",
        text = element_text(size = 14),          # Global text size
        axis.title = element_text(size=16),      # Axis title size
        axis.text = element_text(size=14),       # Axis text size
        legend.title = element_text(size=14),    # Legend title size
        legend.text = element_text(size=12),     # Legend text size
        strip.text = element_text(size=14),      # Facet title size
        plot.title = element_text(size=18))      # Plot title size

cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/figure3_pw_FTICRMS.png",figure3, dpi=300, base_asp = 1.5, base_height = 8, base_width = 14)

##
fticrms_relabund_dates %>%
  filter(Class_detailed == "highly aromatic") %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = date, y = relabund_mean, color = plot)) +
    geom_vline(xintercept = flood_event_datetime,linetype="dashed",color="grey")+ 
   geom_pointrange(aes(ymin=relabund_mean-relabund_sd, ymax=relabund_mean+relabund_sd), shape=17, size=1)+
    ylab("Relative abundance of highly aromatic group (%)")+
    xlab("Date")+
    ylim(0,40)+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()
```

## Figure 4 B 
- Slope Ratio (proxy for HMW/LMW) through year (FW & SW plots, TMP 1 & 2) https://aslopubs.onlinelibrary.wiley.com/doi/abs/10.4319/lo.2008.53.3.0955
- SUVA (using average DOC for the plots?) (FW & SW plots, TMP 1 & 2)

# Supplemental Figures
1. SPC through year (SW plot, TMP 1 & 2)
2. pH through year (FW and SW plot, TMP 2)
3. DO before/during/after event (FW and SW plot, TMP 1 & 2? Discuss with Peter)

# Supplemental Figure 1: SPC timeseries 

```{r spc timeseries}

  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = date, y = relabund_mean, color = plot)) +
    geom_vline(xintercept = flood_event_datetime,linetype="dashed",color="grey")+ 
   geom_pointrange(aes(ymin=relabund_mean-relabund_sd, ymax=relabund_mean+relabund_sd), shape=17, size=1)+
    ylab("Relative abundance of highly aromatic group (%)")+
    xlab("Date")+
    ylim(0,40)+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()
```


# Supplmental Figure 2: pH timeseries 

```{r load pH data}

ph <-  read_rds("~/GitHub/TEMPEST-1-porewater/data/averaged_ph_porewater.rds") 

```

```{r pH timeseries}
ph %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = datetime, y = pH, color = plot)) +
    geom_vline(xintercept = WaterDeliveryStart1,linetype="dashed",color="grey")+ 
    geom_vline(xintercept = WaterDeliveryStart2,linetype="dashed",color="grey")+
   geom_pointrange(aes(ymin=pH-pH_sd, ymax=pH+pH_sd), shape=17, size=1)+
    ylab("porewater pH")+
    xlab("Date")+
    ylim(0,10)+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()
```



