---
title: "Porewater DOC Manuscript Analysis"
author: "AMP"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(ggpubr)
library(viridis)
library(scales)
library(tidyverse)

#clean out your environment 
rm(list = ls())

Anyas_theme = 
        theme_set(theme_bw() +
        theme(panel.background = element_blank(),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(face = "bold", size = 14),
        legend.position = "bottom",
        legend.title = element_blank(),
        strip.text = element_text(size=16)))

plot_order <- c('Control', 'Freshwater','Saltwater')
time_order <- c('Pre', 'Mid','Post')

Anyas_colors = c("springgreen2", "cyan2", "violetred2")

Anyas_colors_figure1 =  c("violetred2", "cyan2","springgreen2")

Treatment_order <- c('Control', 'Artificial Freshwater','Artificial Saltwater')

#double check your wd
getwd()


```

### Set the study dates

```{r event and study dates}
endstudydate = lubridate::as_date("2024-01-31")
startstudydate = lubridate::as_date("2022-05-01")

endstudydate_yr1 = lubridate::as_date("2023-05-31")
startstudydate_yr1 = lubridate::as_date("2022-05-01")

endplotdate_yr1 = lubridate::as_date("2023-04-30")

year1_start = lubridate::as_date("2022-01-01")
year1_stop = lubridate::as_date("2022-12-31")

year2_start = lubridate::as_date("2023-01-01")
year2_stop = lubridate::as_date("2023-12-31")

year3_start = lubridate::as_date("2024-01-01")
year3_stop = lubridate::as_date("2024-12-31")

WaterDeliveryStart2022 = as.POSIXct("2022-06-22 05:30:00", tz = "EST")
WaterDeliveryStop2022 = as.POSIXct("2022-06-22 14:30:00", tz = "EST")

WaterDeliveryStart1 = as.POSIXct("2023-06-06 05:30:00", tz = "EST")
WaterDeliveryStop1 = as.POSIXct("2023-06-06 14:30:00", tz = "EST")

WaterDeliveryStart2 = as.POSIXct("2023-06-07 05:30:00", tz = "EST")
WaterDeliveryStop2 = as.POSIXct("2023-06-07 14:30:00", tz = "EST")

WaterDeliveryStart1_2024 = as.POSIXct("2024-06-11 06:10:00", tz = "EST")
WaterDeliveryStart2_2024 = as.POSIXct("2024-06-12 06:08:00", tz = "EST")
WaterDeliveryStart3_2024 = as.POSIXct("2024-06-13 06:21:00", tz = "EST")

flood_event_datetime <- c(as_date("2022-06-22"))
```

```{r import doc data}

pw_doc_1 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP1_Figure1.rds") %>%
  mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd))

pw_doc_2 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP2_Figure1.rds") %>%
  mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd))

pw_doc_3 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_TMP3_Figure1.rds") %>%
  mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd))

#pw_tmp1 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/doc_porewater_grids_TMP1_forfig1.rds")

#doc_iqr <- readRDS("~/GitHub/TEMPEST-1-porewater/data/doc_porewater_summarystats_forfig1.rds") %>%
#  mutate(collection_datetime = as.POSIXct("2022-05-18 15:59:00", tz= "EST" ),
#         doc_mg_l = NA)

```

# Figure 1: Year 1, 2, 3 DOC Field
Supporting Main Message: This phenomenon can be replicated in the laboratory; AND we can repeat it in the field the second year “reset” 

```{r plot DOC}

Figure1a <- pw_doc_1 %>%
  group_by(adj_month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year)) %>%
  ungroup() %>%
   ggplot(aes(x = adj_month, y = doc_mg_l_mean, color = plot)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab(" ")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16),  axis.title.x = element_blank() )+
          #, axis.title.x = element_blank(), axis.text.x = element_blank() ) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  a) Field Year 1"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(Figure1a)

Figure1b <- pw_doc_2 %>%
  group_by(adj_month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year)) %>%
  ungroup() %>%
    ggplot(aes(x = adj_month, y = doc_mg_l_mean, color = plot)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab("DOC mgC/L")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1),
         text= element_text(size = 16) ) +
       #  , axis.title.x = element_blank(), axis.text.x = element_blank() ) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  b) Field Year 2"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(Figure1b)

Figure1c <- pw_doc_3 %>%
  group_by(adj_month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year)) %>%
  ungroup() %>%
    ggplot(aes(x = adj_month, y = doc_mg_l_mean, color = plot)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab(" ")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16), 
          axis.title.x = element_blank() ) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  c) Field Year 3"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(Figure1c)

Figure1b_LEGEND <- pw_doc_2 %>%
  group_by(adj_month) %>%
  mutate(plot = factor(plot, levels = plot_order),
         Timepoint = cur_group_id(),
         year= as.character(year)) %>%
  ungroup() %>%
    ggplot(aes(x = adj_month, y = doc_mg_l_mean, color = plot)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=17, size = 1)+
    ylab("DOC mgC/L")+
    xlab("Sampling Month Post-Event")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  b) Field Year 2"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

fig1_legend = get_legend(Figure1b_LEGEND)




# Figure1 <- ggarrange(Figure1a, Figure1b, Figure1c, nrow = 3, legend = "bottom", common.legend = TRUE, align = "hv")


# Figure1


# cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/Figure1.png",Figure1, dpi=600,
#                   nrow = 3)

all_hands_fig <- ggarrange(Figure1a, Figure1b, Figure1c, nrow = 1, legend = "bottom", common.legend = TRUE, align = "hv")

all_hands_fig

```

# Supplemental Figures
1. SPC through year (SW plot, TMP 1 & 2)
2. Column Experiment DOC
2. pH through year (FW and SW plot, TMP 2)
3. DO before/during/after event (FW and SW plot, TMP 1 & 2? Discuss with Peter)

Water Isotopes (from Lindsey)


# Supplemental Figure 1: SPC timeseries 
```{r bring in SPC}

pw_ec <- read_rds("~/GitHub/TEMPEST-1-porewater/data/averaged_by_plot_ec_teros_porewater.rds")

pw_ec_all <- read_rds("~/GitHub/TEMPEST-1-porewater/processed data/TMP_PW_EC_15cm_2021-2025_L2.rds")

pw_doc_dates <- read_rds("~/GitHub/TEMPEST-1-porewater/data/dates_DOC_higher_than_variance.rds")

pw_doc_dates
```

```{r set dates based on pw_doc_dates}
tmp1_doc_high = as.POSIXct("2022-07-21 15:59:00", tz = "EST")
tmp2_doc_high = as.POSIXct("2023-08-10 15:53:00", tz = "EST")
tmp3_doc_high = as.POSIXct("2025-01-31 20:52:00", tz = "EST")

```

```{r spc timeseries}

#rectangles from cond_doc_threshold in `1_manuscript_anlaysis_porewaterDOC.Rmd`

pw_ec %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = datetime_est, y = EC_uScm, color = plot)) +
    geom_point()+
    geom_vline(xintercept = tmp1_doc_high,linetype="dashed",color="black")+ 
    geom_vline(xintercept = tmp2_doc_high,linetype="dashed",color="black")+ 
    geom_vline(xintercept = tmp3_doc_high,linetype="dashed",color="black")+ 
     annotate("rect", xmin= min(pw_ec$datetime_est) , xmax = max(pw_ec$datetime_est), ymin = (235.3203 - 28.85163), ymax = (235.3203 + 28.85163), fill= "grey", alpha = 0.2)+
    ylab("mean Porewater Conductivity (uS/cm)")+
    xlab("Date")+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()




```

pw_ec_all %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = datetime_est, y = EC_uScm, color = plot)) +
    geom_point()+
    geom_vline(xintercept = WaterDeliveryStart2022,linetype="dashed",color="grey")+ 
    geom_vline(xintercept = WaterDeliveryStart2,linetype="dashed",color="grey")+ 
    geom_vline(xintercept = WaterDeliveryStart1,linetype="dashed",color="grey")+ 
    ylab("mean Porewater Conductivity (uS/cm)")+
    xlab("Date")+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()



#Supplemental Figure 2: Column Experiment DOC

```{r load col ex}
# exp_doc <- read_csv("~/GitHub/tempest_ionic_strength/Data/Processed Data/DOC/TEMPEST_ColEx_NPOC_TDN_meansforCDOM.csv") %>%
#   mutate(Wash = case_when(is.na(Wash) ~ 0,
#                           TRUE ~ Wash),
#     Treatment = case_when(Treatment == "AS" ~ "Artificial Saltwater",
#                                Treatment == "FW" ~ "Artificial Freshwater",
#                                Treatment == "Control" ~ "Control",
#                                Treatment == "SWE" ~ "Artificial Saltwater",
#                                Treatment == "CL" ~ "Artificial Freshwater",
#                           TRUE ~ Treatment)) %>%
#   mutate(Treatment = factor(Treatment, levels = Treatment_order))

exp_doc <- read_csv("~/GitHub/tempest_ionic_strength/Data/Exp 3 Column Experiment ColEx Data/TEMPEST_Col_Ex_all_chem_averaged_L2.csv") %>%
  mutate(Wash = case_when(is.na(Timepoint) ~ 0,
                          TRUE ~ Timepoint),
    Treatment = case_when(Treatment == "SW" ~ "Artificial Saltwater",
                               Treatment == "FW" ~ "Artificial Freshwater",
                               Treatment == "Control" ~ "Control",
                               Treatment == "SWE" ~ "Artificial Saltwater",
                               Treatment == "CL" ~ "Artificial Freshwater",
                          TRUE ~ Treatment)) %>%
  mutate(Treatment = factor(Treatment, levels = Treatment_order))
```

```{r col ex}
SI2 <- exp_doc %>%
   ggplot(aes(x = Wash, y = doc_mgperL_avg, color = Treatment)) +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mgperL_avg-doc_mgperL_sd, ymax=doc_mgperL_avg+doc_mgperL_sd), shape=17, size = 1)+
    ylab("DOC mg/L")+
    xlab("Treatment Wash Number")+
   # title("Field Manipulation") +
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  Column Experiment"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(SI2)
```


# Supplmental Figure 3: pH 

```{r load pH data}

ph <-  read_rds("~/GitHub/TEMPEST-1-porewater/data/averaged_ph_porewater.rds") 

```


```{r pH timeseries}
ph %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = datetime, y = pH, color = plot)) +
    geom_vline(xintercept = WaterDeliveryStart1,linetype="dashed",color="grey")+ 
    geom_vline(xintercept = WaterDeliveryStart2,linetype="dashed",color="grey")+
    geom_vline(xintercept = WaterDeliveryStart1_2024,linetype="dashed",color="grey")+
   geom_vline(xintercept = WaterDeliveryStart2_2024,linetype="dashed",color="grey")+
   geom_vline(xintercept = WaterDeliveryStart3_2024,linetype="dashed",color="grey")+
   geom_pointrange(aes(ymin=pH-pH_sd, ymax=pH+pH_sd), shape=17, size=1)+
   annotate("rect", xmin= min(ph$datetime) , xmax = max(ph$datetime), ymin = (6.576429 - 0.4931158), ymax = (6.576429 + 0.4931158), fill= "blue", alpha = 0.2)+
    ylab("porewater pH")+
    xlab("Date")+
    ylim(0,10)+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()
```

## SI 3 Option 2 pH vs DOC

```{r load ph year plot avgs}
ph_2 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_ph_porewater_TMP2.rds") %>%
  mutate(pH_sd = case_when(is.na(pH_sd) ~ 0,
                          TRUE ~ pH_sd))

ph_3 <- readRDS("~/GitHub/TEMPEST-1-porewater/data/averaged_ph_porewater_TMP3.rds") %>%
  mutate(pH_sd = case_when(is.na(pH_sd) ~ 0,
                          TRUE ~ pH_sd))

ph_2n3 <- ph_2 %>%
  full_join(ph_3)
```

```{r merging}

ph_doc <- pw_doc_2 %>%
  full_join(pw_doc_3) %>%
  full_join(ph_2n3, by = c("adj_month", "year", "plot")) %>%
  select(-tdn_mg_l_mean, -tdn_mg_l_sd) %>%
  drop_na()
```

```{r plot ph vs doc }

ph_doc %>%
   filter(plot == "Saltwater") %>%
    ggplot(aes(x = pH_mean, y = doc_mg_l_mean)) +
    geom_pointrange(aes(xmin=pH_mean-pH_sd, xmax=pH_mean+pH_sd, ymin= doc_mg_l_mean - doc_mg_l_sd, ymax = doc_mg_l_mean + doc_mg_l_sd), shape=17)+ 
    ylab("porewater DOC")+
    xlab("porewater pH")+
    theme_classic()
```


# Supplemental Figure 4: ICP data from column experiment 

```{r icp column}

exp_col_dat <- read_csv("~/GitHub/tempest_ionic_strength/Data/Exp 3 Column Experiment ColEx Data/TEMPEST_Col_Ex_all_chem_averaged_L2.csv") %>%
  mutate(Wash = case_when(is.na(Timepoint) ~ 0,
                          TRUE ~ Timepoint),
    Treatment = case_when(Treatment == "SW" ~ "Artificial Saltwater",
                               Treatment == "FW" ~ "Artificial Freshwater",
                               Treatment == "Control" ~ "Control",
                               Treatment == "SWE" ~ "Artificial Saltwater",
                               Treatment == "CL" ~ "Artificial Freshwater",
                          TRUE ~ Treatment)) %>%
  mutate(Treatment = factor(Treatment, levels = Treatment_order))


```
Fe, Al, Ca
```{r icp col ex}
SI4_a <- exp_col_dat %>%
  filter(Treatment == "Artificial Saltwater")  %>%
   ggplot(aes(x = Wash, y = Fe_ppm_avg), color = "#C38D18") +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=Fe_ppm_avg-Fe_ppm_sd, ymax=Fe_ppm_avg+Fe_ppm_sd), shape=17, size = 1, color = "#C38D18")+
    ylab("Fe mg/L")+
    xlab(" ")+
   # title("Field Manipulation") +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  Column Experiment"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(SI4_a)

SI4_b <- exp_col_dat %>%
  filter(Treatment == "Artificial Saltwater")  %>%
   ggplot(aes(x = Wash, y = Al_ppm_avg), color = "#DCDCDA") +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=Al_ppm_avg-Al_ppm_sd, ymax=Al_ppm_avg+Al_ppm_sd), shape=17, size = 1, color = "#DCDCDA")+
    ylab("Al mg/L")+
    xlab(" ")+
   # title("Field Manipulation") +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  Column Experiment"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(SI4_b)

SI4_c <- exp_col_dat %>%
  filter(Treatment == "Artificial Saltwater")  %>%
   ggplot(aes(x = Wash, y = Ca_ppm_avg), color = "#009DBD") +
   geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=Ca_ppm_avg-Ca_ppm_sd, ymax=Ca_ppm_avg+Ca_ppm_sd), shape=17, size = 1, color = "#009DBD")+
    ylab("Ca mg/L")+
    xlab("Treatment Wash Number")+
   # title("Field Manipulation") +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    scale_shape_manual(values = c(2, 17))+ 
    theme_classic() +
    theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  Column Experiment"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(SI4_c)

SI4 <- ggarrange(SI4_a, SI4_b, SI4_c, nrow = 3, legend = "bottom", common.legend = TRUE, align = "hv")

SI4

 cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/FigureS4.png",SI4, dpi=600,
                  nrow = 3)
```

# Supplemental Figure 5: ICP data from YR 1 through time 


```{r load ICP data}

icp <-  read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_POOL_ICP_2022_L1.csv") 

```

```{r icp timeseries}
ca <- icp %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = date, y = Ca_mgL_Porewater_average, color = plot)) +
    geom_point(shape=17, size=4)+
    geom_vline(xintercept = as.Date(WaterDeliveryStart2022),linetype="dashed",color="blue")+ 
    geom_vline(xintercept = as.Date("2022-07-21 15:59:00"),linetype="solid",color="grey")+
    ylab("Porewater Ca (mg/L)")+
    xlab("")+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16))

mg <- icp %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = date, y = Mg_mgL_Porewater_average, color = plot)) +
    geom_point(shape=17, size=4)+
    geom_vline(xintercept = as.Date(WaterDeliveryStart2022),linetype="dashed",color="grey")+ 
    ylab("Porewater Mg2+ mg/L")+
    xlab("Date")+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()

na <- icp %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = date, y = Na_mgL_Porewater_average, color = plot)) +
    geom_point(shape=17, size=4)+
    geom_vline(xintercept = as.Date(WaterDeliveryStart2022),linetype="dashed",color="grey")+ 
    ylab("Porewater Na+ mg/L")+
    xlab("Date")+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()

fe <-icp %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = date, y = Fe_mgL_Porewater_average, color = plot)) +
    geom_point(shape=17, size=4)+
    geom_vline(xintercept = as.Date(WaterDeliveryStart2022),linetype="dashed",color="blue")+ 
    geom_vline(xintercept = as.Date("2022-07-21 15:59:00"),linetype="solid",color="grey")+ 
    ylab("Porewater Fe (mg/L)")+
    xlab("")+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))  +
  theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16))

al <- icp %>%
  mutate(plot = factor(plot, levels = plot_order)) %>%
    ggplot(aes(x = date, y = Al_mgL_Porewater_average, color = plot)) +
    geom_point(shape=17, size=4)+
    geom_vline(xintercept = as.Date(WaterDeliveryStart2022),linetype="dashed",color="blue")+ 
    geom_vline(xintercept = as.Date("2022-07-21 15:59:00"),linetype="solid",color="grey")+ 
    ylab("Porewater Al (mg/L)")+
    xlab("Date")+
    scale_color_manual(values=Anyas_colors) +
    theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))  +
  theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1),
          text= element_text(size = 16))

SI5 <- ggarrange(fe, al, ca, nrow = 1, legend = "bottom", common.legend = TRUE, align = "hv")

SI5

 cowplot::save_plot("~/GitHub/TEMPEST-1-porewater/figures/manuscript_figures/FigureS5.png",SI5, dpi=600,
                  nrow = 1)

```
### Supplemental Figure XX: Precip and DOC

```{r precip}
doc_precip <- read_csv("~/GitHub/TEMPEST-1-porewater/data/averaged_doc_porewater_allyrs_withprecip.csv")
precip <- read_csv("~/GitHub/TEMPEST-1-porewater/data/mean_monthly_precip_2006-2026.csv")
```

```{r plots doc rain}
rain_doc <- doc_precip %>%
   ggplot() +
   geom_point(aes(x = stamp, y = doc_mg_l_mean, color= mean_monthly_precip_cm)) +
   facet_wrap(~plot)+
    theme_classic()

rain_doc

precip_study <- precip %>%
  filter(year > 2021)

   ggplot() +
   geom_col(aes(x = stamp, y =  mean_monthly_precip_cm * 100), data = precip_study) +
   geom_point(aes(x = stamp, y = doc_mg_l_mean, color = plot), data = doc_precip)+
         scale_color_manual(values=Anyas_colors) 



```


#graveyard 
```{r graveyard}
Figure1d <- exp_doc %>%
  mutate(Treatment = factor(Treatment, levels = Treatment_order)) %>%
    mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd)) %>%
    ggplot(aes(x = Wash, y = doc_mg_l_mean, color = Treatment)) +
    geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=16, size = 1)+
    ylab("DOC mgC/L")+
    xlab("Experimental Wash")+
   # title("Laboratory Simulation")+
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    theme_classic() +
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  d) Laboratory"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

print(Figure1d)

Figure1d_LEGEND <- exp_doc %>%
  mutate(Treatment = factor(Treatment, levels = Treatment_order)) %>%
    mutate(doc_mg_l_sd = case_when(is.na(doc_mg_l_sd) ~ 0,
                          TRUE ~ doc_mg_l_sd)) %>%
    ggplot(aes(x = Wash, y = doc_mg_l_mean, color = Treatment)) +
    geom_vline(xintercept = 1,linetype="dashed",color="darkgrey")+ 
   geom_pointrange(aes(ymin=doc_mg_l_mean-doc_mg_l_sd, ymax=doc_mg_l_mean+doc_mg_l_sd), shape=16, size = 1)+
    ylab("DOC mgC/L")+
    xlab("Experimental Wash")+
   # title("Laboratory Simulation")+
    ylim(-10,100)+
    scale_color_manual(values=Anyas_colors) +
    scale_x_continuous(breaks = pretty_breaks(), limits= c(0,12)) +
    theme_classic() +
    theme(legend.position = "bottom", panel.border = element_rect(color = "black", fill = NA, size = 1)) +
  geom_text(data = data.frame(x = -Inf, y = Inf, label = "  d) Laboratory"), aes(x, y, label = label), hjust = -0.1, vjust = 2, inherit.aes = FALSE)

fig1c_legend = get_legend(Figure1d_LEGEND)
```


