---
title: "Conductivity"
author: "AMP"
date: "`r Sys.Date()`"
output: html_document
---

This script brings together the teros conductivity and the lysimeter DOC data to enable calculation of relationships between the two for manuscript. 

NOTE: make sure your computer is in EST, because R does what it wants with time zones: 

# 1) Set up environment

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(pacman)
pacman::p_load(tidyverse, # keep things tidy
               googlesheets4, # read_sheet 
               googledrive, # drive_ functions
               here, #helps with file pathing 
               lubridate) #those pesky dates

#double check your wd
getwd()

#clean out your environment 
rm(list = ls())

#Authenticate Google Drive Access:
drive_auth()

Sys.setenv(TZ = "America/New_York")
```

### Set the study dates

```{r event and study dates}
endstudydate = lubridate::as_date("2024-01-31")
startstudydate = lubridate::as_date("2022-05-01")

endstudydate_yr1 = lubridate::as_date("2023-05-31")
startstudydate_yr1 = lubridate::as_date("2022-05-01")

year1_start = lubridate::as_date("2022-01-01")
year1_stop = lubridate::as_date("2022-12-31")

year2_start = lubridate::as_date("2023-01-01")
year2_stop = lubridate::as_date("2023-12-31")

year3_start = lubridate::as_date("2024-01-01")
year3_stop = lubridate::as_date("2024-12-31")

WaterDeliveryStart2022 = as.POSIXct("2022-06-22 05:30:00", tz = "EST")
WaterDeliveryStop2022 = as.POSIXct("2022-06-22 14:30:00", tz = "EST")

WaterDeliveryStart1 = as.POSIXct("2023-06-06 05:30:00", tz = "EST")
WaterDeliveryStop1 = as.POSIXct("2023-06-06 14:30:00", tz = "EST")

WaterDeliveryStart2 = as.POSIXct("2023-06-07 05:30:00", tz = "EST")
WaterDeliveryStop2 = as.POSIXct("2023-06-07 14:30:00", tz = "EST")

plot_order <- c('Control', 'Freshwater','Saltwater')
time_order <- c('Pre', 'Mid','Post')
```



# 2) bring in soil conductivity

```{r load ec}
ec_l2_15cm = readRDS("~/GitHub/TEMPEST-1-porewater/processed data/TMP_PW_EC_15cm_2022-2023_L2.rds") %>%
  filter(depth == "15") %>%
  mutate(date = as_date(datetime_est),
         datetime_est = as.POSIXct(datetime_est)) %>%
  group_by(plot, grid) %>%
   filter(datetime_est < year2_stop & datetime_est > year1_start)

ec_l2_15cm_salt <- ec_l2_15cm %>%
  filter(plot == "Saltwater")

ec_l2_15cm_control <- ec_l2_15cm %>%
  filter(plot == "Control")

ec_l2_15cm_fresh <- ec_l2_15cm %>%
  filter(plot == "Freshwater")

ec_l2_15cm_dailyplotmeans = readRDS("~/GitHub/TEMPEST-1-porewater/processed data/TMP_PW_EC_15cm_2022-2023_L2.rds") %>%
  filter(depth == "15") %>%
  mutate(date = as_date(datetime_est),
         datetime_est = as.POSIXct(datetime_est)) %>%
  group_by(plot, yday(datetime_est)) %>%
   filter(datetime_est < year2_stop & datetime_est > year1_start) %>%
   summarise(datetime_est = mean(datetime_est),
                  EC_uScm_mean = mean(EC_uScm, na.rm = TRUE),
                  EC_uScm_sd = sd(EC_uScm, na.rm = TRUE),
                  .groups = "drop") %>%
  select(plot, datetime_est, EC_uScm_mean, EC_uScm_sd)
  
```

```{r quick plot of S cond}
ec_l2_15cm %>%
  filter(plot == "Control") %>%
    ggplot(aes(x = datetime_est, y = EC_uScm, color = grid)) +
    geom_point()+
    geom_vline(xintercept = WaterDeliveryStart2022,linetype="dashed",color="grey")+ 
    geom_vline(xintercept = WaterDeliveryStart1,linetype="dashed",color="grey")+ 
    geom_vline(xintercept = WaterDeliveryStart2,linetype="dashed",color="grey")+
    ylab("porewater EC uS/cm")+
    xlab("Date")+
    theme_classic()

```


# 4) Load DOC data 

note these will be on google drive eventually but for now just in github:

```{r import doc}
#Porewater DOC data, from L1 google drive: 

doc_l1_2022 <- read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_GRIDSONLY_NPOCTDN_2022_L1.csv")
  
doc_l1_2023 <- read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_GRIDSONLY_NPOCTDN_2023_L1.csv")

doc_l1_pw_grids <- doc_l1_2022 %>%
  full_join(doc_l1_2023) %>%
  select(plot, grid, evacuation_datetime, collection_datetime, doc_mg_l, tdn_mg_l) %>%
 # mutate(date = as_date(collection_datetime)) %>%
  drop_na() %>%
  distinct()#This is dropping the EMPTY lysimeters: and don't need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
```

# 5) filter and summarize cond by date time lysimeter was collecting FOR DOC data:

```{r filter and summarize, message= FALSE}
filter_and_summarize <- function(df, start_date, end_date) {
  df %>%
    filter(datetime_est >= start_date, datetime_est <= end_date) %>%
    summarise(mean_EC_uScm = mean(EC_uScm),
              min_EC_uScm = min(EC_uScm),
              max_EC_uScm = max(EC_uScm),
              .groups = "drop") 
}

ec_during_collection_15cm_salt = doc_l1_pw_grids %>%
  filter(plot == "Saltwater") %>%
  group_by(plot, grid, evacuation_datetime, collection_datetime) %>%
  summarize(mean_EC_uScm_15cm = mean(ec_l2_15cm_salt$EC_uScm[ec_l2_15cm_salt$datetime_est >= first(evacuation_datetime) & ec_l2_15cm_salt$datetime_est <= first(collection_datetime)]),
            min_EC_uScm_15cm = min(ec_l2_15cm_salt$EC_uScm[ec_l2_15cm_salt$datetime_est >= first(evacuation_datetime) & ec_l2_15cm_salt$datetime_est <= first(collection_datetime)]),
            max_EC_uScm_15cm = max(ec_l2_15cm_salt$EC_uScm[ec_l2_15cm_salt$datetime_est >= first(evacuation_datetime) & ec_l2_15cm_salt$datetime_est <= first(collection_datetime)])) %>%
  ungroup()

doc_l1_pw_grids_salt <- doc_l1_pw_grids %>%
  filter(plot == "Saltwater")
  
ec_lys_doc_salt = ec_during_collection_15cm_salt  %>%
  right_join(doc_l1_pw_grids_salt, by=c("plot", "grid", "evacuation_datetime", "collection_datetime")) %>%
  distinct() %>%
  mutate(collection_date= as_date(collection_datetime)) 


#control

ec_during_collection_15cm_control = doc_l1_pw_grids %>%
  filter(plot == "Control") %>%
  group_by(plot, grid, evacuation_datetime, collection_datetime) %>%
  summarize(mean_EC_uScm_15cm = mean(ec_l2_15cm_control$EC_uScm[ec_l2_15cm_control$datetime_est >= first(evacuation_datetime) & ec_l2_15cm_control$datetime_est <= first(collection_datetime)]),
            min_EC_uScm_15cm = min(ec_l2_15cm_control$EC_uScm[ec_l2_15cm_control$datetime_est >= first(evacuation_datetime) & ec_l2_15cm_control$datetime_est <= first(collection_datetime)]),
            max_EC_uScm_15cm = max(ec_l2_15cm_control$EC_uScm[ec_l2_15cm_control$datetime_est >= first(evacuation_datetime) & ec_l2_15cm_control$datetime_est <= first(collection_datetime)])) %>%
  ungroup()

doc_l1_pw_grids_control <- doc_l1_pw_grids %>%
  filter(plot == "Control")
  
ec_lys_doc_control = ec_during_collection_15cm_control  %>%
  right_join(doc_l1_pw_grids_control, by=c("plot", "grid", "evacuation_datetime", "collection_datetime")) %>%
  distinct() %>%
  mutate(collection_date= as_date(collection_datetime)) 
```
```{r quick plot of S cond during lysimeter time}
ec_during_collection_15cm_salt %>%
  filter(plot == "Saltwater") %>%
    ggplot(aes(x = collection_datetime, y = mean_EC_uScm_15cm, color = grid)) +
    geom_point()+
    geom_vline(xintercept = WaterDeliveryStart2022,linetype="dashed",color="grey")+ 
    geom_vline(xintercept = WaterDeliveryStart1,linetype="dashed",color="grey")+ 
    geom_vline(xintercept = WaterDeliveryStart2,linetype="dashed",color="grey")+
    ylab("porewater EC uS/cm")+
    xlab("Date")+
    theme_classic()

ec_during_collection_15cm_control %>%
  filter(plot == "Control") %>%
    ggplot(aes(x = collection_datetime, y = max_EC_uScm_15cm, color = grid)) +
    geom_jitter(width = 15, height = 15) +
    geom_vline(xintercept = WaterDeliveryStart2022,linetype="dashed",color="grey")+ 
    geom_vline(xintercept = WaterDeliveryStart1,linetype="dashed",color="grey")+ 
    geom_vline(xintercept = WaterDeliveryStart2,linetype="dashed",color="grey")+
    ylab("max porewater EC uS/cm")+
    xlab("Date")+
    theme_classic()
```

```{r writeout}

write_csv(ec_lys_doc_salt, "~/GitHub/TEMPEST-1-porewater/processed data/for_fluxes/Lysimeter_EC_DOC_forcorrelations_SALTWATER.csv")

write_csv(ec_lys_doc_control, "~/GitHub/TEMPEST-1-porewater/processed data/for_fluxes/Lysimeter_EC_DOC_forcorrelations_CONTROL.csv")

write_csv(ec_l2_15cm_dailyplotmeans, "~/GitHub/TEMPEST-1-porewater/processed data/for_fluxes/Lysimeter_EC_daily_plot_means.csv")
```


