---
title: "Water Fluxes"
author: "AMP"
date: "`r Sys.Date()`"
output: html_document
---

This script brings together the lysimeter volume data, teros soil moisture, and the lysimeter DOC data to enable calculation of water fluxes for manuscript. 

NOTE: make sure your computer is in EST, because R does what it wants with time zones: 

# 1) Set up environment

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(pacman)
pacman::p_load(tidyverse, # keep things tidy
               googlesheets4, # read_sheet 
               googledrive, # drive_ functions
               here, #helps with file pathing 
               lubridate) #those pesky dates

#double check your wd
getwd()

#clean out your environment 
rm(list = ls())

#Authenticate Google Drive Access:
drive_auth()

Sys.setenv(TZ = "America/New_York")
```

### Set the study dates

```{r event and study dates}
endstudydate = lubridate::as_date("2024-01-31")
startstudydate = lubridate::as_date("2022-05-01")

endstudydate_yr1 = lubridate::as_date("2023-05-31")
startstudydate_yr1 = lubridate::as_date("2022-05-01")

year1_start = lubridate::as_date("2022-01-01")
year1_stop = lubridate::as_date("2022-12-31")

year2_start = lubridate::as_date("2023-01-01")
year2_stop = lubridate::as_date("2023-12-31")

year3_start = lubridate::as_date("2024-01-01")
year3_stop = lubridate::as_date("2024-12-31")

WaterDeliveryStart2022 = as.POSIXct("2022-06-22 05:30:00", tz = "EST")
WaterDeliveryStop2022 = as.POSIXct("2022-06-22 14:30:00", tz = "EST")

WaterDeliveryStart1 = as.POSIXct("2023-06-06 05:30:00", tz = "EST")
WaterDeliveryStop1 = as.POSIXct("2023-06-06 14:30:00", tz = "EST")

WaterDeliveryStart2 = as.POSIXct("2023-06-07 05:30:00", tz = "EST")
WaterDeliveryStop2 = as.POSIXct("2023-06-07 14:30:00", tz = "EST")

plot_order <- c('Control', 'Freshwater','Estuarine-water')
time_order <- c('Pre', 'Mid','Post')
```

# 2) Bring in the cleaned lysimeter volumes metadata

```{r lys volumes}

source("~/GitHub/TEMPEST-1-porewater/scripts/processing_scripts/lysimeter_volumes_forfluxes.R")

#filter these to reasonable range, i.e. you need at least 7 mL for DOC:

volumes_grids_pooled <- volumes_grids_pooled %>%
  filter(Total_Volume_mL > 7) %>%
  distinct(plot, grid, collection_date, evacuation_date, .keep_all = TRUE) 

```

# 3) bring in moisture content

```{r load vwc}
vwc_l2_15cm = readRDS("~/GitHub/TEMPEST-1-porewater/processed data/TMP_PW_VWC_2022-2023_L2.rds") %>%
  filter(depth == "15") %>%
  mutate(date = as_date(datetime_est)) %>%
  group_by(plot, grid) %>%
   filter(datetime_est < year2_stop & datetime_est > year1_start)

vwc_l2_5cm = readRDS("~/GitHub/TEMPEST-1-porewater/processed data/TMP_PW_VWC_2022-2023_L2.rds") %>%
  filter(depth == "5") %>%
  mutate(date = as_date(datetime_est)) %>%
  group_by(plot, grid) %>%
   filter(datetime_est < year2_stop & datetime_est > year1_start)


```

# 4) Load DOC data 

note these will be on google drive eventually but for now just in github:

```{r import doc}
#Porewater DOC data, from L1 google drive: 

doc_l1_2022 <- read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_GRIDSONLY_NPOCTDN_2022_L1.csv")
  
doc_l1_2023 <- read_csv("~/GitHub/TEMPEST-1-porewater/processed data/for_google_drive/TMP_PW_GRIDSONLY_NPOCTDN_2023_L1.csv")

doc_l1_pw_grids <- doc_l1_2022 %>%
  full_join(doc_l1_2023) %>%
  select(plot, grid, evacuation_datetime, collection_datetime, doc_mg_l, tdn_mg_l) %>%
 # mutate(date = as_date(collection_datetime)) %>%
  drop_na() %>%
  distinct()#This is dropping the EMPTY lysimeters: and don't need to specify column here because the only ones with NAs are DOC or TDN and they are both NA at the same time.
```

# 5) filter and summarize VWC by date time lysimeter was collecting FOR DOC data:

```{r filter and summarize, message= FALSE}
filter_and_summarize <- function(df, start_date, end_date) {
  df %>%
    filter(datetime_est >= start_date, datetime_est <= end_date) %>%
    summarise(mean_vwc_m3m3 = mean(vwc_m3m3),
              min_vwc_m3m3 = min(vwc_m3m3),
              max_vwc_m3m3 = max(vwc_m3m3),
              .groups = "drop") 
}

vwc_during_collection_15cm = doc_l1_pw_grids %>%
  group_by(plot, grid, evacuation_datetime, collection_datetime) %>%
  summarize(mean_vwc_m3m3_15cm = mean(vwc_l2_15cm$vwc_m3m3[vwc_l2_15cm$datetime_est >= first(evacuation_datetime) & vwc_l2_15cm$datetime_est <= first(collection_datetime)]),
            min_vwc_m3m3_15cm = min(vwc_l2_15cm$vwc_m3m3[vwc_l2_15cm$datetime_est >= first(evacuation_datetime) & vwc_l2_15cm$datetime_est <= first(collection_datetime)]),
            max_vwc_m3m3_15cm = max(vwc_l2_15cm$vwc_m3m3[vwc_l2_15cm$datetime_est >= first(evacuation_datetime) & vwc_l2_15cm$datetime_est <= first(collection_datetime)])) %>%
  ungroup()

vwc_lys_doc = vwc_during_collection_15cm  %>%
  right_join(doc_l1_pw_grids, by=c("plot", "grid", "evacuation_datetime", "collection_datetime")) %>%
  distinct() %>%
  mutate(collection_date= as_date(collection_datetime)) 
```


```{r merging}

vwc_lys_vol_doc = vwc_lys_doc %>%
  left_join(volumes_grids_pooled, by = c("plot", "grid", "collection_date"), relationship = "many-to-many") %>%
  distinct() %>%
  # There are some samples we have DOC for that WE STILL DON'T HAVE THE RIGHT METADATA FOR, namely missing the volume data. So, we're going to just have to take the mean of the other volumes at that date time, because, there is no way to track it back further. 
  mutate(evacuation_date = case_when(is.na(evacuation_date) ~ as_date(evacuation_datetime),
                                     TRUE ~ evacuation_date)) %>%
  group_by(plot, evacuation_date) %>%
  mutate(Total_Volume_mL = ifelse(is.na(Total_Volume_mL), mean(Total_Volume_mL, na.rm = TRUE), Total_Volume_mL)) %>%
  ungroup() %>%
  #when that gives NAs back due to still not having the right metadata, taking mean at that date across plot:
  group_by(collection_date) %>%
  mutate(Total_Volume_mL = ifelse(is.na(Total_Volume_mL), mean(Total_Volume_mL, na.rm = TRUE), Total_Volume_mL)) %>%
  ungroup() %>%
  #when youre still getting nas for this bad metadata tracking of volumes, now just assume its 40mL: 
  mutate(Total_Volume_mL = ifelse(is.na(Total_Volume_mL), 40, Total_Volume_mL)) %>%
  select(-collection_date, -evacuation_date) %>%
  distinct()

write_csv(vwc_lys_vol_doc, "~/GitHub/TEMPEST-1-porewater/processed data/for_fluxes/Lysimeter_Vols_VWC_DOC_forfluxes.csv")
```



note, there are no 5 cm lysimeters at all of the grid locations, only 15 everywhere. So we are going to focus on 15cm only, but here is the code to bring in the 5 cm, if we want/need. 

vwc_during_collection_5cm = lysimeter_vols %>%
  group_by(plot, grid, evacuation_datetime, collection_datetime) %>%
  rowwise() %>%
  do(data.frame(filter_and_summarize(vwc_l2_5cm, .$evacuation_datetime, .$collection_datetime),
       evacuation_datetime = .$evacuation_datetime,
    collection_datetime = .$collection_datetime)) %>%
  ungroup() %>%
  rename(mean_vwc_m3m3_5cm = mean_vwc_m3m3,
         min_vwc_m3m3_5cm = min_vwc_m3m3,
         max_vwc_m3m3_5cm =max_vwc_m3m3 ) %>%
   group_by(plot, grid, evacuation_datetime, collection_datetime) 

vwc_during_collection_all = vwc_during_collection_15cm %>%
  full_join(vwc_during_collection_5cm, by=c("plot","grid","evacuation_datetime", "collection_datetime"))


<!-- # ```{r import lysimeter data} -->
<!-- inventory_directory <- "https://docs.google.com/spreadsheets/d/1sFWq-WKhemPzbOFInqhCu_Lx0lsO6a_Z/edit#gid=496164093" -->

<!-- directory= file.path(here() %>% dirname(), 'TEMPEST-1-Porewater/data_do_not_commit/water fluxes') -->

<!-- file_path = file.path(directory,"porewaterinventory.xlsx") -->

<!-- drive_download(inventory_directory, path= file_path, overwrite = TRUE) -->

<!-- pooled_vols <- readxl::read_excel(file_path, skip=4, sheet="Porewater - Pooled") %>% -->
<!--   select(Sample_ID, Total_Volume_mL, Grid_C3:Grid_I5, Evacuation_date_YYYMMDD, Collection_Date_YYYYMMDD) %>% -->
<!--   mutate(across(starts_with("Grid"), ~ ifelse(is.na(.),0,.))) %>% -->
<!--   mutate(Plot = stringr::str_extract(Sample_ID, 'FW|SW|C')) %>% -->
<!--    pivot_longer(cols = c("Grid_C3", "Grid_C6", "Grid_F4", "Grid_H3", "Grid_H6", "Grid_B4", "Grid_D5", "Grid_E3", "Grid_F6", "Grid_I5"), -->
<!--     names_to = "grid", -->
<!--     values_to = "Volume_mL") %>% -->
<!--    mutate(plot = case_when(Plot == "C" ~ "Control", -->
<!--                            Plot == "SW" ~ "Estuarine-water", -->
<!--                            Plot == "FW" ~ "Freshwater", -->
<!--                            Plot == "Saltwater" ~ "Estuarine-water", -->
<!--                           TRUE ~ Plot), -->
<!--           grid = str_remove(grid, "Grid_")) %>% -->
<!--   rename(evacuation_date = Evacuation_date_YYYMMDD, -->
<!--          collection_date = Collection_Date_YYYYMMDD) %>% -->
<!--   mutate(evacuation_date = lubridate::as_date(as.character(evacuation_date), format = "%Y%m%d"), -->
<!--          collection_date = lubridate::as_date(as.character(collection_date), format = "%Y%m%d"), -->
<!--          date = stringr::str_extract(Sample_ID, "[0-9]{8}")) %>% -->
<!--   mutate(date = lubridate::as_date(date, format = "%Y%m%d")) %>% -->
<!--   group_by(date) %>% -->
<!--    #manual entry of missing evacuation information from the notes:  -->
<!--   mutate(evacuation_date = case_when( -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2022-05-18") ~ as.Date("2022-05-12"), -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2022-05-20")  ~ as.Date("2022-05-12"),  -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2022-06-13")  ~ as.Date("2022-06-09"), -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2022-06-15")  ~ as.Date("2022-06-09"), -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2022-06-20") ~ as.Date("2022-06-18"), -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2022-06-22") ~ as.Date("2022-06-22"), -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2022-06-23") ~ as.Date("2022-06-22"), -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2022-06-24") ~ as.Date("2022-06-23"), -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2023-06-05") ~ as.Date("2023-06-03"), -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2023-10-05") ~ as.Date("2023-10-02"), -->
<!--     TRUE ~ evacuation_date)) %>% -->
<!--   fill(evacuation_date, collection_date) %>% -->
<!--   #Empty dates for the event dates, those are when collection and evacuation are usually the same date:  -->
<!--   mutate(collection_date = case_when(is.na(collection_date) ~ date, -->
<!--                                      TRUE ~ collection_date), -->
<!--          evacuation_date = case_when(is.na(evacuation_date) ~ date, -->
<!--                                      TRUE ~ evacuation_date)) %>% -->
<!--   ungroup() %>% -->
<!--   select(Sample_ID,plot, grid, Volume_mL, evacuation_date, collection_date) %>% -->
<!--   filter(Volume_mL > 0) -->

<!-- lysimeter_vols = readxl::read_excel(file_path, skip=3, sheet="Porewater - Individual") %>% -->
<!--   select(Sample_ID, Plot, Volume_mL, Evacuation_date_YYYMMDD, Collection_Date_YYYYMMDD, Collection_Start_Time_24hrs, Collection_End_Time_24hrs, EST_EDT) %>% -->
<!--     rename(evacuation_date = Evacuation_date_YYYMMDD, -->
<!--          collection_date = Collection_Date_YYYYMMDD, -->
<!--          plot = Plot) %>% -->
<!--   mutate(evacuation_date = lubridate::as_date(as.character(evacuation_date), format = "%Y%m%d"), -->
<!--          collection_date = lubridate::as_date(as.character(collection_date), format = "%Y%m%d")) %>% -->
<!--     mutate(grid = stringr::str_extract(Sample_ID, "B4|C3|C6|D5|E3|F4|F6|H3|H6|I5")) %>% -->
<!--   mutate(evacuation_date = case_when( -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2022-05-18") ~ as.Date("2022-05-12"), -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2022-05-20")  ~ as.Date("2022-05-12"),  -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2022-06-13")  ~ as.Date("2022-06-09"), -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2022-06-15")  ~ as.Date("2022-06-09"), -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2022-06-20") ~ as.Date("2022-06-18"), -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2022-06-22") ~ as.Date("2022-06-22"), -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2022-06-23") ~ as.Date("2022-06-22"), -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2022-06-24") ~ as.Date("2022-06-23"), -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2023-06-05") ~ as.Date("2023-06-03"), -->
<!--     is.na(evacuation_date) & str_detect(collection_date, "2023-10-05") ~ as.Date("2023-10-02"), -->
<!--     TRUE ~ evacuation_date)) %>%  -->
<!--     fill(evacuation_date, collection_date) %>% -->
<!--     mutate(plot = case_when(plot == "C" ~ "Control", -->
<!--                            plot == "SW" ~ "Estuarine-water", -->
<!--                            plot == "FW" ~ "Freshwater", -->
<!--                            plot == "Contol" ~ "Control", -->
<!--                            plot == "Saltwater" ~ "Estuarine-water", -->
<!--                           TRUE ~ plot)) %>% -->
<!--   group_by(plot, grid, evacuation_date, collection_date) %>% -->
<!--   #fill down exact group matches by group missing values: -->
<!--   fill(evacuation_date, collection_date, Collection_Start_Time_24hrs, Collection_End_Time_24hrs) %>% -->
<!--   ungroup() %>% -->
<!--   #fill down and up by evacuation date match only:  -->
<!--     group_by(plot, grid, evacuation_date) %>% -->
<!--     fill(evacuation_date, collection_date, Collection_Start_Time_24hrs, Collection_End_Time_24hrs, .direction ="downup") %>% -->
<!--   ungroup() %>% -->
<!--   mutate(Collection_Start_Time_24hrs = str_replace(Collection_Start_Time_24hrs, '\\d+', function(m) str_pad(m, 4, pad = '0', side = ("left"))), -->
<!--          Collection_End_Time_24hrs = str_replace(Collection_End_Time_24hrs, '\\d+', function(m) str_pad(m, 4, pad = '0', side = ("left")))) %>% -->
<!--   full_join(pooled_vols, by = join_by(Sample_ID, Volume_mL, evacuation_date, collection_date, grid, plot)) %>% -->
<!-- #fill down and up by evacuation date match only, once joined with the pooled volumes:  -->
<!--     group_by(plot, grid, evacuation_date) %>% -->
<!--     fill(evacuation_date, collection_date, Collection_Start_Time_24hrs, Collection_End_Time_24hrs, .direction ="downup") %>% -->
<!--   ungroup() %>% -->
<!--     #NAs present because some groups have no time recorded at all. Enter in a set time for those: -->
<!--     group_by(plot, grid, evacuation_date, collection_date) %>% -->
<!--     mutate(Collection_Start_Time_24hrs = case_when(is.na(Collection_Start_Time_24hrs) ~ "1159",  -->
<!--                           TRUE ~ Collection_Start_Time_24hrs), -->
<!--            Collection_End_Time_24hrs = case_when(is.na(Collection_End_Time_24hrs) ~ "1159",  -->
<!--                           TRUE ~ Collection_End_Time_24hrs)) %>% #if no time recorded in the metadata sheet, fill in with noon -->
<!--   mutate(Collection_Start_Time_24hrs = strptime(Collection_Start_Time_24hrs, format ="%H%M"), -->
<!--          Collection_Start_Time_24hrs = strftime(Collection_Start_Time_24hrs, "%H:%M:%S"), -->
<!--          Collection_End_Time_24hrs = strptime(Collection_End_Time_24hrs, format ="%H%M"), -->
<!--          Collection_End_Time_24hrs = strftime(Collection_End_Time_24hrs, "%H:%M:%S")) %>% -->
<!--   ungroup() %>% -->
<!--     mutate(evacuation_datetime = as.POSIXct(paste(evacuation_date, Collection_Start_Time_24hrs), format = "%Y-%m-%d %H:%M:%S"), -->
<!--          collection_datetime = as.POSIXct(paste(collection_date, Collection_End_Time_24hrs), format = "%Y-%m-%d %H:%M:%S"), -->
<!--   elapsed_time = collection_datetime - evacuation_datetime) %>% -->
<!--   rename(tz = EST_EDT) %>% -->
<!--    mutate(set_tz = with_tz(collection_datetime, tzone = "America/New_York"), -->
<!--          is_dst = dst(set_tz), -->
<!--     tz = case_when(is.na(tz) &  is_dst == TRUE ~ "EDT", -->
<!--                    is.na(tz) &  is_dst != TRUE ~ "EST", -->
<!--                    TRUE ~ tz)) %>% -->
<!--   select(-set_tz, -is_dst) %>% -->
<!--   group_by(plot, grid, evacuation_date, evacuation_datetime, collection_datetime, elapsed_time, tz) %>% -->
<!--   dplyr::summarise(grid_vols_mL = sum(Volume_mL)) %>% -->
<!--   ungroup() %>% -->
<!--    filter(collection_datetime < year2_stop & collection_datetime > year1_start) -->
<!-- ``` -->

<!-- # ```{r filter and sum, message = FALSE} -->
<!-- # # Function to filter and summarize data for a given collection period  -->
<!-- # filter_and_summarize <- function(df, start_date, end_date) { -->
<!-- #   df %>% -->
<!-- #     filter(datetime_est >= start_date, datetime_est <= end_date) %>% -->
<!-- #     summarise(mean_vwc_m3m3 = mean(vwc_m3m3), -->
<!-- #               min_vwc_m3m3 = min(vwc_m3m3), -->
<!-- #               max_vwc_m3m3 = max(vwc_m3m3), -->
<!-- #               .groups = "keep")  -->
<!-- # } -->
<!-- #  -->
<!-- # vwc_during_collection_15cm = doc_l1_pw_grids %>% -->
<!-- #   group_by(plot, grid, evacuation_datetime, collection_datetime) %>% -->
<!-- #   rowwise() %>% -->
<!-- #   do(data.frame(filter_and_summarize(vwc_l2_15cm, .$evacuation_datetime, .$collection_datetime), -->
<!-- #        evacuation_datetime = .$evacuation_datetime, -->
<!-- #     collection_datetime = .$collection_datetime)) %>% -->
<!-- #   ungroup() %>% -->
<!-- #   rename(mean_vwc_m3m3_15cm = mean_vwc_m3m3, -->
<!-- #          min_vwc_m3m3_15cm = min_vwc_m3m3, -->
<!-- #          max_vwc_m3m3_15cm = max_vwc_m3m3 ) %>% -->
<!-- #     group_by(plot, grid, evacuation_datetime, collection_datetime)  -->
<!-- #  -->
<!-- # # No VWC data for I5 in SW plot between 2023-01-30 to 2023-05-31 due to sensor malfunction  -->
<!-- #  -->
<!-- # vwc_lys_vols = vwc_during_collection_15cm  %>% -->
<!-- #   right_join(doc_l1_pw_grids, by=c("plot", "grid", "evacuation_datetime", "collection_datetime")) %>% -->
<!-- #   distinct() -->
<!-- #  -->
<!-- # ``` -->
<!-- #  -->

