---
title: "TEROS data prep"
author: "PR Modified by AMP"
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: true
knit: "rmarkdown::render"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      message = F,
                      warning = F)

common_tz = "Etc/GMT+5"
```

### General purpose

R markdown for the TEROS data for the project. 

```{r Setup}
## First, load setup script to set up environment
pacman::p_load(tidyverse, 
       #rdrop2, 
       cowplot, 
       lubridate, 
       #ggtukey, # adding letters to boxplots
      # multcompView, # compact letter display for comparing boxplots
       plotly,
       parsedate, # parse_date()
       janitor, # clean_names()
       purrr, # map()
       furrr,  # future_map()
       tictoc, # time stuff
       sf, 
       hms,
       ggallin,
       ggpubr, #stat_compare_means()
       #metR, # geom_contour_fill
       rnaturalearth,
       ggspatial, # north arrow and compass
       googledrive, 
       ggthemes, #theme_map()
      here) 
```

```{r metadata}
#inventory_directory <- "https://docs.google.com/spreadsheets/d/10u29M5Dbkw54Z2DwkVBptVz6Td0SvSld/edit?gid=90806693#gid=90806693"

directory= file.path(here() %>% dirname(), 'TEMPEST-1-Porewater/data_do_not_commit/')

file_path = file.path(directory,"TEROSinventory.xlsx")

#drive_download(inventory_directory, path= file_path, overwrite = TRUE)

teros_inv <- readxl::read_excel(file_path, sheet="Sheet1") %>%
    select(Plot, 'Grid Square', 'Junction Box ID', ID, Depth) %>% 
    rename(Location_Inventory = "Grid Square",
           sensor_id = ID,
           plot = Plot,
           depth = Depth) %>% 
  mutate(depth = as.character(depth),
         plot = case_when(plot == "Control" ~ "Control", 
                          plot == "Seawater" ~ "Saltwater", 
                          plot == "Freshwater" ~ "Freshwater")) %>%
  drop_na()
```

<!-- ```{r data setup} -->
<!-- raw_data_path <- "./data_do_not_commit/teros/" -->

<!-- #TMP -->
<!-- l1_2022_file_list <- drive_ls("https://drive.google.com/drive/u/0/folders/1su94GOgAeP0fo3Xb3X3DHzxx3PDSJct7",  -->
<!--                       pattern = ".csv") -->

<!-- #TMP  -->
<!-- l1_2023_file_list <- drive_ls("https://drive.google.com/drive/u/0/folders/1zwdiXX7f_wv6WS2sG4vg_OE8ELhNKSdo",  -->
<!--                       pattern = ".csv") -->

<!-- #GCW -->
<!-- l1_2022_weather_file_list <- drive_ls("https://drive.google.com/drive/u/0/folders/1PXTn8vIDn5IQiAjYtpubZKnt8svVkLVR",  -->
<!--                       pattern = ".csv") -->
<!-- #GCW -->
<!-- l1_2023_weather_file_list <- drive_ls("https://drive.google.com/drive/u/0/folders/1p4RL7JStJmlPU1x2MXgcayU2rp2XtFGr",  -->
<!--                       pattern = ".csv") -->



<!-- getwd() -->
<!-- ``` -->

If you need to download the data, this is how you can do it. If not, SKIP this chunk (eval = F). 
This is for downloading from google drive. Now getting from ESS-DIVE

<!-- ```{r Pull L1 data from Gdrive, eval = F} -->

<!-- drive_download_ <- function(data){ -->
<!--   #message(paste("Downloading", data$name)) -->
<!--   # you could add an ifelse to only download files it doesn't fine in raw_data_path -->
<!--   drive_download(data$id, overwrite = T, path = paste0(raw_data_path, data$name)) -->
<!-- } -->

<!-- ## Use a for-loop to read in files in a way that I can see what's going on -->
<!-- ## Download data to local. I tried to map() but for some reasons it doesn't work? -->
<!-- for(i in 1:nrow(l1_2022_file_list)){ -->
<!--   drive_download_(l1_2022_file_list %>% slice(i)) -->
<!-- } -->

<!-- for(i in 1:nrow(l1_2023_file_list)){ -->
<!--   drive_download_(l1_2023_file_list %>% slice(i)) -->
<!-- } -->

<!-- for(i in 1:nrow(l1_2022_weather_file_list)){ -->
<!--   drive_download_(l1_2022_weather_file_list %>% slice(i)) -->
<!-- } -->

<!-- for(i in 1:nrow(l1_2023_weather_file_list)){ -->
<!--   drive_download_(l1_2023_weather_file_list %>% slice(i)) -->
<!-- } -->
<!-- ``` -->

#### Read in Data from ESS-DIVE
Download from https://data.ess-dive.lbl.gov/view/doi%3A10.15485%2F2479200

```{r read in data}

raw_data_path <- "./data_do_not_commit/sensor_data"

site <- "TMP" # change this to e.g. "CRC_W" if you want a particular plot
variable <- "soil_EC_15cm"

# Construct a "regular expression" to find the files we want: in this case,
# CSV files starting with the site code above
regex <- paste0("^", site, ".*csv$")

# Get the names of the files we need. Note this assumes that your
# working directory is the main directory of the L1 data
files <- list.files(path=raw_data_path, pattern = regex, recursive = TRUE, full.names = TRUE)

# Helper function to read the files and filter to just the variable we want
# Use readr::read_csv for easy timestamp handling
# Note that we set "col_types" to force the "Plot" column to be read as a
# character; see https://github.com/COMPASS-DOE/sensor-data-pipeline/issues/186
library(readr)
f <- function(f) {
    message("Reading ", basename(f))
    x <- read_csv(paste0(f), col_types = "ccTccccdccii")
    x[x$research_name == variable,]
}

# Read the files
dat <- lapply(files, f)
dat <- do.call("rbind", dat)
```

### high level cleaning 

```{r clean}
#initial clean

data_clean1 <- dat %>%
     filter(F_OOB == 0, #filters out flag, Flag: out of instrumental bounds (logical; 1=TRUE)
            F_OOS == 0) %>% #filters out flag, Flag: sensor listed as out of service (logical; 1=TRUE)
    mutate(datetime_est = force_tz(TIMESTAMP, tzone = common_tz), 
         depth = str_extract(research_name, "\\d+(?=cm)"),
         plot = case_when(Plot == "C" ~ "Control", 
                          Plot == "S" ~ "Saltwater", 
                          Plot == "F" ~ "Freshwater")) %>% #for some reason the FW files are coming in with FALSE not F
  rename(value = Value,
         sensor_id = Sensor_ID,
         location = Location
         ) %>%
  dplyr::select(datetime_est, research_name, plot, depth, sensor_id, location, value, contains("f_"))

write_rds(data_clean1, "~/GitHub/TEMPEST-1-porewater/data_do_not_commit/2021-2024_teros_EC_raw.rds")
```


### Raw data

```{r filter to pw grids}
pwsite_key <- readxl::read_excel("~/GitHub/tempest-system-level-analysis/data/for processing/porewater_sites_complete_key.xlsx") %>%
  select(Plot, Grid) %>%
  unique() %>%
  rename(plot = Plot,
         Location_Inventory = Grid) %>%
  mutate( plot = case_when(plot == "C" ~ "Control", 
                          plot == "SW" ~ "Saltwater", 
                          plot == "FW" ~ "Freshwater"))

grids <- pwsite_key %>%
  dplyr::distinct(Location_Inventory)

grids
```


Remove the teros where we don't have lysimeters
and filter out when lysimeter is out of service 

```{r qc1}

teros_qc1 <- data_clean1 %>% 
  filter((location %in% grids$Location_Inventory)) %>%
  mutate(variable = str_extract(research_name, "(?<=_)[^_]+(?=_)")) 
```

Let's first take a look at all the raw data. 
EC for ionic strength:
```{r, eval=FALSE}
## Plotly plot to look at individual sensors
q <- teros_qc1 %>%
  ggplot(aes(datetime_est, value, linetype= location, color = depth)) + 
  geom_line() + 
  facet_wrap(~plot, ncol = 1)



 plotly::ggplotly(q)
```



<!-- there are some crazy things happening in the control plot H6: -->
<!-- peaks with one data point really high SPC in the 5cm sensor at H6 -->
<!-- "2022-02-20 22:30" -->
<!-- "2022-02-21 09:00" -->
<!-- "2022-12-03 13:45" -->
<!-- "2023-02-25 02:00" -->
<!-- "2023-02-27 04:15" -->
<!-- "2023-03-01 01:30" -->
<!-- "2023-03-01 08:00" -->
<!-- "2023-12-09 05:45" -->
<!-- "2023-12-13 02:30" -->
<!-- "2023-12-13 03:30" -->
<!-- "2023-12-13 21:15" -->



<!-- ```{r scrub the sketch 2} -->
<!-- teros_qc2_EC <- teros_qc1 %>%  -->
<!--   mutate(value = case_when(research_name == "soil_EC_5cm" & plot =="Control" & Location_Inventory == "H6" & datetime_est %in% as.POSIXct(c("2022-02-20 22:30", "2022-02-21 09:00", "2022-12-03 13:45",  -->
<!--                                    "2023-02-25 02:00", "2023-02-27 04:15", "2023-03-01 01:30",  -->
<!--                                    "2023-03-01 08:00", "2023-12-09 05:45", "2023-12-13 02:30", -->
<!--                                    "2023-12-13 03:30", "2023-12-13 21:15"), tz = common_tz)  ~ NA, -->
<!--                            TRUE ~ value)) %>% -->
<!--     mutate(value = case_when(research_name == "soil_EC_30cm" & plot =="Saltwater" & Location_Inventory == "F4" & datetime_est %in% as.POSIXct(c("2023-01-17 03:35"), tz = common_tz)  ~ NA, -->
<!--                            TRUE ~ value)) %>% -->
<!--   mutate(value = case_when(research_name == "soil_EC_5cm" & plot =="Saltwater" & Location_Inventory == "C3" & datetime_est %in% as.POSIXct(c("2023-07-03 05:10"), tz = common_tz)  ~ NA, -->
<!--                            TRUE ~ value)) -->

<!-- ``` -->

```{r EC for SI}

EC_L2 <- teros_qc2_EC %>%
  filter(depth == "15") %>%
  filter(variable == "EC") %>%
  rename(EC_uScm = value,
         grid = Location_Inventory) %>%
  select(datetime_est, plot, grid, depth, research_name, EC_uScm) %>%
  mutate(plot = case_when(plot == "Estuarine-water" ~ "Saltwater",
                          TRUE ~ plot))
  
print(EC_L2)

write_csv(EC_L2, "~/GitHub/TEMPEST-1-porewater/processed data/TMP_PW_EC_15cm_2022-2023_L2.csv")

saveRDS(EC_L2, "~/GitHub/TEMPEST-1-porewater/processed data/TMP_PW_EC_15cm_2022-2023_L2.rds")
  
```
