---
title: "Soil Column DOC Wash Statistical Test"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 5.5
)

required_packages <- c(
  "readxl",
  "dplyr",
  "ggplot2",
  "lmerTest",
  "emmeans",
  "knitr"
)

missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]

if (length(missing_packages) > 0) {
  stop(
    "Install these packages before running this R Markdown file: ",
    paste(missing_packages, collapse = ", ")
  )
}

library(readxl)
library(dplyr)
library(ggplot2)
library(lmerTest)
library(emmeans)
library(knitr)

emmeans::emm_options(lmer.df = "satterthwaite")

format_p_value <- function(p_value) {
  ifelse(
    is.na(p_value),
    NA_character_,
    ifelse(p_value < 0.001, "< 0.001", paste0("= ", format(round(p_value, 3), nsmall = 3)))
  )
}
```

## Analysis Goal

This analysis tests the statement:

> DOC increased markedly in the ASW soil columns after 6 successive washes of low ionic strength water following the initial ASW pulse, with no significant differences in DOC across washes in the control or freshwater column experiments.

The primary analysis treats the dataset as repeated measures, because each soil column was measured across multiple washes. The model is fit separately for each treatment:

```text
DOC ~ wash timepoint + random intercept for Column
```

This assumes `Timepoint == 1` is the initial replicated column wash and that `Timepoint == 7` corresponds to the sixth low-ionic-strength wash after the initial ASW pulse. If the timepoint coding changes, update `post_six_start_timepoint`.

```{r paths-and-inputs}
data_candidates <- c(
  "~/GitHub/tempest_ionic_strength/Data/Exp 3 Column Experiment ColEx Data/ColEx_data_allreps_alltreatments_formatted.xlsx"
)

data_path <- data_candidates[file.exists(data_candidates)][1]

if (is.na(data_path)) {
  stop("Could not find ColEx_data_allreps_alltreatments_formatted.xlsx.")
}

output_dir <- if (dir.exists("vibe_coding")) {
  file.path("vibe_coding", "outputs")
} else {
  "outputs"
}

dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

post_six_start_timepoint <- 7

raw_doc <- read_excel(data_path)

doc_data <- raw_doc %>%
  filter(Column != "Treatment Water") %>%
  filter(Timepoint >= 1) %>%
  mutate(
    Treatment = factor(Treatment, levels = c("ASW", "Control", "Freshwater")),
    Timepoint = as.integer(Timepoint),
    Timepoint_f = factor(Timepoint),
    wash_phase = if_else(
      Timepoint >= post_six_start_timepoint,
      "Low-ionic wash 6 onward",
      "Initial pulse through low-ionic wash 5"
    ),
    wash_phase = factor(
      wash_phase,
      levels = c(
        "Initial pulse through low-ionic wash 5",
        "Low-ionic wash 6 onward"
      )
    ),
    log_doc_mgperL = log(doc_mgperL)
  )

doc_data %>%
  count(Treatment, Timepoint) %>%
  kable(caption = "Replicate count by treatment and wash timepoint.")
```

## Wash Summaries

```{r wash-summary}
wash_summary <- doc_data %>%
  group_by(Treatment, Timepoint) %>%
  summarise(
    n_columns = n(),
    mean_doc_mgperL = mean(doc_mgperL),
    sd_doc_mgperL = sd(doc_mgperL),
    se_doc_mgperL = sd_doc_mgperL / sqrt(n_columns),
    min_doc_mgperL = min(doc_mgperL),
    max_doc_mgperL = max(doc_mgperL),
    .groups = "drop"
  )

write.csv(
  wash_summary,
  file.path(output_dir, "soil_column_DOC_wash_summary.csv"),
  row.names = FALSE
)

wash_summary %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  kable(caption = "Observed DOC summary by treatment and wash.")
```

## Figure 1. DOC Across Successive Washes

```{r wash-figure}
treatment_colors <- c(
  "ASW" = "#B04A5A",
  "Control" = "#4B5563",
  "Freshwater" = "#1F9E89"
)

phase_shade <- data.frame(
  Treatment = factor(levels(doc_data$Treatment), levels = levels(doc_data$Treatment)),
  xmin = post_six_start_timepoint - 0.5,
  xmax = Inf,
  ymin = -Inf,
  ymax = Inf
)

doc_timecourse_plot <- ggplot() +
  geom_rect(
    data = phase_shade,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = "#EEE8D5",
    alpha = 0.6
  ) +
  geom_line(
    data = doc_data,
    aes(x = Timepoint, y = doc_mgperL, group = Column, color = Treatment),
    alpha = 0.35,
    linewidth = 0.5
  ) +
  geom_point(
    data = doc_data,
    aes(x = Timepoint, y = doc_mgperL, color = Treatment),
    alpha = 0.65,
    size = 1.9
  ) +
  geom_errorbar(
    data = wash_summary,
    aes(
      x = Timepoint,
      ymin = mean_doc_mgperL - se_doc_mgperL,
      ymax = mean_doc_mgperL + se_doc_mgperL,
      color = Treatment
    ),
    width = 0.18,
    linewidth = 0.7
  ) +
  geom_line(
    data = wash_summary,
    aes(x = Timepoint, y = mean_doc_mgperL, group = Treatment, color = Treatment),
    linewidth = 1.2
  ) +
  geom_point(
    data = wash_summary,
    aes(x = Timepoint, y = mean_doc_mgperL, color = Treatment),
    size = 2.8
  ) +
  geom_vline(
    xintercept = post_six_start_timepoint - 0.5,
    linetype = "dashed",
    linewidth = 0.5,
    color = "#555555"
  ) +
  facet_wrap(~ Treatment, nrow = 1) +
  scale_color_manual(values = treatment_colors, guide = "none") +
  scale_x_continuous(breaks = sort(unique(doc_data$Timepoint))) +
  labs(
    x = "Wash timepoint",
    y = "DOC (mg/L)",
    title = "DOC across successive soil-column washes",
    subtitle = "Thin lines are individual columns; larger points and error bars are treatment means +/- SE.\nShading starts at low-ionic wash 6 onward."
  ) +
  theme_bw(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 11),
    plot.margin = margin(8, 14, 8, 8),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "#F3F4F6", color = "#D1D5DB"),
    strip.text = element_text(face = "bold")
  )

doc_timecourse_plot

ggsave(
  file.path(output_dir, "soil_column_DOC_by_wash_repeated_columns.png"),
  doc_timecourse_plot,
  width = 10,
  height = 5.5,
  dpi = 300
)

ggsave(
  file.path(output_dir, "soil_column_DOC_by_wash_repeated_columns.pdf"),
  doc_timecourse_plot,
  width = 10,
  height = 5.5
)
```

## Figure 2. Planned Phase Contrast

```{r phase-figure}
phase_column_means <- doc_data %>%
  group_by(Treatment, Column, wash_phase) %>%
  summarise(
    mean_doc_mgperL = mean(doc_mgperL),
    .groups = "drop"
  )

phase_summary <- phase_column_means %>%
  group_by(Treatment, wash_phase) %>%
  summarise(
    n_columns = n(),
    phase_mean_doc_mgperL = mean(mean_doc_mgperL),
    phase_sd_doc_mgperL = sd(mean_doc_mgperL),
    .groups = "drop"
  ) %>%
  mutate(
    mean_doc_mgperL = phase_mean_doc_mgperL,
    sd_doc_mgperL = phase_sd_doc_mgperL,
    se_doc_mgperL = sd_doc_mgperL / sqrt(n_columns),
    phase_mean_doc_mgperL = NULL,
    phase_sd_doc_mgperL = NULL
  )

write.csv(
  phase_summary,
  file.path(output_dir, "soil_column_DOC_phase_summary.csv"),
  row.names = FALSE
)

doc_phase_plot <- ggplot() +
  geom_line(
    data = phase_column_means,
    aes(x = wash_phase, y = mean_doc_mgperL, group = Column, color = Treatment),
    alpha = 0.35,
    linewidth = 0.6
  ) +
  geom_point(
    data = phase_column_means,
    aes(x = wash_phase, y = mean_doc_mgperL, color = Treatment),
    alpha = 0.75,
    size = 2
  ) +
  geom_errorbar(
    data = phase_summary,
    aes(
      x = wash_phase,
      ymin = mean_doc_mgperL - se_doc_mgperL,
      ymax = mean_doc_mgperL + se_doc_mgperL,
      color = Treatment
    ),
    width = 0.12,
    linewidth = 0.8
  ) +
  geom_point(
    data = phase_summary,
    aes(x = wash_phase, y = mean_doc_mgperL, color = Treatment),
    size = 3.2
  ) +
  facet_wrap(~ Treatment, nrow = 1) +
  scale_color_manual(values = treatment_colors, guide = "none") +
  scale_x_discrete(
    labels = c(
      "Initial pulse through low-ionic wash 5" = "Initial pulse\nthrough wash 5",
      "Low-ionic wash 6 onward" = "Low-ionic wash 6\nonward"
    )
  ) +
  labs(
    x = NULL,
    y = "Column mean DOC (mg/L)",
    title = "DOC phase contrast before and after the sixth low-ionic-strength wash",
    subtitle = "Thin lines connect same-column phase means; larger points and error bars are treatment means +/- SE."
  ) +
  theme_bw(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 11),
    axis.text.x = element_text(lineheight = 0.95),
    plot.margin = margin(8, 14, 8, 8),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "#F3F4F6", color = "#D1D5DB"),
    strip.text = element_text(face = "bold")
  )

doc_phase_plot

ggsave(
  file.path(output_dir, "soil_column_DOC_phase_contrast.png"),
  doc_phase_plot,
  width = 10,
  height = 5.5,
  dpi = 300
)

ggsave(
  file.path(output_dir, "soil_column_DOC_phase_contrast.pdf"),
  doc_phase_plot,
  width = 10,
  height = 5.5
)
```

## Primary Repeated-Measures Tests

These models account for repeated measurements from the same column by including a random intercept for `Column`. The raw-scale model keeps the units interpretable as mg/L DOC.

```{r repeated-measures-tests}
response_specs <- data.frame(
  response_column = c("doc_mgperL", "log_doc_mgperL"),
  response_scale = c("raw_mg_per_L", "log_mg_per_L"),
  stringsAsFactors = FALSE
)

timepoint_tests <- data.frame()
phase_tests <- data.frame()
phase_emmeans <- data.frame()
phase_contrasts <- data.frame()

for (response_index in seq_len(nrow(response_specs))) {
  current_response <- response_specs$response_column[response_index]
  current_scale <- response_specs$response_scale[response_index]

  for (current_treatment in levels(doc_data$Treatment)) {
    current_data <- doc_data %>%
      filter(Treatment == current_treatment)

    timepoint_formula <- as.formula(
      paste(current_response, "~ Timepoint_f + (1 | Column)")
    )

    timepoint_fit <- lmer(
      timepoint_formula,
      data = current_data,
      REML = FALSE
    )

    timepoint_anova <- as.data.frame(anova(timepoint_fit, type = 3))
    timepoint_row <- which(rownames(timepoint_anova) == "Timepoint_f")

    timepoint_tests <- bind_rows(
      timepoint_tests,
      data.frame(
        Treatment = current_treatment,
        response_scale = current_scale,
        model = paste(current_response, "~ factor(Timepoint) + (1 | Column)"),
        test = "overall_wash_timepoint_effect",
        numerator_df = timepoint_anova$NumDF[timepoint_row],
        denominator_df = timepoint_anova$DenDF[timepoint_row],
        f_value = timepoint_anova$`F value`[timepoint_row],
        p_value = timepoint_anova$`Pr(>F)`[timepoint_row]
      )
    )

    phase_formula <- as.formula(
      paste(current_response, "~ wash_phase + (1 | Column)")
    )

    phase_fit <- lmer(
      phase_formula,
      data = current_data,
      REML = FALSE
    )

    phase_anova <- as.data.frame(anova(phase_fit, type = 3))
    phase_row <- which(rownames(phase_anova) == "wash_phase")

    phase_tests <- bind_rows(
      phase_tests,
      data.frame(
        Treatment = current_treatment,
        response_scale = current_scale,
        model = paste(current_response, "~ wash_phase + (1 | Column)"),
        test = "planned_phase_effect",
        numerator_df = phase_anova$NumDF[phase_row],
        denominator_df = phase_anova$DenDF[phase_row],
        f_value = phase_anova$`F value`[phase_row],
        p_value = phase_anova$`Pr(>F)`[phase_row]
      )
    )

    current_phase_emmeans <- as.data.frame(emmeans(phase_fit, ~ wash_phase)) %>%
      mutate(
        Treatment = current_treatment,
        response_scale = current_scale
      )

    phase_emmeans <- bind_rows(phase_emmeans, current_phase_emmeans)

    current_phase_contrast <- as.data.frame(
      pairs(emmeans(phase_fit, ~ wash_phase), adjust = "none")
    ) %>%
      mutate(
        Treatment = current_treatment,
        response_scale = current_scale,
        post_six_minus_early_estimate = -estimate,
        post_six_to_early_ratio = if_else(
          current_scale == "log_mg_per_L",
          exp(post_six_minus_early_estimate),
          NA_real_
        )
      )

    phase_contrasts <- bind_rows(phase_contrasts, current_phase_contrast)
  }
}

timepoint_tests <- timepoint_tests %>%
  group_by(response_scale) %>%
  mutate(p_value_bh = p.adjust(p_value, method = "BH")) %>%
  ungroup()

phase_tests <- phase_tests %>%
  group_by(response_scale) %>%
  mutate(p_value_bh = p.adjust(p_value, method = "BH")) %>%
  ungroup()

phase_contrasts <- phase_contrasts %>%
  group_by(response_scale) %>%
  mutate(p_value_bh = p.adjust(p.value, method = "BH")) %>%
  ungroup()

write.csv(
  timepoint_tests,
  file.path(output_dir, "soil_column_DOC_lmer_timepoint_tests.csv"),
  row.names = FALSE
)

write.csv(
  phase_tests,
  file.path(output_dir, "soil_column_DOC_lmer_phase_tests.csv"),
  row.names = FALSE
)

write.csv(
  phase_emmeans,
  file.path(output_dir, "soil_column_DOC_lmer_phase_emmeans.csv"),
  row.names = FALSE
)

write.csv(
  phase_contrasts,
  file.path(output_dir, "soil_column_DOC_lmer_phase_contrasts.csv"),
  row.names = FALSE
)

timepoint_tests %>%
  filter(response_scale == "raw_mg_per_L") %>%
  mutate(
    f_value = round(f_value, 3),
    p_value = format_p_value(p_value),
    p_value_bh = format_p_value(p_value_bh)
  ) %>%
  kable(caption = "Primary repeated-measures test: raw DOC wash-timepoint effect within each treatment.")

phase_contrasts %>%
  filter(response_scale == "raw_mg_per_L") %>%
  transmute(
    Treatment,
    contrast,
    post_six_minus_early_mg_per_L = round(post_six_minus_early_estimate, 3),
    SE = round(SE, 3),
    df = round(df, 1),
    t_ratio = round(t.ratio, 3),
    p_value = format_p_value(p.value),
    p_value_bh = format_p_value(p_value_bh)
  ) %>%
  kable(caption = "Planned phase contrast from the repeated-measures model. Positive estimates mean DOC was higher at low-ionic wash 6 onward.")
```

## Sensitivity Check: Log-DOC Repeated-Measures Tests

The log-DOC model checks whether the conclusions are sensitive to the strong mean-variance relationship visible in the ASW data.

```{r log-sensitivity}
timepoint_tests %>%
  filter(response_scale == "log_mg_per_L") %>%
  mutate(
    f_value = round(f_value, 3),
    p_value = format_p_value(p_value),
    p_value_bh = format_p_value(p_value_bh)
  ) %>%
  kable(caption = "Sensitivity test: log-DOC repeated-measures wash-timepoint effect within each treatment.")

phase_contrasts %>%
  filter(response_scale == "log_mg_per_L") %>%
  transmute(
    Treatment,
    contrast,
    post_six_minus_early_log_difference = round(post_six_minus_early_estimate, 3),
    post_six_to_early_ratio = round(post_six_to_early_ratio, 3),
    SE = round(SE, 3),
    df = round(df, 1),
    t_ratio = round(t.ratio, 3),
    p_value = format_p_value(p.value),
    p_value_bh = format_p_value(p_value_bh)
  ) %>%
  kable(caption = "Sensitivity test: log-DOC planned phase contrast. Ratios above 1 mean DOC was higher at low-ionic wash 6 onward.")
```

## Sensitivity Check: Ordinary One-Way ANOVA

The table below repeats the across-wash test with a simpler one-way ANOVA that treats each measurement as independent. This matches a common visual summary test, but it does not account for repeated measurements from the same column.

```{r ordinary-anova-sensitivity}
ordinary_anova_tests <- data.frame()

for (current_treatment in levels(doc_data$Treatment)) {
  current_data <- doc_data %>%
    filter(Treatment == current_treatment)

  ordinary_fit <- aov(doc_mgperL ~ Timepoint_f, data = current_data)
  ordinary_anova <- summary(ordinary_fit)[[1]]
  ordinary_row <- which(rownames(ordinary_anova) == "Timepoint_f")

  ordinary_anova_tests <- bind_rows(
    ordinary_anova_tests,
    data.frame(
      Treatment = current_treatment,
      model = "doc_mgperL ~ factor(Timepoint)",
      test = "ordinary_one_way_anova_timepoint_effect",
      numerator_df = ordinary_anova$Df[ordinary_row],
      denominator_df = ordinary_anova$Df[ordinary_row + 1],
      f_value = ordinary_anova$`F value`[ordinary_row],
      p_value = ordinary_anova$`Pr(>F)`[ordinary_row]
    )
  )
}

ordinary_anova_tests <- ordinary_anova_tests %>%
  mutate(p_value_bh = p.adjust(p_value, method = "BH"))

write.csv(
  ordinary_anova_tests,
  file.path(output_dir, "soil_column_DOC_ordinary_anova_timepoint_tests.csv"),
  row.names = FALSE
)

ordinary_anova_tests %>%
  mutate(
    f_value = round(f_value, 3),
    p_value = format_p_value(p_value),
    p_value_bh = format_p_value(p_value_bh)
  ) %>%
  kable(caption = "Sensitivity check: ordinary one-way ANOVA across washes within each treatment.")
```

## Interpretation for the Manuscript Sentence

```{r interpretation-table}
raw_timepoint_results <- timepoint_tests %>%
  filter(response_scale == "raw_mg_per_L") %>%
  select(Treatment, repeated_measures_wash_p = p_value, repeated_measures_wash_p_bh = p_value_bh)

raw_phase_results <- phase_contrasts %>%
  filter(response_scale == "raw_mg_per_L") %>%
  transmute(
    Treatment,
    post_six_minus_early_mg_per_L = post_six_minus_early_estimate,
    repeated_measures_phase_p = p.value,
    repeated_measures_phase_p_bh = p_value_bh
  )

ordinary_results <- ordinary_anova_tests %>%
  select(Treatment, ordinary_anova_wash_p = p_value, ordinary_anova_wash_p_bh = p_value_bh)

interpretation_table <- raw_timepoint_results %>%
  left_join(raw_phase_results, by = "Treatment") %>%
  left_join(ordinary_results, by = "Treatment") %>%
  mutate(
    post_six_minus_early_mg_per_L = round(post_six_minus_early_mg_per_L, 3),
    repeated_measures_wash_p = format_p_value(repeated_measures_wash_p),
    repeated_measures_wash_p_bh = format_p_value(repeated_measures_wash_p_bh),
    repeated_measures_phase_p = format_p_value(repeated_measures_phase_p),
    repeated_measures_phase_p_bh = format_p_value(repeated_measures_phase_p_bh),
    ordinary_anova_wash_p = format_p_value(ordinary_anova_wash_p),
    ordinary_anova_wash_p_bh = format_p_value(ordinary_anova_wash_p_bh)
  )

interpretation_table %>%
  kable(caption = "Summary of evidence relevant to the manuscript sentence.")
```

The repeated-measures model supports a strong ASW increase after the sixth low-ionic-strength wash. Freshwater shows no significant repeated-measures wash effect. The Control result depends on the wording of the claim: the planned pre/post phase contrast is not significant, but the full repeated-measures test across all individual washes is significant. Therefore, the sentence should not say that Control had no significant differences across individual washes if the repeated-measures model is the primary test.

A statistically safer version is:

> DOC increased markedly in the ASW soil columns at low-ionic-strength wash 6 onward after the initial ASW pulse. In contrast, the analogous pre/post phase comparison was not significant in the control or freshwater column experiments, although the full repeated-measures test across individual washes detected smaller wash-to-wash variation in the control columns.
