---
title: "Porewater DOC Story Figs"
author: "AMP"
date: "`r Sys.Date()`"
output: html_document
---

# Prep 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotrix)
getwd()
```

```{r event and study dates}
endstudydate = lubridate::as_date("2024-01-31")
startstudydate = lubridate::as_date("2022-05-01")

#water delivery just dates:
water2022 = lubridate::as_date("2022-06-22")
water2023_1 = lubridate::as_date("2023-06-06")
water2023_2 = lubridate::as_date("2023-06-07")

#water delivery dates and times:
WaterDeliveryStart2022 = as.POSIXct("2022-06-22 05:30:00", tz = "EST")
WaterDeliveryStop2022 = as.POSIXct("2022-06-22 14:30:00", tz = "EST")

WaterDeliveryStart1 = as.POSIXct("2023-06-06 05:30:00", tz = "EST")
WaterDeliveryStop1 = as.POSIXct("2023-06-06 14:30:00", tz = "EST")

WaterDeliveryStart2 = as.POSIXct("2023-06-07 05:30:00", tz = "EST")
WaterDeliveryStop2 = as.POSIXct("2023-06-07 14:30:00", tz = "EST")
```

```{r load data}
pw_grids <- readRDS("./processed data/TMP_PW_GRIDSONLY_NPOC_TDN_L1_May2022-Dec2023.rds")

pw_pooled <- readRDS("./processed data/TMP_PW_POOL_NPOC_TDN_L1_May2022-Dec2023.rds")

eems_pooled <- readRDS("./processed data/TMP_PW_POOLED_CDOM_L1_May2022-Dec2023.rds")
eems_grids <- readRDS("./processed data/TMP_PW_grids_CDOM_L1_May2022-Dec2023.rds")

```

```{r set up for making plots}
plot_order <- c('Control Plot', 'Freshwater Plot','Estuarine-water Plot')
Anyas_colors_alpha_order = c("springgreen2", "violetred2", "cyan2")

```

# DOC 

```{r prep data for plotting}

pw_plotting <- pw_grids %>%
  filter(!is.na(doc_mg_l)) %>%
  filter(sample_name != "Lysimeter Empty") %>%
  select(doc_mg_l, tdn_mg_l, plot, Grid, date, time, datetime, Group) %>%
  dplyr::group_by(plot, Group) %>% 
    arrange(plot, date, time) 
  
```


```{r doc summary statistics}

pw_sum_stats <- pw_plotting %>%
   group_by(Group, plot) %>%
   summarise(n=n(),
              doc_mg_l_mean = round(mean(doc_mg_l, na.rm = TRUE), 2),
              doc_mg_l_median = round(median(doc_mg_l, na.rm = TRUE), 2),
              doc_mg_l_sd = round(sd(doc_mg_l, na.rm = TRUE), 2),
              doc_mg_l_se = round(std.error(doc_mg_l, na.rm = TRUE), 2),
              Q1_doc = round(quantile(doc_mg_l, 0.25, na.rm= TRUE),2),
              Q3_doc = round(quantile(doc_mg_l, 0.75, na.rm= TRUE),2),
              min_doc = round(min(doc_mg_l, na.rm= TRUE),2),
              max_doc = round(max(doc_mg_l, na.rm= TRUE),2),
              tdn_mg_l_mean = round(mean(tdn_mg_l, na.rm = TRUE),2),
              tdn_mg_l_median = round(median(tdn_mg_l, na.rm = TRUE),2),
              tdn_mg_l_sd = round(sd(tdn_mg_l, na.rm = TRUE), 2),
              tdn_mg_l_se = round(std.error(tdn_mg_l, na.rm = TRUE), 2),
              Q1_tdn = round(quantile(tdn_mg_l, 0.25, na.rm= TRUE),2),
              Q3_tdn = round(quantile(tdn_mg_l, 0.75, na.rm= TRUE),2),
              min_tdn = round(min(tdn_mg_l, na.rm= TRUE),2),
              max_tdn = round(max(tdn_mg_l, na.rm= TRUE),2),
            .groups = "drop") %>%
    arrange(plot, Group) 
```

 
# RANDOM EXPLORATORY 

Boxplots and Violins to look at the data distribution: 

```{r plot distribution of data}

pw_plotting %>%
    group_by(Group) %>%
# filter(datetime > WaterDeliveryStop2022 & datetime < WaterDeliveryStart1) %>%
  ggplot() +
  #     # Water 2022
 # annotate("rect", xmin = WaterDeliveryStart2022, xmax = WaterDeliveryStop2022, ymin = -Inf, ymax = Inf, fill = "blue")+
  #  # Water 1
 # annotate("rect", xmin = WaterDeliveryStart1, xmax = WaterDeliveryStop1, ymin = -Inf, ymax = Inf, fill = "blue") +
  # #Water 2
 # annotate("rect", xmin = WaterDeliveryStart2 , xmax = WaterDeliveryStop2, ymin = -Inf, ymax = Inf, fill = "blue") +
  #scale_x_datetime(minor_breaks= waiver(),date_minor_breaks = "1 month", date_breaks="2 months",date_labels= '%b %y')+
  geom_violin(aes(x = factor(Group), y =doc_mg_l, color= plot), drop= FALSE)+
  # geom_line(aes(x = plot_date, y =doc_mg_l, color = Plot))+
  ylab("DOC mgC/L")+
  xlab("Date")+
  ylim(0,100)+
  scale_color_manual(values=Anyas_colors_alpha_order) +
  theme_classic() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
         axis.text.x= element_text(size=10, color= "black",angle = 45, hjust = 1)) 
```


levels = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25)

 mutate(plotting_group= case_when(Group == "5-2022" ~ "1",
                                   Group == "2022-06-15" ~ "2",
                                   Group == "2022-06-20" ~ "2",
                                   Group == "2022-06-22" ~ "3", # Event Day
                                   Group == "2022-06-23" ~ "4",
                                   Group == "2022-06-24" ~ "5",
                                   Group == "7-2022" ~ "6",
                                   Group == "8-2022" ~ "7",
                                   Group == "9-2022" ~ "8",
                                   Group == "10-2022" ~ "9",
                                   Group == "11-2022" ~ "10",
                                   Group == "12-2022" ~ "11",
                                   Group == "2-2023" ~ "12",
                                   Group == "4-2023" ~ "13",
                                   Group == "5-2023" ~ "14",
                                   Group == "2023-06-03" ~ "15",
                                   Group == "2023-06-05" ~ "15", 
                                   Group == "2023-06-06" ~ "16", #Event Day
                                   Group == "2023-06-07" ~ "17", #Event Day
                                   Group == "2023-06-08" ~ "18", 
                                   Group == "2023-06-12" ~ "19",
                                   Group == "2023-06-14" ~ "19",
                                   Group == "2023-06-17" ~ "19",
                                   Group == "2023-06-27" ~ "20",
                                   Group == "2023-06-28" ~ "20",
                                   Group == "7-2023" ~ "21",
                                   Group == "8-2023" ~ "22",
                                   Group == "10-2023" ~ "23",
                                   Group == "11-2023" ~ "24",
                                   Group == "12-2023" ~ "25",
                                   TRUE ~ NA)) %>%
         mutate(plotting_group= factor(plotting_group, levels = levels))

all_the_doc <- pw2022event %>%
  full_join(pw2022) %>%
  full_join(pw2023) %>%
  mutate(date = as.POSIXct(date)) %>%
  mutate(Plot = stringr::str_replace(Plot,"SW","Estuarine-water Plot")) %>%
  mutate(Plot = stringr::str_replace(Plot,"FW","Freshwater Plot")) %>%
  mutate(Plot = stringr::str_replace(Plot,"C","Control Plot")) %>%
  mutate(date_formatted = paste0(lubridate::year(date), "-", lubridate::month(date), "-", lubridate::day(date))) %>%
  mutate(date_formatted = lubridate::as_date(date_formatted)) %>%
  mutate(plot_date = case_when( lubridate::month(date_formatted) == lubridate::month(event_2023) ~ paste0(lubridate::year(date), "-", lubridate::month(date), "-", lubridate::day(date)),
                               lubridate::month(date_formatted) == lubridate::month(event_2023) ~ paste0(lubridate::year(date), "-", lubridate::month(date), "-", lubridate::day(date)),
          TRUE ~  paste0(lubridate::year(date), "-", lubridate::month(date), "-", "01")),
        month = lubridate::month(date),
        year = lubridate::year(date)) %>%
  mutate(plot_date = lubridate::with_tz(plot_date, "UTC")) %>%
  group_by(Plot, plot_date, month, year) %>% 
  filter(Grid != "POOL") %>%
  filter(!npoc_flag %in% "omitted for high dilution and blank values") %>%
  filter(!tdn_flag %in% "omitted for high dilution and blank values") %>%
  filter(!sample_name %in% "Lysimeter Empty") %>%
  summarise(n=n(),
              doc_mg_l_sd = round(sd(doc_mg_l, na.rm = TRUE), 2),
            doc_mg_l = round(mean(doc_mg_l, na.rm = TRUE), 2),
            tdn_mg_l_sd = round(sd(tdn_mg_l, na.rm = TRUE), 2),
            tdn_mg_l = round(mean(tdn_mg_l, na.rm = TRUE),2),
            .groups = "drop") %>%
    arrange(Plot, year, month)

all_eems <- eems %>%
   mutate(date = as.POSIXct(date)) %>%
  mutate(Plot = stringr::str_replace(Plot,"SW","Estuarine-water Plot")) %>%
  mutate(Plot = stringr::str_replace(Plot,"FW","Freshwater Plot")) %>%
  mutate(Plot = stringr::str_replace(Plot,"C","Control Plot")) %>%
  mutate(date_formatted = paste0(lubridate::year(date), "-", lubridate::month(date), "-", lubridate::day(date))) %>%
  mutate(date_formatted = lubridate::as_date(date_formatted)) %>%
  mutate(plot_date = case_when( lubridate::month(date_formatted) == lubridate::month(event_2023) ~ paste0(lubridate::year(date), "-", lubridate::month(date), "-", lubridate::day(date)),
                               lubridate::month(date_formatted) == lubridate::month(event_2023) ~ paste0(lubridate::year(date), "-", lubridate::month(date), "-", lubridate::day(date)),
          TRUE ~  paste0(lubridate::year(date), "-", lubridate::month(date), "-", "01")),
        month = lubridate::month(date),
        year = lubridate::year(date)) %>%
  mutate(plot_date = lubridate::with_tz(plot_date, "UTC")) %>%
  group_by(Plot, plot_date, month, year) %>% 
    arrange(Plot, year, month)
    

# DOC Plots 

Figure X for Take Home Message: There is a prolonged and sustained release of carbon following exposure to salinity.

```{r timeseries plots, warning=FALSE}

plot <- pw_sum_stats %>%
  ggplot() +
       # Water 2022
  geom_vline("line", xintercept = water2022,  color = "blue")+
  annotate("text", label= "2022 Event", x = lubridate::as_date("2022-06-24"), y = 100, size=6, color = "blue") +
    # Water 1
  geom_vline("line", xintercept = water2023_1,  color = "blue")+
   #Water 2
  geom_vline("line", xintercept = water2023_2,  color = "blue")+
   annotate("text", label= "2023 Events", x = lubridate::as_date("2023-06-17"), y = 100, size=6, color = "blue") +
 scale_x_date(minor_breaks= waiver(),date_minor_breaks = "1 month", date_breaks="2 months",date_labels= '%b %Y')+
 geom_errorbar(aes(x= Group, ymin= min_doc, ymax=  max_doc, color = plot))+
  geom_point(aes(x= Group, y= doc_mg_l_mean, color = plot), size = 2)+
 labs(x = "Date", y = "DOC mgC/L", color = "Plot") +
 # ylim(0,100)+
  scale_color_manual(values=Anyas_colors_alpha_order) +
  theme_classic() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
         axis.text.x= element_text(size=12, color= "black",angle = 45, hjust = 1),
        axis.text =  element_text(size = 14, color= "black"),
        axis.title = element_text(size = 14, color= "black")) 

plot

cowplot::save_plot("./figures/doc_timeseries_2022-2023.pdf", plot, dpi=300, base_aspect_ratio = 2:1)
```

# EEMS

```{r doc summary statistics}

eems_sum_stats <- eems_pooled %>%
   dplyr::bind_rows(eems_grids) %>%
  group_by(plot, Group) %>%
  summarise(n=n(),
              Sr_mean = round(mean(Sr, na.rm = TRUE), 2),
              Sr_min = round(min(Sr, na.rm= TRUE),2),
              Sr_max = round(max(Sr, na.rm= TRUE),2),
              Sr_mean = round(mean(Sr, na.rm = TRUE), 2),
              Sr_min = round(min(Sr, na.rm= TRUE),2),
              Sr_max = round(max(Sr, na.rm= TRUE),2),
              FI_mean = round(mean(FI, na.rm = TRUE), 2),
              FI_min = round(min(FI, na.rm= TRUE),2),
              FI_max = round(max(FI, na.rm= TRUE),2),
              FI_mean = round(mean(FI, na.rm = TRUE), 2),
              FI_min = round(min(FI, na.rm= TRUE),2),
              FI_max = round(max(FI, na.rm= TRUE),2),
            .groups = "drop") %>%
    arrange(plot, Group)

```

```{r timeseries plots, warning=FALSE}

Sr_plot <- eems_sum_stats %>%
  ggplot() +
       # Water 2022
  geom_vline("line", xintercept = water2022,  color = "blue")+
  annotate("text", label= "2022 Event", x = lubridate::as_date("2022-08-08"), y = 2, size=6, color = "blue") +
    # Water 1
  geom_vline("line", xintercept = water2023_1,  color = "blue")+
   #Water 2
  geom_vline("line", xintercept = water2023_2,  color = "blue")+
   annotate("text", label= "2023 Events", x = lubridate::as_date("2023-06-17"), y = 2, size=6, color = "blue") +
 scale_x_date(minor_breaks= waiver(),date_minor_breaks = "1 month", date_breaks="2 months",date_labels= '%b %Y')+
 geom_errorbar(aes(x= Group, ymin= Sr_min, ymax=  Sr_max, color = plot))+
  geom_point(aes(x= Group, y= Sr_mean, color = plot), size = 2)+
 labs(x = "Date", y = "Slope Ratio", color = "Plot") +
  scale_color_manual(values=Anyas_colors_alpha_order) +
  theme_classic() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
         axis.text.x= element_text(size=12, color= "black",angle = 45, hjust = 1),
        axis.text =  element_text(size = 14, color= "black"),
        axis.title = element_text(size = 14, color= "black")) 

Sr_plot

cowplot::save_plot("./figures/Sr_timeseries_2022-2023.pdf", Sr_plot, dpi=300, base_aspect_ratio = 2:1)

FI_plot <- eems_sum_stats %>%
  ggplot() +
       # Water 2022
  geom_vline("line", xintercept = water2022,  color = "blue")+
  annotate("text", label= "2022 Event", x = lubridate::as_date("2022-08-08"), y = 2, size=6, color = "blue") +
    # Water 1
  geom_vline("line", xintercept = water2023_1,  color = "blue")+
   #Water 2
  geom_vline("line", xintercept = water2023_2,  color = "blue")+
   annotate("text", label= "2023 Events", x = lubridate::as_date("2023-06-17"), y = 2, size=6, color = "blue") +
 scale_x_date(minor_breaks= waiver(),date_minor_breaks = "1 month", date_breaks="2 months",date_labels= '%b %Y')+
 geom_errorbar(aes(x= Group, ymin= FI_min, ymax=  FI_max, color = plot))+
  geom_point(aes(x= Group, y= FI_mean, color = plot), size = 2)+
 labs(x = "Date", y = "Fluorescence Index", color = "Plot") +
  scale_color_manual(values=Anyas_colors_alpha_order) +
  theme_classic() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
         axis.text.x= element_text(size=12, color= "black",angle = 45, hjust = 1),
        axis.text =  element_text(size = 14, color= "black"),
        axis.title = element_text(size = 14, color= "black")) 

FI_plot

cowplot::save_plot("./figures/FI_timeseries_2022-2023.pdf", FI_plot, dpi=300, base_aspect_ratio = 2:1)
```